### 2013-02-07 latex/sig-alternate.tex
Latex template up

% This is "sig-alternate.tex" V2.0 May 2012
% This file should be compiled with V2.5 of "sig-alternate.cls" May 2012
%
% This example file demonstrates the use of the 'sig-alternate.cls'
% V2.5 LaTeX2e document class file. It is for those submitting
% articles to ACM Conference Proceedings WHO DO NOT WISH TO
% STRICTLY ADHERE TO THE SIGS (PUBS-BOARD-ENDORSED) STYLE.
% The 'sig-alternate.cls' file will produce a similar-looking,
% albeit, 'tighter' paper resulting in, invariably, fewer pages.
%
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V2.5) produces:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) NO page numbers
%
% as against the acm_proc_article-sp.cls file which
% DOES NOT produce 1) thru' 3) above.
%
% Using 'sig-alternate.cls' you have control, however, from within
% the source .tex file, over both the CopyrightYear
% (defaulted to 200X) and the ACM Copyright Data
% (defaulted to X-XXXXX-XX-X/XX/XX).
% e.g.
% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the copyright line.
%
% ---------------------------------------------------------------------------------------------------------------
% This .tex source is an example which *does* use
% the .bib file (from which the .bbl file % is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission, you *NEED* to 'insert'
% your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% ================= IF YOU HAVE QUESTIONS =======================
% Questions regarding the SIGS styles, SIGS policies and
% procedures, Conferences etc. should be sent to
% Adrienne Griscti (griscti@acm.org)
%
% Technical questions _only_ to
% Gerald Murray (murray@hq.acm.org)
% ===============================================================
%
% For tracking purposes - this is V2.0 - May 2012

\documentclass{sig-alternate}

\begin{document}
%
% --- Author Metadata here ---
\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{Alternate {\ttlit ACM} SIG Proceedings Paper in LaTeX
Format\titlenote{(Produces the permission block, and
copyright information). For use with
SIG-ALTERNATE.CLS. Supported by ACM.}}
\subtitle{[Extended Abstract]
\titlenote{A full version of this paper is available as
\textit{Author's Guide to Preparing ACM SIG Proceedings Using
\LaTeX$2_\epsilon$\ and BibTeX} at
\texttt{www.acm.org/eaddress.htm}}}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

\numberofauthors{8} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Ben Trovato\titlenote{Dr.~Trovato insisted his name be first.}\\
       \affaddr{Institute for Clarity in Documentation}\\
       \affaddr{1932 Wallamaloo Lane}\\
       \affaddr{Wallamaloo, New Zealand}\\
       \email{trovato@corporation.com}
% 2nd. author
\alignauthor
G.K.M. Tobin\titlenote{The secretary disavows
any knowledge of this author's actions.}\\
       \affaddr{Institute for Clarity in Documentation}\\
       \affaddr{P.O. Box 1212}\\
       \affaddr{Dublin, Ohio 43017-6221}\\
       \email{webmaster@marysville-ohio.com}
% 3rd. author
\alignauthor Lars Th{\o}rv{\"a}ld\titlenote{This author is the
one who did all the really hard work.}\\
       \affaddr{The Th{\o}rv{\"a}ld Group}\\
       \affaddr{1 Th{\o}rv{\"a}ld Circle}\\
       \affaddr{Hekla, Iceland}\\
       \email{larst@affiliation.org}
\and  % use '\and' if you need 'another row' of author names
% 4th. author
\alignauthor Lawrence P. Leipuner\\
       \affaddr{Brookhaven Laboratories}\\
       \affaddr{Brookhaven National Lab}\\
       \affaddr{P.O. Box 5000}\\
       \email{lleipuner@researchlabs.org}
% 5th. author
\alignauthor Sean Fogarty\\
       \affaddr{NASA Ames Research Center}\\
       \affaddr{Moffett Field}\\
       \affaddr{California 94035}\\
       \email{fogartys@amesres.org}
% 6th. author
\alignauthor Charles Palmer\\
       \affaddr{Palmer Research Laboratories}\\
       \affaddr{8600 Datapoint Drive}\\
       \affaddr{San Antonio, Texas 78229}\\
       \email{cpalmer@prl.com}
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
\date{30 July 1999}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

\maketitle
\begin{abstract}
This paper provides a sample of a \LaTeX\ document which conforms,
somewhat loosely, to the formatting guidelines for
ACM SIG Proceedings. It is an {\em alternate} style which produces
a {\em tighter-looking} paper and was designed in response to
concerns expressed, by authors, over page-budgets.
It complements the document \textit{Author's (Alternate) Guide to
Preparing ACM SIG Proceedings Using \LaTeX$2_\epsilon$\ and Bib\TeX}.
This source file has been written with the intention of being
compiled under \LaTeX$2_\epsilon$\ and BibTeX.

The developers have tried to include every imaginable sort
of ``bells and whistles", such as a subtitle, footnotes on
title, subtitle and authors, as well as in the text, and
every optional component (e.g. Acknowledgments, Additional
Authors, Appendices), not to mention examples of
equations, theorems, tables and figures.

To make best use of this sample document, run it through \LaTeX\
and BibTeX, and compare this source code with the printed
output produced by the dvi file. A compiled PDF version
is available on the web page to help you with the
`look and feel'.
\end{abstract}

% A category with the (minimum) three required fields
\category{H.4}{Information Systems Applications}{Miscellaneous}
%A category including the fourth, optional field follows...
\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

\terms{Theory}

\keywords{ACM proceedings, \LaTeX, text tagging}

\section{Introduction}

Sociam machines classification.

%ACKNOWLEDGMENTS are optional
\section{Acknowledgments}

SOCIAM ack here

%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%APPENDICES are optional

\balancecolumns % GM June 2007
% That's all folks!
\end{document}

### 2013-02-08 latex/sig-alternate.tex
Removed template ACM content from .tex file

@@
 % --- Author Metadata here ---
-\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
+\conferenceinfo{Social Machines Workshop}{2013, Paris, France}
 %\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
@@
 
-\title{Alternate {\ttlit ACM} SIG Proceedings Paper in LaTeX
-Format\titlenote{(Produces the permission block, and
-copyright information). For use with
-SIG-ALTERNATE.CLS. Supported by ACM.}}
-\subtitle{[Extended Abstract]
-\titlenote{A full version of this paper is available as
-\textit{Author's Guide to Preparing ACM SIG Proceedings Using
-\LaTeX$2_\epsilon$\ and BibTeX} at
-\texttt{www.acm.org/eaddress.htm}}}
+\title{A Framework for Classification of Social Machines}
+%\subtitle{[Extended Abstract]
+%\titlenote{A full version of this paper is available as
+%\textit{Author's Guide to Preparing ACM SIG Proceedings Using
+%\LaTeX$2_\epsilon$\ and BibTeX} at
+%\texttt{www.acm.org/eaddress.htm}}}
 %
@@
 
-\numberofauthors{8} %  in this sample file, there are a *total*
+\numberofauthors{1} %  in this sample file, there are a *total*
 % of EIGHT authors. SIX appear on the 'first-page' (for formatting
@@
 \alignauthor
-Ben Trovato\titlenote{Dr.~Trovato insisted his name be first.}\\
-       \affaddr{Institute for Clarity in Documentation}\\
-       \affaddr{1932 Wallamaloo Lane}\\
-       \affaddr{Wallamaloo, New Zealand}\\
-       \email{trovato@corporation.com}
-% 2nd. author
-\alignauthor
-G.K.M. Tobin\titlenote{The secretary disavows
-any knowledge of this author's actions.}\\
-       \affaddr{Institute for Clarity in Documentation}\\
-       \affaddr{P.O. Box 1212}\\
-       \affaddr{Dublin, Ohio 43017-6221}\\
-       \email{webmaster@marysville-ohio.com}
-% 3rd. author
-\alignauthor Lars Th{\o}rv{\"a}ld\titlenote{This author is the
-one who did all the really hard work.}\\
-       \affaddr{The Th{\o}rv{\"a}ld Group}\\
-       \affaddr{1 Th{\o}rv{\"a}ld Circle}\\
-       \affaddr{Hekla, Iceland}\\
-       \email{larst@affiliation.org}
-\and  % use '\and' if you need 'another row' of author names
-% 4th. author
-\alignauthor Lawrence P. Leipuner\\
-       \affaddr{Brookhaven Laboratories}\\
-       \affaddr{Brookhaven National Lab}\\
-       \affaddr{P.O. Box 5000}\\
-       \email{lleipuner@researchlabs.org}
-% 5th. author
-\alignauthor Sean Fogarty\\
-       \affaddr{NASA Ames Research Center}\\
-       \affaddr{Moffett Field}\\
-       \affaddr{California 94035}\\
-       \email{fogartys@amesres.org}
-% 6th. author
-\alignauthor Charles Palmer\\
-       \affaddr{Palmer Research Laboratories}\\
-       \affaddr{8600 Datapoint Drive}\\
-       \affaddr{San Antonio, Texas 78229}\\
-       \email{cpalmer@prl.com}
+Authors\\
+       \affaddr{Web and Internet Science Group}\\
+       \affaddr{University of Southampton}\\
+       \affaddr{Southampton, UK}\\
+       \email{\{a,b,c,d,nrs\}@ecs.soton.ac.uk}
 }
@@
 % in the \additional authors block, viz.
-\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
-email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
-(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
-\date{30 July 1999}
+%\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
+%email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
+%(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
+%\date{30 July 1999}
 % Just remember to make sure that the TOTAL number of authors
@@
 \begin{abstract}
-This paper provides a sample of a \LaTeX\ document which conforms,
-somewhat loosely, to the formatting guidelines for
-ACM SIG Proceedings. It is an {\em alternate} style which produces
-a {\em tighter-looking} paper and was designed in response to
-concerns expressed, by authors, over page-budgets.
-It complements the document \textit{Author's (Alternate) Guide to
-Preparing ACM SIG Proceedings Using \LaTeX$2_\epsilon$\ and Bib\TeX}.
-This source file has been written with the intention of being
-compiled under \LaTeX$2_\epsilon$\ and BibTeX.
 
-The developers have tried to include every imaginable sort
-of ``bells and whistles", such as a subtitle, footnotes on
-title, subtitle and authors, as well as in the text, and
-every optional component (e.g. Acknowledgments, Additional
-Authors, Appendices), not to mention examples of
-equations, theorems, tables and figures.
+This paper provides a framework for...
 
-To make best use of this sample document, run it through \LaTeX\
-and BibTeX, and compare this source code with the printed
-output produced by the dvi file. A compiled PDF version
-is available on the web page to help you with the
-`look and feel'.
 \end{abstract}
@@
 % A category with the (minimum) three required fields
-\category{H.4}{Information Systems Applications}{Miscellaneous}
+%\category{H.4}{Information Systems Applications}{Miscellaneous}
 %A category including the fourth, optional field follows...
-\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]
+%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]
 
-\terms{Theory}
+%\terms{Theory}
 
-\keywords{ACM proceedings, \LaTeX, text tagging}
+%\keywords{ACM proceedings, \LaTeX, text tagging}
 

### 2013-02-09 latex/sig-alternate.tex
added basic paper structure
including bullet points for sections 1 and 2. examples need to be
further discussed

@@
 
-Sociam machines classification.
+What are social machines and why are they important
+
+An emerging, interdisciplinary field of research and development with a growing number of social machine instances and studies about them
+
+We propose a classification framework for social machines. Targeted audiences
+
+* research community: shared terminology, set of key features and their possible instantiations based on the analysis of a relevant set of 50 (?) social machines (both popular systems as well as long tail systems); facilitates the study of design and evolution patterns; supports research on how communities emerge and develop and organizational studies (mechanism design, incentives); different research approaches can be better compared and aligned.
+
+* developer community: provides a theoretical grounding for the realization of new systems, informs system design and operation (what features should be implemented, what features are more important than others); enables predictions on user behavior and community development.
+
+The framework is based on interviews with XXX researchers covering XXX systems and was preliminary evaluated with respect to ...
+
+\section{Methodology}
+
+How was the classification framework defined
+
+* interviews with XXX researchers on key descriptive features of systems they are familiar with
+
+* consolidation of features listed for a resulting set of 50 (?) social machines
+
+* grid analysis to identify groups of similar systems
+
+* evaluation with respect to correctness, completeness, clarity, extendability in user survey, additional data set of XXX systems
+
+\section{Classification framework}
+
+First explain main principles: Distinction between application system and enabling technology; contributions, actions, activities, purpose, etc
+
+\section{Usage examples}
+Wikipedia
+
+GalaxyZoo (?)
+
+Ushahidi
+
+Facebook
+
+MechanicalTurk
+
+GWAPs
+
+\section{Related work}
+
+existing classification systems in CSCW, collective intelligence, social computing systems
+
+\section{Conclusions and future work}
+
+Detail evaluation plans
+
+Feedback from different disciplines, and from the developers community
 

### 2013-02-14 latex/sig-alternate.tex
added indicative length for each section
Added start of description, and added polyarchy figure

@@
 \documentclass{sig-alternate}
+
+\usepackage{url}
 
@@
        \affaddr{Southampton, UK}\\
-       \email{\{a,b,c,d,nrs\}@ecs.soton.ac.uk}
+       \email{\{a,b,c,d,wh,nrs\}@ecs.soton.ac.uk}
 }
@@
 
-\section{Introduction}
+\section{Introduction (0.5 page)}
 
@@
 
-\section{Methodology}
+\section{Methodology (1 page)}
 
@@
 
-\section{Classification framework}
+\section{Classification framework (2 pages)}
 
@@
 
-\section{Usage examples}
+\subsection{The Hierarchy/Polyarchical relationship of social machines}
+
+While determining a set of social machines, one clear distinction was between social machine frameworks, such as MediaWiki and Ushahidi, which enable
+social machines to be created, and instantiations of those frameworks into social machines that have been created for a specific purpose, such as Wikipedia
+and Ushahidi Haiti. However, through further investigation, we discovered that the distinction between framework and instantiations was not the only
+differentiator. Specifically, that some social machines could also have sub-machines or communities within them, that are working to solve different problems,
+within the context of the larger machine. For example, Wikipedia is an instantiation of the MediaWiki framework, but there are a large number of
+communities within Wikipedia that work on domain-specific areas, such as those contributing in areas of science and technology\footnote{Visualisation of Wikipedia Science Communities: \url{http://www.olihb.com/WikiCommunities/}}.
+However, we must also note that wikipedia is primarily split into communities of language, with the English, German and French wikipedias having the largest number of articles.
+It is therefore possible to analyse wikipedia into a hierarchy, from MediaWiki, to the languages of Wikipedia, down to the domain-specific communities.
+We explored this idea further, to include other social machines, and links where social machines utilise other social machines in order to function.
+For example, the social machine used by the Obama '12 US Presidential campaign~\cite{obamakieron}, which relied on social machines such as Twitter and Facebook.
+A resulting polyarchy is shown in Figure \ref{polyarchy}, which classifies levels as Infrastructure, Frameworks, Services and Causes/Groups.
+
+\begin{figure*}[htb]
+\begin{center}
+\includegraphics[width=18cm]{img/polyarchy.pdf}
+\caption{A polyarchy of Social Machines, illustrating the infrastructure and frameworks used by social machines, and machine-machine usage.} \label{polyarchy}
+\end{center}
+\end{figure*}
+
+\subsection{Artefacts of Social Machines}
+
+Our next observation was that there are a number of shared artefact types that can be classified within the context of social machines. Specifically that
+there are ``intentions'' that motivate the use of social machines, ``actions'' that are directly performed by participants of social machines, and ``contributions``
+that result from the use of social machines. We have taken these abstract concepts, and mapped them to a number of well-known social machines in order
+to demonstrate that these concepts can be used to compare and contrast different machines.
+
+\section{Usage examples (0.5 page)}
 Wikipedia
@@
 
-\section{Related work}
+\section{Related work (0.5 pages)}
 
@@
 
-\section{Conclusions and future work}
+\section{Conclusions and future work (0.5 pages)}
 
@@
 
-SOCIAM ack here
+This work is supported under SOCIAM: The Theory and Practice of Social Machines.
+The SOCIAM Project is funded by the UK Engineering and Physical Sciences Research Council (EPSRC) under grant number EP/J017728/1 and comprises the Universities of Southampton, Oxford and Edinburgh.
 

### 2013-02-14 latex/sig-alternate.tex
Added more about framework, and added motivation figure

@@
 For example, the social machine used by the Obama '12 US Presidential campaign~\cite{obamakieron}, which relied on social machines such as Twitter and Facebook.
+
+Looking at the set of social machines in a polyarchy leads to a broad/specific relationship emerging that lets us talk about behaviours at various levels of granularity.
+We propose looking at nested machines with ``The Web'' as one of several potential roots, with the next level down consisting of sub-platforms (such as Facebook, Twitter, MediaWiki, Ushahid platforms) that spawn more specific social machines.
 A resulting polyarchy is shown in Figure \ref{polyarchy}, which classifies levels as Infrastructure, Frameworks, Services and Causes/Groups.
@@
 
+This approach enables us to start with a more detailed analysis of certain levels over others;
+and seeing what similarities flow up and down the polyarchy. For example, what do
+specific instances of Ushahidi/Zooniverse/MediaWiki have in common with other instances, and
+how to they differ dynamically?
+
+\subsection{Multi-faceted motivation, and who-gets-what benefit}
+
+We identified a hierarchy of constructs pertaining to the motivation people have for using social
+machines with top level constructs of:
+
+\begin{enumerate}
+\item Hedonistic
+\item Financial
+\item To be informed
+\item To help others
+\end{enumerate}
+
+The sub constructs are illustrated in Figure \ref{motivation}, and we relate them to Maslow's hierarchy of needs~\cite{maslow}, (e.g., self-actualisation, security, respect of others), in various ways.
+
+\begin{figure*}[htb]
+\begin{center}
+\includegraphics[width=18cm]{img/motivation.pdf}
+\caption{Motivation hierarchy of participants using Social Machines, related to Maslow's hierarchy of needs.} \label{motivation}
+\end{center}
+\end{figure*}
+
+For each of these motivations, different roles that participate may have different {\it core motivations}, such as:
+
+\begin{enumerate}
+\item Benefit to contributor
+\item Benefit to moderator
+\item Benefit to system operator/host (e.g, Google, Facebook, Amazon)
+\item To affiliates of the host (e.g., Amazon affiliates, YouTube partners)
+\item To society as a whole
+\item To a contributor's social network
+\end{enumerate}
+
 \subsection{Artefacts of Social Machines}
@@
 to demonstrate that these concepts can be used to compare and contrast different machines.
+
+[more here]
+
+\subsection{Metrics of participants and usage}
+
+Finally, we identified several contracts that are relevant for all social machines.
+First, there are construct that relate to ``analytics'' that can be measured at a point in time, and may change over time.
+Specifically:
+
+\begin{enumerate}
+\item Size
+\item Maturity
+\item Competitors
+\item Diversity of user base
+\item Popularity over time
+\item Frequency of participation
+\end{enumerate}
+
+Next, there are constructs regarding the design affordances of the social machine:
+
+\begin{enumerate}
+\item Does it afford anonymity
+\item Does it afford quality control or moderation, and in what ways
+\item What types of contribution does it afford
+\item Openness of resulting data set
+\end{enumerate}
+
+In the middle we have a mixture of the result of how the design affordances actually were realised by the social computation, such as how users co-opted / incorporated / interpreted the constraints and:
+
+\begin{enumerate}
+\item worked within them
+\item worked around them
+\item gamed them
+\item extended and enhanced them
+\item derived new structures and conventions (e.g., hashtags)
+\end{enumerate}
+
+The result is what we call `participation' constructs including:
+
+\begin{enumerate}
+\item Roles emerging in the social machine  (domain expert, troll/spammer, hub, personal gainer)
+\item Quality of contributions
+\item Content of contributions
+\item Global / cultural reach
+\item Task reach --- what do people use it for? 	
+\item Active/Passive participation roles 
+\end{enumerate}
 

### 2013-02-15 latex/sig-alternate.tex
Added more about framework, put affordances in a table

@@
 
+In addition to relationships to Maslow's hierarchy of needs, which drive the motivations,
+we suggest that there are also resulting actions from each motivation. Thus, it should be
+possible to follow basic human needs to motivations in social machines, through to the
+actual actions (and therefore the necessary software feature support) that are made by
+participants in social machines. Our theory is that this path may enable designers of
+social machines to satisfy human needs by implemented specific features into their social
+machines, or to identify potential gaps in their serving of needs and motivations due
+to lack of implemented features.
+
 \subsection{Artefacts of Social Machines}
@@
 
-Finally, we identified several contracts that are relevant for all social machines.
+Finally, we identified several constructs that are relevant for all social machines.
 First, there are construct that relate to ``analytics'' that can be measured at a point in time, and may change over time.
-Specifically:
+We have documented these constructs in Table \ref{table:constructs}.
+In the left column we list construct regarding the design affordances of the social machine.
+In the middle column we have a mixture of the result of how the design affordances actually were realised by the social computation, such as how users co-opted / incorporated / interpreted the constraints.
+Finally, in the right column we list analytical measures of social machines which are subject to change over time.
+
+\begin{table*}[htbp]
+\begin{center}
+\begin{tabular}{|p{6.5cm}|p{5.5cm}|p{4.5cm}|}
+\hline
+Designed Affordances & Participatory Affordances & Analytics \\
+\hline
+Does it afford anonymity & worked within them & Size \\
+Does it afford quality control or moderation, and in what ways & worked around them & Maturity \\
+What types of contribution does it afford & gamed them & Competitors \\
+Openness of resulting data set & extended and enhanced them & Diversity of user base \\
+ & derived new structures and conventions (e.g., hashtags) & Population over time \\
+ &  & Frequency of participation \\
+\hline
+\end{tabular}
+\end{center}
+\caption{Constructs of social machines, listing affordances that are designed into machines, affordances that have resulted from use, and analytical constructs that can be measured and change over time.} \label{table:constructs}
+\end{table*}%
+
+The result is what we call `participation' constructs. These constructs arise from use of a
+social machine, and are not necessarily predictable; they include:
 
 \begin{enumerate}
-\item Size
-\item Maturity
-\item Competitors
-\item Diversity of user base
-\item Popularity over time
-\item Frequency of participation
-\end{enumerate}
-
-Next, there are constructs regarding the design affordances of the social machine:
-
-\begin{enumerate}
-\item Does it afford anonymity
-\item Does it afford quality control or moderation, and in what ways
-\item What types of contribution does it afford
-\item Openness of resulting data set
-\end{enumerate}
-
-In the middle we have a mixture of the result of how the design affordances actually were realised by the social computation, such as how users co-opted / incorporated / interpreted the constraints and:
-
-\begin{enumerate}
-\item worked within them
-\item worked around them
-\item gamed them
-\item extended and enhanced them
-\item derived new structures and conventions (e.g., hashtags)
-\end{enumerate}
-
-The result is what we call `participation' constructs including:
-
-\begin{enumerate}
-\item Roles emerging in the social machine  (domain expert, troll/spammer, hub, personal gainer)
+\item Roles emerging in the social machine.
+   \indent In addition to designed and implemented roles such as {\it moderator} and {\it administrator}, emerging roles include motivations of the participant, such as {\it trolls}, {\it spammers}, {\it domain experts}, {\it social nexus}, and {\it re-blogger}.
 \item Quality of contributions
+    \indent For any social machine, the quality of contributions will vary. While some
+    issues with quality can be predicted and protected against (e.g., through 
+    multi-user result agreement on Mechanical Turk~\cite{ipeirotis2010quality}), some
+    particular issues with quality cannot always be predicted ahead of time.
 \item Content of contributions
+    \indent While a system operator may design their social machine to generate specific
+    types of content, participants may subvert the system in order to generate other types,
+    or to concentrate on areas that were unforeseen at the time of launch, but are now seen
+    as useful to the operators.
 \item Global / cultural reach
+    \indent 
 \item Task reach --- what do people use it for?

### 2013-02-15 latex/sig-alternate.tex
Added "Open Access" as a SM, and updated figure to show it

@@
 We propose looking at nested machines with ``The Web'' as one of several potential roots, with the next level down consisting of sub-platforms (such as Facebook, Twitter, MediaWiki, Ushahid platforms) that spawn more specific social machines.
-A resulting polyarchy is shown in Figure \ref{polyarchy}, which classifies levels as Infrastructure, Frameworks, Services and Causes/Groups.
+A resulting polyarchy is shown in Figure~\ref{polyarchy}, which classifies levels as Infrastructure, Frameworks, Services and Causes/Groups.
 
@@
 
-The sub constructs are illustrated in Figure \ref{motivation}, and we relate them to Maslow's hierarchy of needs~\cite{maslow}, (e.g., self-actualisation, security, respect of others), in various ways.
+The sub constructs are illustrated in Figure~\ref{motivation}, and we relate them to Maslow's hierarchy of needs~\cite{maslow}, (e.g., self-actualisation, security, respect of others), in various ways.
 
@@
 First, there are construct that relate to ``analytics'' that can be measured at a point in time, and may change over time.
-We have documented these constructs in Table \ref{table:constructs}.
+We have documented these constructs in Table~\ref{table:constructs}.
 In the left column we list construct regarding the design affordances of the social machine.
@@
 
+\subsection{Causes as Social Machines}
+
+One originally unforeseen classification is that of the causes that utilise Social Machines.
+Through a particularly popular cause, large numbers of participants can be mobilised, for
+either a short period of time, or for long-term engagement, using multiple social machines
+in order to further their cause. Examples of causes can be seen in Figure~\ref{polyarchy},
+at the lowest level of the polyarchy. One such example is the cause of ``Open Access'' to
+academic publications~\cite{harnad2001self}. This cause mobilises academic authors to
+self-archive their own papers online so they are available for free, in addition to hosted
+versions behind publishers pay-walls. In order to achieve this aim, the social machine of
+``open access'' typically uses the social machines of institutional repositories provided
+by the authors' associated institutions. As with other social machines, these social
+machines typically use off-the-shelf implementations (in this case the most popular are
+EPrints~\cite{eprints} and DSpace~\cite{dspace}), which run as web sites on the world wide
+web. This cause is mature, growing year-on-year and has global coverage.
+
+[more about this?]
+
 \section{Related work (0.5 pages)}

### 2013-02-15 latex/sig-alternate.tex
Added particapatory constructs

@@
 \begin{enumerate}
-\item Roles emerging in the social machine.
-   \indent In addition to designed and implemented roles such as {\it moderator} and {\it administrator}, emerging roles include motivations of the participant, such as {\it trolls}, {\it spammers}, {\it domain experts}, {\it social nexus}, and {\it re-blogger}.
-\item Quality of contributions
-    \indent For any social machine, the quality of contributions will vary. While some
+\item {\bf Roles emerging in the social machine}
+   \newline In addition to designed and implemented roles such as {\it moderator} and {\it administrator}, emerging roles include motivations of the participant, such as {\it trolls}, {\it spammers}, {\it domain experts}, {\it social nexus}, and {\it re-blogger}.
+\item {\bf Quality of contributions}
+    \newline For any social machine, the quality of contributions will vary. While some
     issues with quality can be predicted and protected against (e.g., through 
@@
     particular issues with quality cannot always be predicted ahead of time.
-\item Content of contributions
-    \indent While a system operator may design their social machine to generate specific
+\item {\bf Content of contributions}
+    \newline While a system operator may design their social machine to generate specific
     types of content, participants may subvert the system in order to generate other types,
     or to concentrate on areas that were unforeseen at the time of launch, but are now seen
-    as useful to the operators.
-\item Global / cultural reach
-    \indent 
-\item Task reach --- what do people use it for?
-\item Active/Passive participation roles 
+    as useful to the operators. It is also possible that multiple social machines can be
+    combined/linked by users to further expand the types of contributions they produce.
+\item {\bf Global / cultural reach}
+    \newline A social machine may naturally focus on a narrow geographic area, for example
+    an Ushahidi-based social machine for local elections~\cite{meier2008crisis} or
+    natural disasters~\cite{morrow2011independent}. Such machines are unlikely to be
+    re-purposed by participants outside of that area. However, some machines have been
+    created for a single cultural/global area, but have potential global relevance that was
+    unforeseen early on. An example is the fall-off of English-language users of the
+    social network ``Orkut,'' which gained significant usage elsewhere, particularly in
+    Brazil.
+\item {\bf Task reach --- what do people use it for?}
+    \newline In addition to breaking cultural barriers, user have also shown that they can
+    use social machines for tasks that were not set up by the system operators. For example
+    the identification of ``Green Pea Galaxies'' using the GalaxyZoo social machine for
+    astronomy~\cite{greenpea}. Users were not originally asked to identify these types, but
+    a number of users identified them, and the software was modified to include them,
+    resulting in their discovery as a new type of galaxy.
+\item {\bf Active/Passive participation roles}
+    \newline When it comes to participation in social machines, there are levels of
+    participation that can be classified as being ``active'' and being ``passive.'' These
+    roles may not be particularly designed into systems, but may be exhibited over time.
+    Particularly noteworthy examples can be seen in news and link aggregation social
+    machines such as Reddit and Digg, where users can both submit links and vote on links.
+    The frequency of each of these activities can vary wildly, with a small
+    minority of users controlling the links that rise to the top~\cite{digg}.
 \end{enumerate}

### 2013-02-15 latex/sig-alternate.tex
Methodology added

@@
 
-* research community: shared terminology, set of key features and their possible instantiations based on the analysis of a relevant set of 50 (?) social machines (both popular systems as well as long tail systems); facilitates the study of design and evolution patterns; supports research on how communities emerge and develop and organizational studies (mechanism design, incentives); different research approaches can be better compared and aligned.
+* research community: shared terminology, set of key features and their possible instantiations based on the analysis of a relevant set of 50 (?) social machines (both popular systems as well as long tail systems); facilitates the study of design and evolution patterns; supports research on how communities emerge and develop and organisational studies (mechanism design, incentives); different research approaches can be better compared and aligned.
 
-* developer community: provides a theoretical grounding for the realization of new systems, informs system design and operation (what features should be implemented, what features are more important than others); enables predictions on user behavior and community development.
+* developer community: provides a theoretical grounding for the realisation of new systems, informs system design and operation (what features should be implemented, what features are more important than others); enables predictions on user behaviour and community development.
 
@@
 
-How was the classification framework defined
+%How was the classification framework defined
+%
+%* interviews with XXX researchers on key descriptive features of systems they are familiar with
+%
+%* consolidation of features listed for a resulting set of 50 (?) social machines
+%
+%* grid analysis to identify groups of similar systems
+%
+%* evaluation with respect to correctness, completeness, clarity, extendability in user survey, additional data set of XXX systems
+%
 
-* interviews with XXX researchers on key descriptive features of systems they are familiar with
+In order to identify and classify social machines, we first embarked on a principled
+knowledge elicitation process~\cite{knowledgeelicitation}, to determine the {\it elements}
+(i.e., the social machines) and the {\it constructs} (i.e., the classifying factors) that
+we can use to classify social machines. We used the {\it repertory grid} elicitation
+technique~\cite{kelly} in order to derive an initial set of elements and constructs. We
+asked ten Computer Science researchers familiar with social machines to create their own
+repertory grids, and populate the elements from their own knowledge, and to create the
+constructs using the standard repertory grid elicitation technique. This exercise led
+to 10 grids, the union of which comprised a total of 56 unique elements (social machines)
+and 117 different constructs.
 
-* consolidation of features listed for a resulting set of 50 (?) social machines
+To make the set of social machines easier to refer to and more approachable, we performed
+an initial grid analysis of the machines, clustering them into related user-centered
+themes, of: Social Networks, (Micro)Blogging, News Aggregators, Image Boards, Crowd Science,
+Answer Gardens, Community Watch, Health and Wellbeing Support, Action and Investigation,
+Opinion Sharing, Video Sharing, Photo Sharing, Code Sharing, Art Sharing, Crowdsourcing
+Platforms, Mash-up Systems, Crowdsourcing Toolkits and Platforms.
 
-* grid analysis to identify groups of similar systems
+While determining the intersection of elements was straightforward, the consolidation of the
+constructs required a more thorough process. To consolidate the constructs we embarked on a manual process of grouping the constructs
+into rough clusters, based around the areas they cover. For example, we used construct
+clusters including: Users, Motivation, Popularity, Technology, and Purpose. We then examined
+each construct to determine which were equivalent, and whether we could re-word or subsume
+existing constructs to cover the same aspects as a construct, with the aim to remove any
+redundancy, to cut the list of constructs down to a manageable number, but to ensure that
+all of the constructs that were elicited were somehow represented in the final set. This
+process involved four of the authors discussing the choices, and finally agreeing on a
+consolidated set of 31 constructs.
 
-* evaluation with respect to correctness, completeness, clarity, extendability in user survey, additional data set of XXX systems
+Even with this reduced set of constructs, if a single person were to classify all 56
+elements, this would still require 1736 individual decisions, and hence is unfeasible to
+expect full coverage of the elements using all of these constructs, despite how representative
+they are.
+
+[something here about doing 2 grids each and verifying the results?]
+
+To extract greater usefulness, and the ability to use our constructs to classify social
+machines we re-evaluated the constructs and the semantics that they represent into richer
+constructs that are not constrained by the 1--5 rating of repertory grid constructs.
 
@@
 
-First explain main principles: Distinction between application system and enabling technology; contributions, actions, activities, purpose, etc
+%First explain main principles: Distinction between application system and enabling technology; contributions, actions, activities, purpose, etc
 
-\subsection{The Hierarchy/Polyarchical relationship of social machines}
+To create a classification framework, it is necessary to first determine the aspects of
+classification. We will first discuss the relationships between social
+machines, enabling technologies and platforms; we move on to discuss the multi-faceted
+nature of motivation within social machines; next we evaluate the artefacts of social
+machines: contributions, actions, activities and purpose; finally we discuss the metrics of
+participation, and the difference between designed and emergent properties of social
+machines.
+
+\subsection{The Hierarchical/Polyarchical relationship of social machines}
 
@@
 Looking at the set of social machines in a polyarchy leads to a broad/specific relationship emerging that lets us talk about behaviours at various levels of granularity.
-We propose looking at nested machines with ``The Web'' as one of several potential roots, with the next level down consisting of sub-platforms (such as Facebook, Twitter, MediaWiki, Ushahid platforms) that spawn more specific social machines.
+We propose looking at nested machines with ``The Web'' as one of several potential roots, with the next level down consisting of sub-platforms (such as Facebook, Twitter, MediaWiki, Ushahidi platforms) that spawn more specific social machines.
 A resulting polyarchy is shown in Figure~\ref{polyarchy}, which classifies levels as Infrastructure, Frameworks, Services and Causes/Groups.
@@
 \section{Usage examples (0.5 page)}
+
 Wikipedia

### 2013-02-15 latex/sig-alternate.tex
Added all consolidated constructs to the table (needs tidying)

@@
 
-\begin{table*}[htbp]
+\begin{table*}[htb]
 \begin{center}
-\begin{tabular}{|p{6.5cm}|p{5.5cm}|p{4.5cm}|}
+\begin{tabular}{|p{18cm}|}
 \hline
-Designed Affordances & Participatory Affordances & Analytics \\
+{\bf Designed Affordances} \\
 \hline
-Does it afford anonymity & worked within them & Size \\
-Does it afford quality control or moderation, and in what ways & worked around them & Maturity \\
-What types of contribution does it afford & gamed them & Competitors \\
-Openness of resulting data set & extended and enhanced them & Diversity of user base \\
- & derived new structures and conventions (e.g., hashtags) & Population over time \\
- &  & Frequency of participation \\
+Domain specificity [1 - Specific / 5 - General] vs \\
+\hspace{0.66cm} Generality of Purpose \\
+Visibility and importance of a participants `reputation' [1 - Not important / 5- Important] \\
+\hspace{0.66cm} Explicit representation of participant reputation [1 - No Reputations / 5 - Reputations] \\
+Participants can define new types of contributions [1 - Not possible / 5 - Complete freedom] \\
+Degree of open source software used/created [1 - Proprietary / 5 - Entirely open source] \\
+Degree of openness and availability of data created in the system (APIs/Dumps) [1 - Closed/unavailable data / 5 - Completely open data] \\
+Variety of actions participants can perform [1 - Single Action / 5 - Large variety of actions] \\
+\hline
+{\bf Participatory Affordances} \\
+\hline
+(Level of social features) - Supports social interaction [1 - No social features / 5 - Large variety of social features] \\
+How often a participant participates (times a day, etc) [1 - Many times a day / 5 Seldom] \\
+Importance of Timely Participation [1 - Unimportant / 5 - Timely] \\
+Generality of Audience [1 - Niche / 5 - General] \\
+Participant anonymity [1 - No Anonymity / 5 - Complete Anonymity] \\
+Participant autonomy (1 - No autonomy / 5 - Complete autonomy) \\
+Quality of participant contributions [1 - Low quality / 5 - High quality] \\
+(Different roles clearly defined) - \\
+\hspace{0.66cm} Clear separation of roles/responsibilities among participants - [1 - Single role - Everyone is the same) / 5 - People assume different roles/responsibilities] \\
+	Extent of hierarchical organisation of roles [1 - No Hierarchy / 5 - Deep/Well-defined Hierarchy] \\
+Variety of types of contributions [1 - Single Type / 5 - Large variety of types] \\
+Participants are motivated by extrinsic award  [1 - Disagree / 5 - Agree] \\
+Participants are intrinsically motivated [1 - Disagree / 5 - Agree] \\
+Service/Platform derives benefit from participant participation [1 - Disagree /  5- Agree] \\
+Participants directly benefit from participating [1 - Disagree / 5 - Agree] \\
+Participants' friends/social network benefit from participation [1 - Disagree / 5 - Agree] \\
+Participation is for the benefit of a specific person/group other than the participant or platform owner [1-Disagree / 5-Agree] \\
+Participation is for the benefit of society/the world at large  [1 - Disagree / 5 - Agree] \\
+Participation is done to `get something done' [1-Disagree/5-Agree] \\
+Participation is done `for fun' [1-Disagree/5-Agree] \\
+Participation is done to exchange knowledge [1-Disagree/5-Agree] \\
+Participation is done to be social [1-Disagree/5-Agree] \\
+Participation is done via mobile devices [1-Never / 5 - Often] \\
+Participants' (geographic) location is used by the service [1-No / 5 - Very] \\
+\hline
+{\bf Analytics} \\
+\hline
+Popularity [1 - Unpopular / 5 - Very Popular] \\
+Maturity [1 - New / 5 - Mature] \\
+Ratio of passive to active participants [1 - Passive / 5 - Active] \\
+Number of competitors [1 - Not Many / 5 - Many] \\
+Extent of global geographic coverage [1 - Highly localised / 5 - Global]\\
 \hline

### 2013-02-20 latex/paper_new.tex
New structure of the paper
Following discussions with Wendy

--- 2013-02-15 latex/sig-alternate.tex
@@
 % --- Author Metadata here ---
-\conferenceinfo{Social Machines Workshop}{2013, Paris, France}
+\conferenceinfo{Workshop on the theory and practice of social machines @ WWW2013}{2013, Rio de Janeiro, Brazil}
 %\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
@@
 
-\title{A Framework for Classification of Social Machines}
+\title{Towards a classification framework for social machines}
 %\subtitle{[Extended Abstract]
@@
 
-\section{Introduction (0.5 page)}
+\section{Social machines: an introduction}
+Once upon a time 'machines' were programmed by programmers and used by users. The Web, and the
+massive participation in it, has changed this relationship: we now see configurations of people interacting with content exposed via the Web and with each other, typified by social Web sites.
+Rather than drawing a line through such Web-based systems to separate the human and machine parts (as Computer Science has traditionally done), we can now draw a line around them  and  treat  each  such  assembly  as  a  'social  machine'  ñ  a  machine  comprising  digital  and  human
+components and the interactions between them.
+This crucial transition in thinking acknowledges the reality of today's computing systems and is essential to underpin any understanding of the science and engineering of our future socio-technical systems: we view an ecosystem of co-evolving social machines.
 
-What are social machines and why are they important
+The  typical  social Web  site  is  a  social machine, with  its user-generated  content  and  social network evolving dynamically  and  hand-in-hand.
+More  than  just  'data' or  'content',  these sites have  some notion of  'social artefacts' around which we forge the links in the social network, like photos on flickr or movies on YouTube, and they  can also be  flourishing  spaces  for human  ingenuity.
+Today we  see  a  divide  between  conventional  computer  systems  dedicated  to  data-  and  computation-intensive tasks, and social computing like the Web 2.0 sites.
+But many of our challenges, from healthcare to transport, need  both  of  these  axes:  a  sophisticated  combination  of  data-intensive  automation  and  deep  social involvement.
+This suggests the need for new social machines to tackle new problems, and to achieve this we
+need to understand the science and engineering of social machines.
 
-An emerging, interdisciplinary field of research and development with a growing number of social machine instances and studies about them
+A  traditional  database,  with  users  looking  up  records  independently  of  each  other  and  an  administrator responsible for content, has some of the right ingredients but there is really no social component in this strict provider-consumer relationship.
+When we fill out a form (e.g.
+health information, birds spotted in my garden) it  is minimally  a  social machine  because humans  are  part  of  the  'social  computation'.
+This  becomes  richer when  the database  is  curated by members  of  the  community  (e.g.
+Wikipedia) and when  the  social network adds value implicitly (amazon) or explicitly (Facebook).
+We anticipate many categories of social machine in this continuum, and they may manifest themselves as Web site or app, or indeed in the Internet of Things.
 
-We propose a classification framework for social machines. Targeted audiences
+An emerging, interdisciplinary field of research and development with a growing number of social machine instances and studies about them.
+
+We propose a classification framework for social machines.
+
+Targeted audiences
 
@@
 
-\section{Methodology (1 page)}
+Dimensions of the classification
 
-%How was the classification framework defined
-%
-%* interviews with XXX researchers on key descriptive features of systems they are familiar with
-%
-%* consolidation of features listed for a resulting set of 50 (?) social machines
-%
-%* grid analysis to identify groups of similar systems
-%
-%* evaluation with respect to correctness, completeness, clarity, extendability in user survey, additional data set of XXX systems
-%
+* tasks and types of contributions: what does the system do, how do participants contribute: from supplying data and creating content to solving problems or validating solutions.
 
+* social network and interaction: roles, hierarchies of roles, profiling
+
+* motivation and incentives: system features influencing user behavior and the social network dynamics
+
+* human-machine interaction: how are human contributions used to compute things, including quality control, spam, workflow, aggregation of results
+
+People  supply  or  refine  data.    Systems  in  this  category  use  people  as,  effectively,  sophisticated  sensors.
+Examples of  this  sort of  system  are Galaxy  Zoo  [www.galaxyzoo.org] where people  classify  galaxies with
+collective performance as good as professional astronomers and reCaptcha [www.google.com/recaptcha]
+where people refine data describing digital manuscripts.
+Åú People are elementary problem solvers. Systems in this category typically find innovative ways of getting
+people to perform a task that are easy for them but for which no effective computer algorithm is available.
+Examples  of  this  sort  of  system  are Mechanical  Turk  [www.mturk.com]  and  the  related  Turkit  system
+[groups.csail.mit.edu/uid/turkit]  that  allows  these  actions  to  be  strung  together  iteratively,  thus  giving  a
+basic form of composition.
+Åú People generate/test partial  solutions. Systems  in  this  category provide  incentives  for people  to perform
+complex  problem  solving  that  generates  a  design  to  be  tested  automatically  and/or  tests  automatically
+generated  designs.    Examples  include  EteRNA  [eterna.cmu.edu]  in which  participants  design RNA  that  is
+evaluated automatically using simulators and then refined via human effort; Matlab programming contests
+[www.mathworks.co.uk/matlabcentral/contest] where people compete for the best solution as judged by automated  tests  that  compute  the  efficiency  of  the  code;  or  Scratch  [scratch.mit.edu]  in  which  simple programs are created by some participants and modified by others in a remixing process.
+
+\section{Comparison with related fields}
+
+Web 2.0
+
+Collective Intelligence, Wisdom of the crowds
+
+Open Innovation
+
+In  social machines  the  social and  the  technical are equal  partners. While  crowdsourcing  could  be  viewed  as  'human  subroutines'  and  'citizen  sensors',  the
+social machine perspective is one of incentivised participation and not humans enslaved to computers.
+
+Crowdsourcing, human computation
+
+CSCW, social computing
+
+\section{Classification of social machines}
+
+\subsection{Methodology}
 In order to identify and classify social machines, we first embarked on a principled
@@
 
-\section{Classification framework (2 pages)}
+\subsection{Classification dimensions}
 
-%First explain main principles: Distinction between application system and enabling technology; contributions, actions, activities, purpose, etc
-
-To create a classification framework, it is necessary to first determine the aspects of
-classification. We will first discuss the relationships between social
-machines, enabling technologies and platforms; we move on to discuss the multi-faceted
-nature of motivation within social machines; next we evaluate the artefacts of social
-machines: contributions, actions, activities and purpose; finally we discuss the metrics of
-participation, and the difference between designed and emergent properties of social
-machines.
-
-\subsection{The Hierarchical/Polyarchical relationship of social machines}
-
+\subsubsection{The polyarchical relationship of social machines}
 While determining a set of social machines, one clear distinction was between social machine frameworks, such as MediaWiki and Ushahidi, which enable
@@
 
-\subsection{Multi-faceted motivation, and who-gets-what benefit}
+One originally unforeseen classification is that of the causes that utilise Social Machines.
+Through a particularly popular cause, large numbers of participants can be mobilised, for
+either a short period of time, or for long-term engagement, using multiple social machines
+in order to further their cause. Examples of causes can be seen in Figure~\ref{polyarchy},
+at the lowest level of the polyarchy. One such example is the cause of ``Open Access'' to
+academic publications~\cite{harnad2001self}. This cause mobilises academic authors to
+self-archive their own papers online so they are available for free, in addition to hosted
+versions behind publishers pay-walls. In order to achieve this aim, the social machine of
+``open access'' typically uses the social machines of institutional repositories provided
+by the authors' associated institutions. As with other social machines, these social
+machines typically use off-the-shelf implementations (in this case the most popular are
+EPrints~\cite{eprints} and DSpace~\cite{dspace}), which run as web sites on the world wide
+web. This cause is mature, growing year-on-year and has global coverage.
+
+\subsubsection{Multi-faceted motivation, and who-gets-what benefit}
 
@@
 
-\subsection{Artefacts of Social Machines}
+\subsubsection{Artefacts of Social Machines}
 
@@
 
-\subsection{Metrics of participants and usage}
+\subsubsection{Metrics of participants and usage}
 
@@
 
-\begin{table*}[htb]
+\begin{table*}[htbp]
 \begin{center}
@@
 
-\section{Usage examples (0.5 page)}
+\subsection{Using the classification framework}
 
-Wikipedia
+In order to demonstrate usage of our framework's constructs, we ranked our social machines
+by their Alexa ranking, and performed a full repertory grid elicitation exercise. The
+results are illustrated in a grid and accompanying dendrograms for element and construct
+similarity in Figure~\ref{dendrogram}.
 
-GalaxyZoo (?)
+\begin{figure*}
+\begin{center}
+\includegraphics[width=18cm]{img/dendrogram.png}
+\caption{Dendrogram from a repertory grid exercise of the top 20 (ranked by Alexa) social machines from our element set, against our consolidated constructs.} \label{dendrogram}
+\end{center}
+\end{figure*}
 
-Ushahidi
+[needs some discussion on what this shows.]
 
-Facebook
+\section{Future work: building a social machine around the classification framework}
 
-MechanicalTurk
-
-GWAPs
-
-\subsection{Causes as Social Machines}
-
-One originally unforeseen classification is that of the causes that utilise Social Machines.
-Through a particularly popular cause, large numbers of participants can be mobilised, for
-either a short period of time, or for long-term engagement, using multiple social machines
-in order to further their cause. Examples of causes can be seen in Figure~\ref{polyarchy},
-at the lowest level of the polyarchy. One such example is the cause of ``Open Access'' to
-academic publications~\cite{harnad2001self}. This cause mobilises academic authors to
-self-archive their own papers online so they are available for free, in addition to hosted
-versions behind publishers pay-walls. In order to achieve this aim, the social machine of
-``open access'' typically uses the social machines of institutional repositories provided
-by the authors' associated institutions. As with other social machines, these social
-machines typically use off-the-shelf implementations (in this case the most popular are
-EPrints~\cite{eprints} and DSpace~\cite{dspace}), which run as web sites on the world wide
-web. This cause is mature, growing year-on-year and has global coverage.
-
-[more about this?]
-
-\section{Related work (0.5 pages)}
-
-existing classification systems in CSCW, collective intelligence, social computing systems
-
-\section{Conclusions and future work (0.5 pages)}
-
-Detail evaluation plans
-
-Feedback from different disciplines, and from the developers community
-
-%ACKNOWLEDGMENTS are optional
 \section{Acknowledgments}

### 2013-02-21 latex/paper_new.tex
related work section

@@
 \section{Comparison with related fields}
+
+A Social Machine includes human and machine two major components, therefore, the study of social machine is an interdisciplinary endeavour.
+Since the Turning machine was invented, there has been a `paradigm shift' in using computers, where the computer facilitates human communication rather acts as a purely calculative device.
+The Computer-supported Collaboration Work (CSCW) emerged in 1982 represent such shift.
+It addresses ``how collaborative activities and their coordination can be supported by means of computer systems'' later evolved in a more broader sense of  Computer-supported collaboration (CSC).
+The Social computing is another shift, which is a generic term refers systems that support ``the gathering, representation, processing, use, and dissemination of information that is distributed across social collectivities such as teams, communities, organizations, and markets''.
+To compare with the CSC, the social computing has more focus in computing rather than broader sense of
+
+CSCW (1982)	"how collaborative activities and their coordination can be supported by means of computer systems."
+
+Social Computing (1994)	the system that support the gathering, representation, processing, use, and dissemination of information that is distributed across social collectivities such as teams, communities, organizations, and markets.
+
+Open Innovation (2003)	Paradigm in how companies commercialize trial knowledge.
+
+Is a paradigm that assumes that firms can and should use external ideas as well as internal ideas, and internal and external path to market, as the firms look to advance their technology.
+It combines internal and external ideas into architectures and systems whose requirements are defined by a business model.
+The Wisdom of Crowds (2004)
+	About the aggregation of information in groups, resulting in decisions that, he argues, are often better than could have been made by any single member of the group. 
+Human Computation (2005)	.	“...a paradigm for utilizing human processing power to solve problems that computers cannot yet solve.”
+
+Crowdsourcing (2006)	“Crowdsourcing is the act of taking a job traditionally
+performed by a designated agent (usually an employee) and outsourcing it to an undefined, generally large group of people in the form of an open call.”
+.	
+Collective Intelligence (2009)	"... groups of individuals doing things collectively that seem intelligent.”
+
+The Wisdom of Crowds described by Surowiecki, It is idea of the aggregation of information in groups, resulting in decisions that are often better than could have been made by any single member of the group.
+
+The Open Innovation also describe the similar idea on how companies commercialise trial knowledge - instead of only use internal ideas, it should also use external ideas to advance their technology.
 

### 2013-02-21 latex/paper_new.tex
Rewrote webgrid elicitation explanation, wrote up cluster analysis and explained what we did. Re-flowed to remove chronological-focused flow.

@@
 massive participation in it, has changed this relationship: we now see configurations of people interacting with content exposed via the Web and with each other, typified by social Web sites.
-Rather than drawing a line through such Web-based systems to separate the human and machine parts (as Computer Science has traditionally done), we can now draw a line around them  and  treat  each  such  assembly  as  a  'social  machine'  ñ  a  machine  comprising  digital  and  human
+Rather than drawing a line through such Web-based systems to separate the human and machine parts (as Computer Science has traditionally done), we can now draw a line around them  and  treat  each  such  assembly  as  a  'social  machine'    a  machine  comprising  digital  and  human
 components and the interactions between them.
@@
 where people refine data describing digital manuscripts.
-Åú People are elementary problem solvers. Systems in this category typically find innovative ways of getting
+ People are elementary problem solvers. Systems in this category typically find innovative ways of getting
 people to perform a task that are easy for them but for which no effective computer algorithm is available.
@@
 basic form of composition.
-Åú People generate/test partial  solutions. Systems  in  this  category provide  incentives  for people  to perform
+ People generate/test partial  solutions. Systems  in  this  category provide  incentives  for people  to perform
 complex  problem  solving  that  generates  a  design  to  be  tested  automatically  and/or  tests  automatically
@@
 	About the aggregation of information in groups, resulting in decisions that, he argues, are often better than could have been made by any single member of the group. 
-Human Computation (2005)	.	“...a paradigm for utilizing human processing power to solve problems that computers cannot yet solve.”
+Human Computation (2005)	.	Ò...a paradigm for utilizing human processing power to solve problems that computers cannot yet solve.Ó
 
-Crowdsourcing (2006)	“Crowdsourcing is the act of taking a job traditionally
-performed by a designated agent (usually an employee) and outsourcing it to an undefined, generally large group of people in the form of an open call.”
+Crowdsourcing (2006)	ÒCrowdsourcing is the act of taking a job traditionally
+performed by a designated agent (usually an employee) and outsourcing it to an undefined, generally large group of people in the form of an open call.Ó
 .	
-Collective Intelligence (2009)	"... groups of individuals doing things collectively that seem intelligent.”
+Collective Intelligence (2009)	"... groups of individuals doing things collectively that seem intelligent.Ó
 
@@
 \subsection{Methodology}
+
 In order to identify and classify social machines, we first embarked on a principled
@@
 we can use to classify social machines. We used the {\it repertory grid} elicitation
-technique~\cite{kelly} in order to derive an initial set of elements and constructs. We
-asked ten Computer Science researchers familiar with social machines to create their own
-repertory grids, and populate the elements from their own knowledge, and to create the
-constructs using the standard repertory grid elicitation technique. This exercise led
+technique~\cite{kelly} in order to derive an initial set of elements and constructs.
+
+To elicit knowledge using a repertory grid, software is used to ask users to describe constructs
+that differentiate between elements. For example, prompting the user to create a construct that
+differentiates between {\it GalaxyZoo} and {\it Wikipedia}. The user describes the opposing poles of 
+the construct, in this case, the user may decide on ``For Science'' and ``For Socialising''. The
+user then rates every element from 1 to 5 on this construct, where 1 represents an element that is
+``For Science'', and 5 if the element is ``For Socialising''; elements can also be rated between
+the poles if they may correspond to both or neither. The repertory grid software identifies
+elements that are the least differentiated when requesting new constructs, in an attempt to
+elicit as much knowledge as possible from the user. Ten Computer Science researchers familiar with social machines created their own
+repertory grids, and populate the elements from their own knowledge, and created the
+constructs using the standard repertory grid elicitation technique described above. This exercise led
 to 10 grids, the union of which comprised a total of 56 unique elements (social machines)
@@
 
-To make the set of social machines easier to refer to and more approachable, we performed
-an initial grid analysis of the machines, clustering them into related user-centered
+In addition to repertory grid analysis, we also employed a classification analysis of social
+machines, by extracting the most-tagged terms of each social network from the online
+bookmarking service {\it Delicious.com}, and thus achieving a crowd sourced cluster of
+classifications. This resulting in user-centered cluster 
 themes, of: Social Networks, (Micro)Blogging, News Aggregators, Image Boards, Crowd Science,

### 2013-02-21 latex/paper_new.tex
Added more about using the framework

@@
 In order to demonstrate usage of our framework's constructs, we ranked our social machines
-by their Alexa ranking, and performed a full repertory grid elicitation exercise. The
+by their Alexa ranking, and performed a full repertory grid elicitation exercise over them.
+This process involved ranking each social machine against the constructs, rating each construct
+from 1 to 5. In total, 680 ranking were made, and the
 results are illustrated in a grid and accompanying dendrograms for element and construct
@@
 
-[needs some discussion on what this shows.]
+From the dendrograms we can get a sense of similarities and separation of social machines. For
+instance, it is clear that YouTube, Vimeo, Reddit and Digg are similar, which is logical, because
+they are all social machines that focus on sharing content. Likewise, Quora and StackOverflow
+are also similar, and again, this makes sense, because they are both question answering social
+machines. There is also a dendrogram for the constructs, to show correlations between them. This
+dendrogram indicates that there is a correlation between the ``Single Participant Roles'' and
+``No Hierarchy in Participant Roles'' constructs, which is because they refer to the same
+aspect of social machines. Another, but less obvious similarity is between the ``Large variety of
+social features'' and ``Benefit NOT for the world/society'', which indicates that of the social
+machines we have rated, there is a correlation between social features and social machines that
+do not benefit the world at large, and vice-versa.
+
+We envisage two key initial uses for this framework: for classifying social machines; and for
+predicting behaviour of new/future social machines. For classifying social machines, it will
+enable direct comparisons to be made with existing machines, as well as provide a platform
+for discussions about social machines. We also think that the data from a large-scale classification
+exercise may be useful for researchers who wish to apply prediction techniques (such as SVMs)
+to social machines, their behaviour and impact.
 

### 2013-02-21 latex/paper_new.tex
Simplified motivation section, removed Fig2

@@
 
-We identified a hierarchy of constructs pertaining to the motivation people have for using social
-machines with top level constructs of:
+%We identified a hierarchy of constructs pertaining to the motivation people have for using social
+%machines with top level constructs of:
+%
+%\begin{enumerate}
+%\item Hedonistic
+%\item Financial
+%\item To be informed
+%\item To help others
+%\end{enumerate}
+%
+%The sub constructs are illustrated in Figure~\ref{motivation}, and we relate them to Maslow's hierarchy of needs~\cite{maslow}, (e.g., self-actualisation, security, respect of others), in various ways.
+%
+%\begin{figure*}[htb]
+%\begin{center}
+%\includegraphics[width=18cm]{img/motivation.pdf}
+%\caption{Motivation hierarchy of participants using Social Machines, related to Maslow's hierarchy of needs.} \label{motivation}
+%\end{center}
+%\end{figure*}
 
-\begin{enumerate}
-\item Hedonistic
-\item Financial
-\item To be informed
-\item To help others
-\end{enumerate}
-
-The sub constructs are illustrated in Figure~\ref{motivation}, and we relate them to Maslow's hierarchy of needs~\cite{maslow}, (e.g., self-actualisation, security, respect of others), in various ways.
-
-\begin{figure*}[htb]
-\begin{center}
-\includegraphics[width=18cm]{img/motivation.pdf}
-\caption{Motivation hierarchy of participants using Social Machines, related to Maslow's hierarchy of needs.} \label{motivation}
-\end{center}
-\end{figure*}
-
-For each of these motivations, different roles that participate may have different {\it core motivations}, such as:
+The motivations of users can be identified and mapped to existing human needs and motivations
+systems, such as Maslow's hierarchy of human needs~\cite{maslow} (e.g., self-actualisation, security
+respect of others). For each of these motivations, different roles that participate may have different {\it core motivations}, such as:
 

### 2013-02-21 latex/paper_new.tex
motivation links in

@@
 
+Motivation in specific social machines has been studied, including motivation of participants
+in: Wikipedia~\cite{kuznetsov2006motivations}; Open Source Software~\cite{lakhani2003hackers};
+Peer-to-peer filesharing~\cite{p2p}; Tagging~\cite{tagging}; and Virtual
+Communities~\cite{ardichvili2003motivation,Moore:2007:UMM:1235000.1235035}.
+
+[relate this work back to ours briefly.]
+
 \subsubsection{Artefacts of Social Machines}

### 2013-02-21 latex/paper_new.tex
related work

@@
 The Social computing is another shift, which is a generic term refers systems that support ``the gathering, representation, processing, use, and dissemination of information that is distributed across social collectivities such as teams, communities, organizations, and markets''.
-To compare with the CSC, the social computing has more focus in computing rather than broader sense of
+To compare with the CSC, the social computing has more focus in `computing' rather than `collaboration'.
+Similarly, the Human Computation is another paradigm towards this computational direction, it utilize human processing power to solve problems that computers cannot yet solve.
+All these three mentioned fields has overlapping in the their concepts and they are only mainly popularised in the computer science.
+In the Web 2.0 era, there are few new concepts/buzzword emerged
 

### 2013-02-21 latex/paper_new.tex
related work section update

--- 2013-02-21 latex/paper_new.tex
@@
 The Computer-supported Collaboration Work (CSCW) emerged in 1982 represent such shift.
-It addresses ``how collaborative activities and their coordination can be supported by means of computer systems'' later evolved in a more broader sense of  Computer-supported collaboration (CSC).
+It addresses ``how collaborative activities and their coordination can be supported by means of computer systems''.
+This concept later evolved into a more broader sense of  Computer-supported collaboration (CSC).
 The Social computing is another shift, which is a generic term refers systems that support ``the gathering, representation, processing, use, and dissemination of information that is distributed across social collectivities such as teams, communities, organizations, and markets''.
-To compare with the CSC, the social computing has more focus in computing rather than broader sense of
+To compare with the CSC, the social computing has more focus in `computing' rather than `collaboration'.
+Similarly, the Human Computation is another paradigm towards this computational direction, which utilize human processing power to solve problems that computers cannot yet solve.
+All these three mentioned fields has overlapping in the their concepts and they are mainly popularised in the computer science.
+In the Web 2.0 era, there are few new broader concepts/buzzword emerged on the idea of utilising the human power which are not only used in computer science, but also in other fields such as social science, business etc.
+For example, crowding sourcing is an idea that taking a job traditionally performed by a designated agent and outsourcing it to an undefined, generally large group of people in the form of an open call.
+As it is argued in Surowieckis the wisdom of crowds that the aggregation of information in groups, resulting in decisions that are often better than could have been made by any single member of group.
+Open Innovation advocate the similar idea with empathising on the innovation in firms.
+
+Open Innovation is another example,
 

### 2013-02-21 latex/paper_new.tex
related work section update

@@
 For example, crowding sourcing is an idea that taking a job traditionally performed by a designated agent and outsourcing it to an undefined, generally large group of people in the form of an open call.
-As it is argued in Surowieckis the wisdom of crowds that the aggregation of information in groups, resulting in decisions that are often better than could have been made by any single member of group.
 Open Innovation advocate the similar idea with empathising on the innovation in firms.
+Surowiecki also argued in the book of the wisdom of crowds that the aggregation of information in groups, resulting in decisions that are often better than could have been made by any single member of group.
 

### 2013-02-21 latex/paper_new.tex
related work section updates

@@
 Surowiecki also argued in the book of the wisdom of crowds that the aggregation of information in groups, resulting in decisions that are often better than could have been made by any single member of group.
+Collective Intelligence described by Malone, defines a even more broadly term that ``groups of individuals doing things collectively that seem intelligent.''
+The relationship between human computation, crowdsourcing, social computing and collective intelligence is discussed in . 
+In  social machines  the  social and  the  technical are equal  partners.
+While  crowdsourcing  could  be  viewed  as  'human  subroutines'  and  'citizen  sensors',  the social machine perspective is one of incentivised participation and not humans enslaved to computers, as described in Figure \ref{socialmachine}.
 
-Open Innovation is another example,
-
-CSCW (1982)	"how collaborative activities and their coordination can be supported by means of computer systems."
-
-Social Computing (1994)	the system that support the gathering, representation, processing, use, and dissemination of information that is distributed across social collectivities such as teams, communities, organizations, and markets.
-
-Open Innovation (2003)	Paradigm in how companies commercialize trial knowledge.
-
-Is a paradigm that assumes that firms can and should use external ideas as well as internal ideas, and internal and external path to market, as the firms look to advance their technology.
-It combines internal and external ideas into architectures and systems whose requirements are defined by a business model.
-The Wisdom of Crowds (2004)
-	About the aggregation of information in groups, resulting in decisions that, he argues, are often better than could have been made by any single member of the group. 
-Human Computation (2005)	.	Ò...a paradigm for utilizing human processing power to solve problems that computers cannot yet solve.Ó
-
-Crowdsourcing (2006)	ÒCrowdsourcing is the act of taking a job traditionally
-performed by a designated agent (usually an employee) and outsourcing it to an undefined, generally large group of people in the form of an open call.Ó
-.	
-Collective Intelligence (2009)	"... groups of individuals doing things collectively that seem intelligent.Ó
-
-The Wisdom of Crowds described by Surowiecki, It is idea of the aggregation of information in groups, resulting in decisions that are often better than could have been made by any single member of the group.
-
-The Open Innovation also describe the similar idea on how companies commercialise trial knowledge - instead of only use internal ideas, it should also use external ideas to advance their technology.
-
-Web 2.0
-
-Collective Intelligence, Wisdom of the crowds
-
-Open Innovation
-
-In  social machines  the  social and  the  technical are equal  partners. While  crowdsourcing  could  be  viewed  as  'human  subroutines'  and  'citizen  sensors',  the
-social machine perspective is one of incentivised participation and not humans enslaved to computers.
-
-Crowdsourcing, human computation
-
-CSCW, social computing
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=8cm]{img/socialmachine.JPG}
+\caption{Social Machine} \label{socialmachine}
+\end{center}
+\end{figure}
 

### 2013-02-21 latex/paper_new.tex
related work section updates

@@
 A Social Machine includes human and machine two major components, therefore, the study of social machine is an interdisciplinary endeavour.
-Since the Turning machine was invented, there has been a `paradigm shift' in using computers, where the computer facilitates human communication rather acts as a purely calculative device.
-The Computer-supported Collaboration Work (CSCW) emerged in 1982 represent such shift.
+Since the Turning machine was invented, there has been a `paradigm shift' \cite{kuhn1996structure} in using computers, where the computer facilitates human communication rather acts as a purely calculative device.
+The Computer-supported Collaboration Work (CSCW) emerged in 1984 represent such shift.
 It addresses ``how collaborative activities and their coordination can be supported by means of computer systems''.
 This concept later evolved into a more broader sense of  Computer-supported collaboration (CSC).
-The Social computing is another shift, which is a generic term refers systems that support ``the gathering, representation, processing, use, and dissemination of information that is distributed across social collectivities such as teams, communities, organizations, and markets''.
+The Social computing \cite{parameswaran2007research} is another shift, which is a generic term refers systems that support ``the gathering, representation, processing, use, and dissemination of information that is distributed across social collectivities such as teams, communities, organizations, and markets''.
 To compare with the CSC, the social computing has more focus in `computing' rather than `collaboration'.
-Similarly, the Human Computation is another paradigm towards this computational direction, which utilize human processing power to solve problems that computers cannot yet solve.
+Similarly, the Human Computation \cite{von2009human} is another paradigm towards this computational direction, which utilize human processing power to solve problems that computers cannot yet solve.
 All these three mentioned fields has overlapping in the their concepts and they are mainly popularised in the computer science.
 In the Web 2.0 era, there are few new broader concepts/buzzword emerged on the idea of utilising the human power which are not only used in computer science, but also in other fields such as social science, business etc.
-For example, crowding sourcing is an idea that taking a job traditionally performed by a designated agent and outsourcing it to an undefined, generally large group of people in the form of an open call.
-Open Innovation advocate the similar idea with empathising on the innovation in firms.
+For example, crowding sourcing \cite{howe2006crowdsourcing} is an idea that taking a job traditionally performed by a designated agent and outsourcing it to an undefined, generally large group of people in the form of an open call.
+Open Innovation \cite{chesbrough2008open} advocates the similar idea with empathising on the innovation in firms.
 Surowiecki also argued in the book of the wisdom of crowds that the aggregation of information in groups, resulting in decisions that are often better than could have been made by any single member of group.

### 2013-02-21 latex/paper_new.tex
related work section updates

@@
 Open Innovation \cite{chesbrough2008open} advocates the similar idea with empathising on the innovation in firms.
-Surowiecki also argued in the book of the wisdom of crowds that the aggregation of information in groups, resulting in decisions that are often better than could have been made by any single member of group.
-Collective Intelligence described by Malone, defines a even more broadly term that ``groups of individuals doing things collectively that seem intelligent.''
-The relationship between human computation, crowdsourcing, social computing and collective intelligence is discussed in . 
+Surowiecki also argued in the book of the wisdom of crowds \cite{surowiecki2005wisdom} that the aggregation of information in groups, resulting in decisions that are often better than could have been made by any single member of group.
+Collective Intelligence described by Malone \cite{malone2009harnessing}, defines a even more broadly term that ``groups of individuals doing things collectively that seem intelligent.''
+The relationship between human computation, crowdsourcing, social computing and collective intelligence is discussed in \cite{quinn2011human}. 
 In  social machines  the  social and  the  technical are equal  partners.

### 2013-02-21 latex/paper_new.tex
related work section updates

@@
 It addresses ``how collaborative activities and their coordination can be supported by means of computer systems''.
-This concept later evolved into a more broader sense of  Computer-supported collaboration (CSC).
-The Social computing \cite{parameswaran2007research} is another shift, which is a generic term refers systems that support ``the gathering, representation, processing, use, and dissemination of information that is distributed across social collectivities such as teams, communities, organizations, and markets''.
+This concept later evolved into a broader sense of  Computer-supported collaboration (CSC).
+The Social computing \cite{parameswaran2007research} is another shift, which refers systems that support ``the gathering, representation, processing, use, and dissemination of information that is distributed across social collectivities such as teams, communities, organizations, and markets''.
 To compare with the CSC, the social computing has more focus in `computing' rather than `collaboration'.

### 2013-02-21 latex/paper_new.tex
related work section updates

@@
 Similarly, the Human Computation \cite{von2009human} is another paradigm towards this computational direction, which utilize human processing power to solve problems that computers cannot yet solve.
-All these three mentioned fields has overlapping in the their concepts and they are mainly popularised in the computer science.
-In the Web 2.0 era, there are few new broader concepts/buzzword emerged on the idea of utilising the human power which are not only used in computer science, but also in other fields such as social science, business etc.
-For example, crowding sourcing \cite{howe2006crowdsourcing} is an idea that taking a job traditionally performed by a designated agent and outsourcing it to an undefined, generally large group of people in the form of an open call.
+All these three mentioned fields overlap in their concepts and they are mainly popularised in the computer science.
+In the Web 2.0 era, there were few generic concepts/buzzword emerged on the idea of utilising the human power which are not only used in computer science, but also in other fields such as social science, business etc.
+Crowding sourcing \cite{howe2006crowdsourcing} as an example, describes the idea that taking a job traditionally performed by a designated agent and outsourcing it to an undefined, generally large group of people in the form of an open call.
 Open Innovation \cite{chesbrough2008open} advocates the similar idea with empathising on the innovation in firms.

### 2013-02-21 latex/paper_new.tex
related work section updates

@@
 Crowding sourcing \cite{howe2006crowdsourcing} as an example, describes the idea that taking a job traditionally performed by a designated agent and outsourcing it to an undefined, generally large group of people in the form of an open call.
-Open Innovation \cite{chesbrough2008open} advocates the similar idea with empathising on the innovation in firms.
+Open Innovation \cite{chesbrough2008open} advocates the similar idea with empathizing on the innovation in firms.
 Surowiecki also argued in the book of the wisdom of crowds \cite{surowiecki2005wisdom} that the aggregation of information in groups, resulting in decisions that are often better than could have been made by any single member of group.
-Collective Intelligence described by Malone \cite{malone2009harnessing}, defines a even more broadly term that ``groups of individuals doing things collectively that seem intelligent.''
-The relationship between human computation, crowdsourcing, social computing and collective intelligence is discussed in \cite{quinn2011human}. 
+Collective Intelligence described by Malone \cite{malone2009harnessing}, defines an even more broadly concept that ``groups of individuals doing things collectively that seem intelligent.''
+The concept overlapping between human computation, crowdsourcing, social computing and collective intelligence is discussed in \cite{quinn2011human}. 
 In  social machines  the  social and  the  technical are equal  partners.

### 2013-02-21 latex/paper_new.tex
related work section updates

@@
 In  social machines  the  social and  the  technical are equal  partners.
-While  crowdsourcing  could  be  viewed  as  'human  subroutines'  and  'citizen  sensors',  the social machine perspective is one of incentivised participation and not humans enslaved to computers, as described in Figure \ref{socialmachine}.
+While  crowdsourcing  could  be  viewed  as  'human  subroutines'  and  'citizen  sensors',  the social machine perspective is one of incentivised participation and not humans enslaved to computers.
+We define the scope of social machine in a conceptual digram in Figure \ref{socialmachine}.
 

### 2013-02-21 latex/paper_new.tex
related work section updates

@@
 
-A Social Machine includes human and machine two major components, therefore, the study of social machine is an interdisciplinary endeavour.
+A Social Machine includes human and machine two major components, therefore the study of social machine is an interdisciplinary endeavour.
 Since the Turning machine was invented, there has been a `paradigm shift' \cite{kuhn1996structure} in using computers, where the computer facilitates human communication rather acts as a purely calculative device.
-The Computer-supported Collaboration Work (CSCW) emerged in 1984 represent such shift.
+The Computer-supported Collaboration Work (CSCW) \cite{grudin1994computer} emerged in 1984 represents such shift.
 It addresses ``how collaborative activities and their coordination can be supported by means of computer systems''.
@@
 To compare with the CSC, the social computing has more focus in `computing' rather than `collaboration'.
-Similarly, the Human Computation \cite{von2009human} is another paradigm towards this computational direction, which utilize human processing power to solve problems that computers cannot yet solve.
-All these three mentioned fields overlap in their concepts and they are mainly popularised in the computer science.
+Similarly, the Human Computation \cite{von2009human} is another paradigm towards the computational direction, which utilize human processing power to solve problems that computers cannot yet solve.
+All these three mentioned fields overlap in their concepts and are mainly popularised in the computer science.
 In the Web 2.0 era, there were few generic concepts/buzzword emerged on the idea of utilising the human power which are not only used in computer science, but also in other fields such as social science, business etc.
@@
 \includegraphics[width=8cm]{img/socialmachine.JPG}
-\caption{Social Machine} \label{socialmachine}
+\caption{The scope of the Social Machine} \label{social machine}
 \end{center}

### 2013-02-22 latex/paper_new.tex
related work section updates
Removed headyness

@@
 \section{Social machines: an introduction}
-Once upon a time 'machines' were programmed by programmers and used by users. The Web, and the
-massive participation in it, has changed this relationship: we now see configurations of people interacting with content exposed via the Web and with each other, typified by social Web sites.
-Rather than drawing a line through such Web-based systems to separate the human and machine parts (as Computer Science has traditionally done), we can now draw a line around them  and  treat  each  such  assembly  as  a  'social  machine'    a  machine  comprising  digital  and  human
-components and the interactions between them.
-This crucial transition in thinking acknowledges the reality of today's computing systems and is essential to underpin any understanding of the science and engineering of our future socio-technical systems: we view an ecosystem of co-evolving social machines.
+Once upon a time 'machines' were programmed by programmers and used by users.
+The success of the Web has changed this relationship: we now see configurations of people interacting with content and with each other, typified by social Web sites.
+Rather than drawing a line through such Web-based systems to separate the human and machine parts (as Computer Science has traditionally done), we can now draw a line around them  and  treat  each  such  assembly  as  a  'social  machine'  -  a  machine  in which  digital  and  human components are seamlessly interwoven.
+This crucial transition in thinking acknowledges the reality of current socio-technical systems, and is essential to underpin any understanding of the science and engineering of their future development towards pervasive ecosystems of co-evolving social machines.
 
-The  typical  social Web  site  is  a  social machine, with  its user-generated  content  and  social network evolving dynamically  and  hand-in-hand.
-More  than  just  'data' or  'content',  these sites have  some notion of  'social artefacts' around which we forge the links in the social network, like photos on flickr or movies on YouTube, and they  can also be  flourishing  spaces  for human  ingenuity.
-Today we  see  a  divide  between  conventional  computer  systems  dedicated  to  data-  and  computation-intensive tasks, and social computing like the Web 2.0 sites.
-But many of our challenges, from healthcare to transport, need  both  of  these  axes:  a  sophisticated  combination  of  data-intensive  automation  and  deep  social involvement.
-This suggests the need for new social machines to tackle new problems, and to achieve this we
-need to understand the science and engineering of social machines.
+A number of prominent examples aside, today we still see a divide between conventional IT systems dedicated to data- and computation-intensive tasks, and Web $2.0$ sites offering some combination of well-known participatory features, in which user-generated content and the underlying social network evolve dynamically and hand-in-hand.
+As technology becomes more and more ubiquitous, many of the challenges we will witness in any domain of our life, business, and society will need solutions that rely on both of these axes: a sophisticated  combination  of  data-intensive, complex automation and deep community involvement.
+This suggests the need for new types of systems to tackle these challenges, and to achieve requires a through understanding of the science and engineering of social machines.
 
+As a first step, we propose a classification framework for social machines, which can be used to describe, study, and compare this expanding field of interdisciplinary research and the increasing number of systems combining human and computational intelligence to serve purposes as diverse as traffic management, healthcare, media and journalism, or design and natural sciences.
+
+In this paper we introduce the current version of the framework, which is the result of a knowledge elicitation process performed internally in our lab.
+Our aim is to present it to the broader community and extend and revise it based on their feedback in order to achieve a shared understanding, terminology, and set of descriptive features to enable a systematic study of existing and future social machines.
+Using a common descriptive classification allows researchers to observe the effects of specific technical properties of a system on social behavior, identify design and evolution patterns, learn about social network formation and dynamics, and devise incentive mechanisms to encourage large-scale participation.
+A second audience group targeted by our work are developers of social machines.
+For them, the framework may provide a theoretical grounding for the realization of new systems, informing system design and operation in terms of critical features and predicted user behavior and community development.
+
+\section{What are social machines?}
+Social machines can be basically characterized as assemblies of human and automatic components and their interaction.
 A  traditional  database,  with  users  looking  up  records  independently  of  each  other  and  an  administrator responsible for content, has some of the right ingredients but there is really no social component in this strict provider-consumer relationship.
-When we fill out a form (e.g.
-health information, birds spotted in my garden) it  is minimally  a  social machine  because humans  are  part  of  the  'social  computation'.
-This  becomes  richer when  the database  is  curated by members  of  the  community  (e.g.
-Wikipedia) and when  the  social network adds value implicitly (amazon) or explicitly (Facebook).
+When we fill out a form (e.g., health information, birds spotted in my garden) it  is minimally  a  social machine  because humans  are  part  of  the  'social  computation'.
+This  becomes  richer when  the database  is  curated by members  of  the  community  (e.g., Wikipedia) and when  the  social network adds value implicitly (Amazon) or explicitly (Facebook).
 We anticipate many categories of social machine in this continuum, and they may manifest themselves as Web site or app, or indeed in the Internet of Things.
 
-An emerging, interdisciplinary field of research and development with a growing number of social machine instances and studies about them.
+To describe social machines we roughly rely on four different categories of classification features:
 
-We propose a classification framework for social machines.
+\begin{description}
+  \item [The types of tasks undertaken by the social component and the resulting contributions]: social machines, just as any computing system, are built with a given purpose in mind, and exhibit functionality by which this purpose can be achieved.
+Users and the emerging social network interact with the system and with each other to carry out specific tasks by which they contribute to the 'computation'.
+This includes the creation and refinement of digital artifacts by executing given actions, possibly in collaboration with other participants.
+For example, in GalaxyZoo\footnote{\url{http://www.galaxyzoo.org}} people classify galaxies with collective performance as good as professional astronomers, while in reCaptcha\footnote{\url{http://www.google.com/recaptcha}} they refine the data that describes digital manuscripts.
+Sometimes, systems look for innovative ways of getting people to engage with the systems, including game narratives, like in EteRNA\footnote{\url{http://eterna.cmu.edu}}, in which  participants design RNA  that is evaluated automatically using simulators and then refined via human effort, or monetary rewards, like in Amazon's Mechanical Turk, a platform in which users undertake routine (micro) tasks such as object classification, audio transcription and text translation.
 
-Targeted audiences
+  \item [The roles of the participants and the social network they form]:
 
-* research community: shared terminology, set of key features and their possible instantiations based on the analysis of a relevant set of 50 (?) social machines (both popular systems as well as long tail systems); facilitates the study of design and evolution patterns; supports research on how communities emerge and develop and organisational studies (mechanism design, incentives); different research approaches can be better compared and aligned.
+  \item [The motivating factors and incentive mechanisms]: system features influencing user behavior and the social network dynamics
 
-* developer community: provides a theoretical grounding for the realisation of new systems, informs system design and operation (what features should be implemented, what features are more important than others); enables predictions on user behaviour and community development.
-
-The framework is based on interviews with XXX researchers covering XXX systems and was preliminary evaluated with respect to ...
-
-Dimensions of the classification
-
-* tasks and types of contributions: what does the system do, how do participants contribute: from supplying data and creating content to solving problems or validating solutions.
-
-* social network and interaction: roles, hierarchies of roles, profiling
-
-* motivation and incentives: system features influencing user behavior and the social network dynamics
-
-* human-machine interaction: how are human contributions used to compute things, including quality control, spam, workflow, aggregation of results
-
-People  supply  or  refine  data.    Systems  in  this  category  use  people  as,  effectively,  sophisticated  sensors.
-Examples of  this  sort of  system  are Galaxy  Zoo  [www.galaxyzoo.org] where people  classify  galaxies with
-collective performance as good as professional astronomers and reCaptcha [www.google.com/recaptcha]
-where people refine data describing digital manuscripts.
- People are elementary problem solvers. Systems in this category typically find innovative ways of getting
-people to perform a task that are easy for them but for which no effective computer algorithm is available.
-Examples  of  this  sort  of  system  are Mechanical  Turk  [www.mturk.com]  and  the  related  Turkit  system
-[groups.csail.mit.edu/uid/turkit]  that  allows  these  actions  to  be  strung  together  iteratively,  thus  giving  a
-basic form of composition.
- People generate/test partial  solutions. Systems  in  this  category provide  incentives  for people  to perform
-complex  problem  solving  that  generates  a  design  to  be  tested  automatically  and/or  tests  automatically
-generated  designs.    Examples  include  EteRNA  [eterna.cmu.edu]  in which  participants  design RNA  that  is
-evaluated automatically using simulators and then refined via human effort; Matlab programming contests
-[www.mathworks.co.uk/matlabcentral/contest] where people compete for the best solution as judged by automated  tests  that  compute  the  efficiency  of  the  code;  or  Scratch  [scratch.mit.edu]  in  which  simple programs are created by some participants and modified by others in a remixing process.
+  \item [Interaction between social and machine component]: how are human contributions used to compute things, including quality control, spam, workflow, aggregation of results
+\end{description}
 
@@
 While  crowdsourcing  could  be  viewed  as  'human  subroutines'  and  'citizen  sensors',  the social machine perspective is one of incentivised participation and not humans enslaved to computers.
+We define the scope of social machine in a venn digram in Figure \ref{socialmachine}.
+
+The study of social machines is naturally an interdisciplinary endeavor.
+Since the Turning machine was invented, we have experienced a paradigm shift in computer science from computers as purely calculative devices into a worldwide interconnected computer infrastructure which facilitates almost any type of human endeavor.
+The Computer-supported Collaboration Work (CSCW) \cite{grudin1994computer} emerged in 1984 represents such shift.
+It addresses ``how collaborative activities and their coordination can be supported by means of computer systems''.
+This concept later evolved into a broader sense of  Computer-supported collaboration (CSC).
+The Social computing \cite{parameswaran2007research} is another shift, which refers systems that support ``the gathering, representation, processing, use, and dissemination of information that is distributed across social collectivities such as teams, communities, organizations, and markets''.
+To compare with the CSC, the social computing has more focus in `computing' rather than `collaboration'.
+Similarly, the Human Computation \cite{von2009human} is another paradigm towards the computational direction, which utilize human processing power to solve problems that computers cannot yet solve.
+All these three mentioned fields overlap in their concepts and are mainly popularised in the computer science.
+In the Web 2.0 era, there were few generic concepts/buzzword emerged on the idea of utilising the human power which are not only used in computer science, but also in other fields such as social science, business etc.
+Crowding sourcing \cite{howe2006crowdsourcing} as an example, describes the idea that taking a job traditionally performed by a designated agent and outsourcing it to an undefined, generally large group of people in the form of an open call.
+Open Innovation \cite{chesbrough2008open} advocates the similar idea with empathizing on the innovation in firms.
+Surowiecki also argued in the book of the wisdom of crowds \cite{surowiecki2005wisdom} that the aggregation of information in groups, resulting in decisions that are often better than could have been made by any single member of group.
+Collective Intelligence described by Malone \cite{malone2009harnessing}, defines an even more broadly concept that ``groups of individuals doing things collectively that seem intelligent.''
+The concept overlapping between human computation, crowdsourcing, social computing and collective intelligence is discussed in \cite{quinn2011human}.
+In  social machines  the  social and  the  technical are equal  partners.
+While  crowdsourcing  could  be  viewed  as  'human  subroutines'  and  'citizen  sensors',  the social machine perspective is one of incentivised participation and not humans enslaved to computers.
 We define the scope of social machine in a conceptual digram in Figure \ref{socialmachine}.
@@
 \begin{center}
-\includegraphics[width=8cm]{img/socialmachine.JPG}
-\caption{The scope of the Social Machine} \label{social machine}
+\includegraphics[width=8cm]{img/socialmachinescope.png}
+\caption{Social machines and related areas} \label{social machine}
 \end{center}
@@
 
-In order to identify and classify social machines, we first embarked on a principled
-knowledge elicitation process~\cite{knowledgeelicitation}, to determine the {\it elements}
-(i.e., the social machines) and the {\it constructs} (i.e., the classifying factors) that
-we can use to classify social machines. We used the {\it repertory grid} elicitation
-technique~\cite{kelly} in order to derive an initial set of elements and constructs.
+The classification framework presented in this paper was created through knowledge elicitation ~\cite{knowledgeelicitation}. In particular we used the {\it repertory grid} elicitation
+technique~\cite{kelly} in order to derive an initial set of {\it elements}, which represent instances of social machines, and {\it constructs}, which capture their most important characteristics.
 
-To elicit knowledge using a repertory grid, software is used to ask users to describe constructs
-that differentiate between elements. For example, prompting the user to create a construct that
-differentiates between {\it GalaxyZoo} and {\it Wikipedia}. The user describes the opposing poles of 
-the construct, in this case, the user may decide on ``For Science'' and ``For Socialising''. The
-user then rates every element from 1 to 5 on this construct, where 1 represents an element that is
-``For Science'', and 5 if the element is ``For Socialising''; elements can also be rated between
-the poles if they may correspond to both or neither. The repertory grid software identifies
-elements that are the least differentiated when requesting new constructs, in an attempt to
-elicit as much knowledge as possible from the user. Ten Computer Science researchers familiar with social machines created their own
-repertory grids, and populate the elements from their own knowledge, and created the
-constructs using the standard repertory grid elicitation technique described above. This exercise led
-to 10 grids, the union of which comprised a total of 56 unique elements (social machines)
-and 117 different constructs.
+In repertory grid elicitation a software tool is used to ask users to describe constructs
+that differentiate between elements - for example,  prompting the user to create a construct that
+differentiates between {\it GalaxyZoo} and {\it Facebook}, as two prominent example of social machines our classification framework should apply to.
+The user describes the opposing poles of the construct - in this case, the user may decide on 'For Science' and 'For Socialising' in order to capture the core distinction between the two system.
+The user then rates every element with value from $1$ to $5$ on this construct, where $1$ represents an element that is purely
+'For Science', and $5$ one that mainly serves  'socialising' purposes.
+Elements can also be rated with values between the poles if they may correspond to both or neither.
+The repertory grid software then identifies elements that are the least differentiated when requesting new constructs, in an attempt to elicit as much knowledge as possible from the user.
+We applied this technique to create an initial set of examples of social machines and to classify them according to specific features.
+To do so, we asked then computer science researchers familiar with the field to create their own repertory grids, and populate the elements from their own knowledge, and created the constructs using the standard repertory grid elicitation technique described above.
+This exercise led to $10$ grids, the union of which comprised a total of $56$ unique elements (social machines)
+and $117$ different constructs (classifying factors).
+As the aim of this initial phase was to understand how people perceive the notion of social machine and their most distinctive properties, we allowed the participants to choose the social machines they are familiar with and describe them in their own terms.
 
-In addition to repertory grid analysis, we also employed a classification analysis of social
-machines, by extracting the most-tagged terms of each social network from the online
-bookmarking service {\it Delicious.com}, and thus achieving a crowd sourced cluster of
-classifications. This resulting in user-centered cluster 
-themes, of: Social Networks, (Micro)Blogging, News Aggregators, Image Boards, Crowd Science,
-Answer Gardens, Community Watch, Health and Wellbeing Support, Action and Investigation,
-Opinion Sharing, Video Sharing, Photo Sharing, Code Sharing, Art Sharing, Crowdsourcing
-Platforms, Mash-up Systems, Crowdsourcing Toolkits and Platforms.
+%To enrich the resulting list of classifying constructs we then extracted the most frequent tags used in connection to the $56$ systems mentioned by the participants from the online bookmarking service {\it delicious.com}, thus achieving a crowdsourced social-machine classification containing the following clusters: Social Networks, (Micro)Blogging, News Aggregators, Image Boards, Crowd Science, Answer Gardens, Community Watch, Health and Wellbeing Support, Action and Investigation, Opinion Sharing, Video Sharing, Photo Sharing, Code Sharing, Art Sharing, Crowdsourcing Platforms, Mash-up Systems, and Crowdsourcing Toolkits and Platforms.
 
 While determining the intersection of elements was straightforward, the consolidation of the
-constructs required a more thorough process. To consolidate the constructs we embarked on a manual process of grouping the constructs
-into rough clusters, based around the areas they cover. For example, we used construct
+constructs required a more thorough process. We embarked on a manual process of grouping the constructs into rough clusters, based around the areas they cover. For example, we used construct
 clusters including: Users, Motivation, Popularity, Technology, and Purpose. We then examined
@@
 process involved four of the authors discussing the choices, and finally agreeing on a
-consolidated set of 31 constructs.
+consolidated set of $31$ constructs.
 
-Even with this reduced set of constructs, if a single person were to classify all 56
-elements, this would still require 1736 individual decisions, and hence is unfeasible to
-expect full coverage of the elements using all of these constructs, despite how representative
-they are.
-
-[something here about doing 2 grids each and verifying the results?]
-
-To extract greater usefulness, and the ability to use our constructs to classify social
-machines we re-evaluated the constructs and the semantics that they represent into richer
-constructs that are not constrained by the 1--5 rating of repertory grid constructs.
+%Even with this reduced set of constructs, if a single person were to classify all 56
+%elements, this would still require 1736 individual decisions, and hence is unfeasible to
+%expect full coverage of the elements using all of these constructs, despite how representative
+%they are.
+%
+%[something here about doing 2 grids each and verifying the results?]
+%
+%To extract greater usefulness, and the ability to use our constructs to classify social
+%machines we re-evaluated the constructs and the semantics that they represent into richer
+%constructs that are not constrained by the 1-5 rating of repertory grid constructs.
 

### 2013-02-22 latex/paper_new.tex
Split dendrograms into two images

@@
 \hline
-Domain specificity [1 - Specific / 5 - General] vs \\
-\hspace{0.66cm} Generality of Purpose \\
+Domain specificity [1 - Specific / 5 - General] \\
 Visibility and importance of a participants `reputation' [1 - Not important / 5- Important] \\
-\hspace{0.66cm} Explicit representation of participant reputation [1 - No Reputations / 5 - Reputations] \\
+Explicit representation of participant reputation [1 - No Reputations / 5 - Reputations] \\
 Participants can define new types of contributions [1 - Not possible / 5 - Complete freedom] \\
@@
 \hline
-(Level of social features) - Supports social interaction [1 - No social features / 5 - Large variety of social features] \\
+Supports social interaction [1 - No social features / 5 - Large variety of social features] \\
 How often a participant participates (times a day, etc) [1 - Many times a day / 5 Seldom] \\
@@
 Quality of participant contributions [1 - Low quality / 5 - High quality] \\
-(Different roles clearly defined) - \\
-\hspace{0.66cm} Clear separation of roles/responsibilities among participants - [1 - Single role - Everyone is the same) / 5 - People assume different roles/responsibilities] \\
+Clear separation of roles/responsibilities among participants - [1 - Single role - Everyone is the same) / 5 - People assume different roles/responsibilities] \\
 Extent of hierarchical organisation of roles [1 - No Hierarchy / 5 - Deep/Well-defined Hierarchy] \\
@@
 
-\begin{figure*}
+\begin{figure}
 \begin{center}
-\includegraphics[width=18cm]{img/dendrogram.png}
-\caption{Dendrogram from a repertory grid exercise of the top 20 (ranked by Alexa) social machines from our element set, against our consolidated constructs.} \label{dendrogram}
+\includegraphics[width=4cm]{img/dendrogram-elements.png}
+\caption{Dendrogram of elements (social machines) from a repertory grid exercise of the top 20 (ranked by Alexa) social machines from our element set, against our consolidated constructs.} \label{dendrogram}
 \end{center}
-\end{figure*}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=8cm]{img/dendrogram-constructs.png}
+\caption{Dendrogram of constructs from a repertory grid exercise of the top 20 (ranked by Alexa) social machines from our element set, against our consolidated constructs.} \label{dendrogram}
+\end{center}
+\end{figure}
 

### 2013-02-22 latex/paper_new.tex
Polyarchy simplified and reduced in size.
Added refined list of constructs, removed sections and added outline of new section

@@
 
-To describe social machines we roughly rely on four different categories of classification features:
-
-\begin{description}
-  \item [The types of tasks undertaken by the social component and the resulting contributions]: social machines, just as any computing system, are built with a given purpose in mind, and exhibit functionality by which this purpose can be achieved.
-Users and the emerging social network interact with the system and with each other to carry out specific tasks by which they contribute to the 'computation'.
-This includes the creation and refinement of digital artifacts by executing given actions, possibly in collaboration with other participants.
-For example, in GalaxyZoo\footnote{\url{http://www.galaxyzoo.org}} people classify galaxies with collective performance as good as professional astronomers, while in reCaptcha\footnote{\url{http://www.google.com/recaptcha}} they refine the data that describes digital manuscripts.
-Sometimes, systems look for innovative ways of getting people to engage with the systems, including game narratives, like in EteRNA\footnote{\url{http://eterna.cmu.edu}}, in which  participants design RNA  that is evaluated automatically using simulators and then refined via human effort, or monetary rewards, like in Amazon's Mechanical Turk, a platform in which users undertake routine (micro) tasks such as object classification, audio transcription and text translation.
-
-  \item [The roles of the participants and the social network they form]:
-
-  \item [The motivating factors and incentive mechanisms]: system features influencing user behavior and the social network dynamics
-
-  \item [Interaction between social and machine component]: how are human contributions used to compute things, including quality control, spam, workflow, aggregation of results
-\end{description}
-
 \section{Comparison with related fields}
@@
 
-\subsection{Methodology}
+\subsection{Methodology} \label{meth}
 
@@
 
-\subsection{Classification dimensions}
+\subsection{The Polyarchical Relationship of Social Machines}
 
-\subsubsection{The polyarchical relationship of social machines}
 While determining a set of social machines, one clear distinction was between social machine frameworks, such as MediaWiki and Ushahidi, which enable
@@
 
-\begin{figure*}[htb]
+\begin{figure}[htb]
 \begin{center}
-\includegraphics[width=18cm]{img/polyarchy.pdf}
+\includegraphics[width=8.5cm]{img/polyarchy.pdf}
 \caption{A polyarchy of Social Machines, illustrating the infrastructure and frameworks used by social machines, and machine-machine usage.} \label{polyarchy}
 \end{center}
-\end{figure*}
+\end{figure}
 
@@
 
-\subsubsection{Multi-faceted motivation, and who-gets-what benefit}
-
-%We identified a hierarchy of constructs pertaining to the motivation people have for using social
-%machines with top level constructs of:
+%\subsubsection{Multi-faceted motivation, and who-gets-what benefit}
+%
+%%We identified a hierarchy of constructs pertaining to the motivation people have for using social
+%%machines with top level constructs of:
+%%
+%%\begin{enumerate}
+%%\item Hedonistic
+%%\item Financial
+%%\item To be informed
+%%\item To help others
+%%\end{enumerate}
+%%
+%%The sub constructs are illustrated in Figure~\ref{motivation}, and we relate them to Maslow's hierarchy of needs~\cite{maslow}, (e.g., self-actualisation, security, respect of others), in various ways.
+%%
+%%\begin{figure*}[htb]
+%%\begin{center}
+%%\includegraphics[width=18cm]{img/motivation.pdf}
+%%\caption{Motivation hierarchy of participants using Social Machines, related to Maslow's hierarchy of needs.} \label{motivation}
+%%\end{center}
+%%\end{figure*}
+%
+%The motivations of users can be identified and mapped to existing human needs and motivations
+%systems, such as Maslow's hierarchy of human needs~\cite{maslow} (e.g., self-actualisation, security
+%respect of others). For each of these motivations, different roles that participate may have different {\it core motivations}, such as:
 %
 %\begin{enumerate}
-%\item Hedonistic
-%\item Financial
-%\item To be informed
-%\item To help others
+%\item Benefit to contributor
+%\item Benefit to moderator
+%\item Benefit to system operator/host (e.g, Google, Facebook, Amazon)
+%\item To affiliates of the host (e.g., Amazon affiliates, YouTube partners)
+%\item To society as a whole
+%\item To a contributor's social network
 %\end{enumerate}
 %
-%The sub constructs are illustrated in Figure~\ref{motivation}, and we relate them to Maslow's hierarchy of needs~\cite{maslow}, (e.g., self-actualisation, security, respect of others), in various ways.
+%In addition to relationships to Maslow's hierarchy of needs, which drive the motivations,
+%we suggest that there are also resulting actions from each motivation. Thus, it should be
+%possible to follow basic human needs to motivations in social machines, through to the
+%actual actions (and therefore the necessary software feature support) that are made by
+%participants in social machines. Our theory is that this path may enable designers of
+%social machines to satisfy human needs by implemented specific features into their social
+%machines, or to identify potential gaps in their serving of needs and motivations due
+%to lack of implemented features.
 %
-%\begin{figure*}[htb]
-%\begin{center}
-%\includegraphics[width=18cm]{img/motivation.pdf}
-%\caption{Motivation hierarchy of participants using Social Machines, related to Maslow's hierarchy of needs.} \label{motivation}
-%\end{center}
-%\end{figure*}
+%Motivation in specific social machines has been studied, including motivation of participants
+%in: Wikipedia~\cite{kuznetsov2006motivations}; Open Source Software~\cite{lakhani2003hackers};
+%Peer-to-peer filesharing~\cite{p2p}; Tagging~\cite{tagging}; and Virtual
+%Communities~\cite{ardichvili2003motivation,Moore:2007:UMM:1235000.1235035}.
+%
+%[relate this work back to ours briefly.]
 
-The motivations of users can be identified and mapped to existing human needs and motivations
-systems, such as Maslow's hierarchy of human needs~\cite{maslow} (e.g., self-actualisation, security
-respect of others). For each of these motivations, different roles that participate may have different {\it core motivations}, such as:
+%\subsubsection{Artefacts of Social Machines}
+%
+%Our next observation was that there are a number of shared artefact types that can be classified within the context of social machines. Specifically that
+%there are ``intentions'' that motivate the use of social machines, ``actions'' that are directly performed by participants of social machines, and ``contributions``
+%that result from the use of social machines. We have taken these abstract concepts, and mapped them to a number of well-known social machines in order
+%to demonstrate that these concepts can be used to compare and contrast different machines.
+%
+%[more here]
 
-\begin{enumerate}
-\item Benefit to contributor
-\item Benefit to moderator
-\item Benefit to system operator/host (e.g, Google, Facebook, Amazon)
-\item To affiliates of the host (e.g., Amazon affiliates, YouTube partners)
-\item To society as a whole
-\item To a contributor's social network
-\end{enumerate}
-
-In addition to relationships to Maslow's hierarchy of needs, which drive the motivations,
-we suggest that there are also resulting actions from each motivation. Thus, it should be
-possible to follow basic human needs to motivations in social machines, through to the
-actual actions (and therefore the necessary software feature support) that are made by
-participants in social machines. Our theory is that this path may enable designers of
-social machines to satisfy human needs by implemented specific features into their social
-machines, or to identify potential gaps in their serving of needs and motivations due
-to lack of implemented features.
-
-Motivation in specific social machines has been studied, including motivation of participants
-in: Wikipedia~\cite{kuznetsov2006motivations}; Open Source Software~\cite{lakhani2003hackers};
-Peer-to-peer filesharing~\cite{p2p}; Tagging~\cite{tagging}; and Virtual
-Communities~\cite{ardichvili2003motivation,Moore:2007:UMM:1235000.1235035}.
-
-[relate this work back to ours briefly.]
-
-\subsubsection{Artefacts of Social Machines}
-
-Our next observation was that there are a number of shared artefact types that can be classified within the context of social machines. Specifically that
-there are ``intentions'' that motivate the use of social machines, ``actions'' that are directly performed by participants of social machines, and ``contributions``
-that result from the use of social machines. We have taken these abstract concepts, and mapped them to a number of well-known social machines in order
-to demonstrate that these concepts can be used to compare and contrast different machines.
-
-[more here]
-
-\subsubsection{Metrics of participants and usage}
-
-Finally, we identified several constructs that are relevant for all social machines.
+%\subsubsection{Metrics of participants and usage}
+%
+%Finally, we identified several constructs that are relevant for all social machines.
 First, there are construct that relate to ``analytics'' that can be measured at a point in time, and may change over time.
@@
 
-\begin{table*}[htbp]
+\subsection{Constructs}
+
+The four clusters of constructs generated by the process described in Section~\ref{meth}
+are shown in full in Table \ref{table:constructs}, and are detailed below.
+
+\begin{table}[htb]
 \begin{center}
-\begin{tabular}{|p{18cm}|}
+\begin{tabular}{|p{8cm}|}
 \hline
-{\bf Designed Affordances} \\
+{\bf Tasks} \\
 \hline
-Domain specificity [1 - Specific / 5 - General] \\
-Visibility and importance of a participants `reputation' [1 - Not important / 5- Important] \\
-Explicit representation of participant reputation [1 - No Reputations / 5 - Reputations] \\
-Participants can define new types of contributions [1 - Not possible / 5 - Complete freedom] \\
-Degree of open source software used/created [1 - Proprietary / 5 - Entirely open source] \\
-Degree of openness and availability of data created in the system (APIs/Dumps) [1 - Closed/unavailable data / 5 - Completely open data] \\
-Variety of actions participants can perform [1 - Single Action / 5 - Large variety of actions] \\
+Variety of types of contributions \\
+Domain specificity \\
+Participation is done to `get something done' \\
+Participation is done `for fun' \\
+Participation is done to exchange knowledge \\
+Participation is done to be social \\
+Participants are intrinsically motivated \\
+Participants are motivated by extrinsic reward \\
 \hline
-{\bf Participatory Affordances} \\
+{\bf Users and Roles} \\
 \hline
-Supports social interaction [1 - No social features / 5 - Large variety of social features] \\
-How often a participant participates (times a day, etc) [1 - Many times a day / 5 Seldom] \\
-Importance of Timely Participation [1 - Unimportant / 5 - Timely] \\
-Generality of Audience [1 - Niche / 5 - General] \\
-Participant anonymity [1 - No Anonymity / 5 - Complete Anonymity] \\
-Participant autonomy (1 - No autonomy / 5 - Complete autonomy) \\
-Quality of participant contributions [1 - Low quality / 5 - High quality] \\
-Clear separation of roles/responsibilities among participants - [1 - Single role - Everyone is the same) / 5 - People assume different roles/responsibilities] \\
-Extent of hierarchical organisation of roles [1 - No Hierarchy / 5 - Deep/Well-defined Hierarchy] \\
-Variety of types of contributions [1 - Single Type / 5 - Large variety of types] \\
-Participants are motivated by extrinsic award  [1 - Disagree / 5 - Agree] \\
-Participants are intrinsically motivated [1 - Disagree / 5 - Agree] \\
-Service/Platform derives benefit from participant participation [1 - Disagree /  5- Agree] \\
-Participants directly benefit from participating [1 - Disagree / 5 - Agree] \\
-Participants' friends/social network benefit from participation [1 - Disagree / 5 - Agree] \\
-Participation is for the benefit of a specific person/group other than the participant or platform owner [1-Disagree / 5-Agree] \\
-Participation is for the benefit of society/the world at large  [1 - Disagree / 5 - Agree] \\
-Participation is done to `get something done' [1-Disagree/5-Agree] \\
-Participation is done `for fun' [1-Disagree/5-Agree] \\
-Participation is done to exchange knowledge [1-Disagree/5-Agree] \\
-Participation is done to be social [1-Disagree/5-Agree] \\
-Participation is done via mobile devices [1-Never / 5 - Often] \\
-Participants' (geographic) location is used by the service [1-No / 5 - Very] \\
+Generality of Audience \\
+Participant autonomy \\
+Participant anonymity \\
+Extent of hierarchical organisation of roles \\
+Clear separation of roles/responsibilities among participants \\
 \hline
-{\bf Analytics} \\
+{\bf Participation} \\
 \hline
-Popularity [1 - Unpopular / 5 - Very Popular] \\
-Maturity [1 - New / 5 - Mature] \\
-Ratio of passive to active participants [1 - Passive / 5 - Active] \\
-Number of competitors [1 - Not Many / 5 - Many] \\
-Extent of global geographic coverage [1 - Highly localised / 5 - Global]\\
+Participants' (geographic) location is used by the service \\
+Participation is done via: \\
+\hspace{1cm} mobile devices \\
+\hspace{1cm} web browsers \\
+\hspace{1cm} apps \\
+\hspace{1cm} sensors/sensing (location sensing and wearable devices) \\
+Participation is done in a mobile context \\
+\hline
+{\bf Computation, Output and Benefit} \\
+\hline
+Output solves a computational problem \\
+Output requires human insight \\
+Output is a subjective appraisal \\
+Participants directly benefit from participating \\
+Service/Platform derives benefit from participant participation \\
+Participants'' friends/social network benefit from participation \\
+Participation is for the benefit of a specific person/group other than the participant or platform owner \\
+Participation is for the benefit of society/the world at large  \\
 \hline
@@
 \end{center}
-\caption{Constructs of social machines, listing affordances that are designed into machines, affordances that have resulted from use, and analytical constructs that can be measured and change over time.} \label{table:constructs}
-\end{table*}%
+\caption{Consolidated constructs of social machines.} \label{table:constructs}
+\end{table}
 
-The result is what we call `participation' constructs. These constructs arise from use of a
-social machine, and are not necessarily predictable; they include:
+\subsubsection{Tasks}
 
-\begin{enumerate}
-\item {\bf Roles emerging in the social machine}
-   \newline In addition to designed and implemented roles such as {\it moderator} and {\it administrator}, emerging roles include motivations of the participant, such as {\it trolls}, {\it spammers}, {\it domain experts}, {\it social nexus}, and {\it re-blogger}.
-\item {\bf Quality of contributions}
-    \newline For any social machine, the quality of contributions will vary. While some
-    issues with quality can be predicted and protected against (e.g., through
-    multi-user result agreement on Mechanical Turk~\cite{ipeirotis2010quality}), some
-    particular issues with quality cannot always be predicted ahead of time.
-\item {\bf Content of contributions}
-    \newline While a system operator may design their social machine to generate specific
-    types of content, participants may subvert the system in order to generate other types,
-    or to concentrate on areas that were unforeseen at the time of launch, but are now seen
-    as useful to the operators. It is also possible that multiple social machines can be
-    combined/linked by users to further expand the types of contributions they produce.
-\item {\bf Global / cultural reach}
-    \newline A social machine may naturally focus on a narrow geographic area, for example
-    an Ushahidi-based social machine for local elections~\cite{meier2008crisis} or
-    natural disasters~\cite{morrow2011independent}. Such machines are unlikely to be
-    re-purposed by participants outside of that area. However, some machines have been
-    created for a single cultural/global area, but have potential global relevance that was
-    unforeseen early on. An example is the fall-off of English-language users of the
-    social network ``Orkut,'' which gained significant usage elsewhere, particularly in
-    Brazil.
-\item {\bf Task reach --- what do people use it for?}
-    \newline In addition to breaking cultural barriers, user have also shown that they can
-    use social machines for tasks that were not set up by the system operators. For example
-    the identification of ``Green Pea Galaxies'' using the GalaxyZoo social machine for
-    astronomy~\cite{greenpea}. Users were not originally asked to identify these types, but
-    a number of users identified them, and the software was modified to include them,
-    resulting in their discovery as a new type of galaxy.
-\item {\bf Active/Passive participation roles}
-    \newline When it comes to participation in social machines, there are levels of
-    participation that can be classified as being ``active'' and being ``passive.'' These
-    roles may not be particularly designed into systems, but may be exhibited over time.
-    Particularly noteworthy examples can be seen in news and link aggregation social
-    machines such as Reddit and Digg, where users can both submit links and vote on links.
-    The frequency of each of these activities can vary wildly, with a small
-    minority of users controlling the links that rise to the top~\cite{digg}.
-\end{enumerate}
+{\bf The types of tasks undertaken by the social component and the resulting contributions:}Social machines, just as any computing system, are built with a given purpose in mind, and exhibit functionality by which this purpose can be achieved.
+Users and the emerging social network interact with the system and with each other to carry out specific tasks by which they contribute to the 'computation'.
+This includes the creation and refinement of digital artifacts by executing given actions, possibly in collaboration with other participants.
+For example, in GalaxyZoo\footnote{\url{http://www.galaxyzoo.org}} people classify galaxies with collective performance as good as professional astronomers, while in reCaptcha\footnote{\url{http://www.google.com/recaptcha}} they refine the data that describes digital manuscripts.
+Sometimes, systems look for innovative ways of getting people to engage with the systems, including game narratives, like in EteRNA\footnote{\url{http://eterna.cmu.edu}}, in which  participants design RNA  that is evaluated automatically using simulators and then refined via human effort, or monetary rewards, like in Amazon's Mechanical Turk, a platform in which users undertake routine (micro) tasks such as object classification, audio transcription and text translation.
+
+\subsubsection{Users and Roles}
+
+{\bf The roles of the participants and the social network they form}
+
+\subsubsection{Participation (How users interact with the Social Machine)}
+
+\subsubsection{Computation, Output and Benefit}
+
+%The result is what we call `participation' constructs. These constructs arise from use of a
+%social machine, and are not necessarily predictable; they include:
+%
+%\begin{enumerate}
+%\item {\bf Roles emerging in the social machine}
+%   \newline In addition to designed and implemented roles such as {\it moderator} and {\it administrator}, emerging roles include motivations of the participant, such as {\it trolls}, {\it spammers}, {\it domain experts}, {\it social nexus}, and {\it re-blogger}.
+%\item {\bf Quality of contributions}
+%    \newline For any social machine, the quality of contributions will vary. While some
+%    issues with quality can be predicted and protected against (e.g., through
+%    multi-user result agreement on Mechanical Turk~\cite{ipeirotis2010quality}), some
+%    particular issues with quality cannot always be predicted ahead of time.
+%\item {\bf Content of contributions}
+%    \newline While a system operator may design their social machine to generate specific
+%    types of content, participants may subvert the system in order to generate other types,
+%    or to concentrate on areas that were unforeseen at the time of launch, but are now seen
+%    as useful to the operators. It is also possible that multiple social machines can be
+%    combined/linked by users to further expand the types of contributions they produce.
+%\item {\bf Global / cultural reach}
+%    \newline A social machine may naturally focus on a narrow geographic area, for example
+%    an Ushahidi-based social machine for local elections~\cite{meier2008crisis} or
+%    natural disasters~\cite{morrow2011independent}. Such machines are unlikely to be
+%    re-purposed by participants outside of that area. However, some machines have been
+%    created for a single cultural/global area, but have potential global relevance that was
+%    unforeseen early on. An example is the fall-off of English-language users of the
+%    social network ``Orkut,'' which gained significant usage elsewhere, particularly in
+%    Brazil.
+%\item {\bf Task reach --- what do people use it for?}
+%    \newline In addition to breaking cultural barriers, user have also shown that they can
+%    use social machines for tasks that were not set up by the system operators. For example
+%    the identification of ``Green Pea Galaxies'' using the GalaxyZoo social machine for
+%    astronomy~\cite{greenpea}. Users were not originally asked to identify these types, but
+%    a number of users identified them, and the software was modified to include them,
+%    resulting in their discovery as a new type of galaxy.
+%\item {\bf Active/Passive participation roles}
+%    \newline When it comes to participation in social machines, there are levels of
+%    participation that can be classified as being ``active'' and being ``passive.'' These
+%    roles may not be particularly designed into systems, but may be exhibited over time.
+%    Particularly noteworthy examples can be seen in news and link aggregation social
+%    machines such as Reddit and Digg, where users can both submit links and vote on links.
+%    The frequency of each of these activities can vary wildly, with a small
+%    minority of users controlling the links that rise to the top~\cite{digg}.
+%\end{enumerate}
 
@@
 
-\begin{figure}
+\begin{figure}[htb]
 \begin{center}

### 2013-02-24 latex/paper_new.tex
working on 4.3

@@
 
+%% Constructs =============================================================
+
 \subsection{Constructs}
 
-The four clusters of constructs generated by the process described in Section~\ref{meth}
-are shown in full in Table \ref{table:constructs}, and are detailed below.
+The clustering process described in Section~\ref{meth} yielded both a
+reduction in the overall number of constructs (in particular an
+elimination of overrepresented areas) but also identificaiton of four
+simple themes: sets of constructs pertaining to users and roles within
+each machine, second, the nature of tasks, problems and activities
+performed through and using the machine, third, motivation for
+participation, finally, the contexts of participation.  Each of these
+clusters contain 5-8 constructs as illustrated in Table
+\ref{table:constructs}; we describe each of these themes in the
+following subsections.
 
@@
 \begin{tabular}{|p{8cm}|}
-\hline
-{\bf Tasks} \\
-\hline
-Variety of types of contributions \\
-Domain specificity \\
-Participation is done to `get something done' \\
-Participation is done `for fun' \\
-Participation is done to exchange knowledge \\
-Participation is done to be social \\
-Participants are intrinsically motivated \\
-Participants are motivated by extrinsic reward \\
 \hline
@@
 \hline
-{\bf Participation} \\
+{\bf Tasks} \\
 \hline
-Participants' (geographic) location is used by the service \\
+Task involves subjective appraisal of content \\
+Task solves (a defnable) computation problem/set of problems\\
+Domain-specificity of tasks \\
+Service/platform derives benefit from participation \\
+Variation in types of contributions and tasks \\
+{\bf Motivation} \\
+Participants are intrinsically motivated:\\
+\hspace{1cm} to gain/share knowledge \\
+\hspace{1cm} to ``get something done'' \\
+\hspace{1cm} to ``for fun/entertainment''\\
+\hspace{1cm} to ``be social'' \\
+Participation is motivated by extrinsic reward (payment, status etc) \\
+Participation is done for the benefit of a specific third-party \\
+Participation is done for the benefit of society as a whole \\
+\hline
+{\bf Context of Participation} \\
+\hline
+Participants' (geographic) location is relevant to the service \\
 Participation is done via: \\
@@
 \hline
-{\bf Computation, Output and Benefit} \\
-\hline
-Output solves a computational problem \\
-Output requires human insight \\
-Output is a subjective appraisal \\
-Participants directly benefit from participating \\
-Service/Platform derives benefit from participant participation \\
-Participants'' friends/social network benefit from participation \\
-Participation is for the benefit of a specific person/group other than the participant or platform owner \\
-Participation is for the benefit of society/the world at large  \\
-\hline
 \end{tabular}
@@
 
-\subsubsection{Tasks}
+\subsubsection{Users and Roles}
+The first set of constructs pertain to the human participants of
+social machines and the way(s) participants are organised within them.
+First, whether participants consistute members of the general public,
+or are of a specific demongraphic, external group or organisation,
+occupation, or of a particular expertise.  Second, pertain to how
+constrained interaction is among members of the machine (lack of
+\emph{autonomy}), and the degree to which participants are anonymous.
+
+\subsubsection{Tasks, Problems and Activiites}}
+
+The first set of constructs pertain to \emph{what} human components
+of the social machine do within the system.  These constructs pertain
+to the
 
@@
 
-{\bf The roles of the participants and the social network they form}
+% {\bf The roles of the participants and the social network they form}
 
-\subsubsection{Participation (How users interact with the Social Machine)}
+% \subsubsection{Participation (How users interact with the Social Machine)}
 
-\subsubsection{Computation, Output and Benefit}
+\subsubsection{Participation - \emph{How and when user participate}}
+
+\subsubsection{Problems Solved and Benefit Derived}
 

### 2013-02-24 latex/paper_new.tex
working on 4.3

@@
 social machines and the way(s) participants are organised within them.
-First, whether participants consistute members of the general public,
-or are of a specific demongraphic, external group or organisation,
-occupation, or of a particular expertise.  Second, pertain to how
-constrained interaction is among members of the machine (lack of
-\emph{autonomy}), and the degree to which participants are anonymous.
+One construct identifies whether participants constitute members of
+the general public, or are of a specific demongraphic, external group
+or organisation, occupation, or of a particular expertise.  Two
+constructs, \emph{autonomy} and \emph{anonymity} pertain to the degree
+to which interaction is constrained among participants, and to the
+degree to which participants' identities are used and disclosed within
+the machine. The final two constructs pertain to the roles: first,
+as to whether multiple roles are pre-differentiated by the machine or
+whether all participants essentially initially assume the same role,
+second whether roles (pre-defined or emergent) are archical.
 

### 2013-02-24 latex/paper_new.tex
working on 4.3

@@
 \hline
+Task involves creative production or creation of content \\
 Task involves subjective appraisal of content \\
@@
 
-\subsubsection{Tasks, Problems and Activiites}}
+\subsubsection{Tasks, Problems and Activiites}
+The first set of constructs pertain to \emph{what} the participants do
+individually and collectively within the system.  This task could
+involve creation of content, subjective appraisal of content, posing
+of problems, group solving of problems, and so on.  Tasks relate to a
+pre-defined problem space or domain, or may be support a general,
+broad collection of activities.  Finally, a question of whether a
+direct benefit is derived by the platform/machine provider other than
+the benefits gained by participants through participation.
 
-The first set of constructs pertain to \emph{what} human components
-of the social machine do within the system.  These constructs pertain
-to the
-
-{\bf The types of tasks undertaken by the social component and the resulting contributions:}Social machines, just as any computing system, are built with a given purpose in mind, and exhibit functionality by which this purpose can be achieved.
+%% {\bf The types of tasks undertaken by the social component and the resulting contributions:}Social machines, just as any computing system, are built with a given purpose in mind, and exhibit functionality by which this purpose can be achieved.
 Users and the emerging social network interact with the system and with each other to carry out specific tasks by which they contribute to the 'computation'.

### 2013-02-24 latex/paper_new.tex
working on 4.3

@@
 Variation in types of contributions and tasks \\
+\hline
 {\bf Motivation} \\
+\hline
 Participants are intrinsically motivated:\\

### 2013-02-24 latex/paper_new.tex
working on 4.3

@@
 \hline
-Task involves creative production or creation of content \\
-Task involves subjective appraisal of content \\
-Task solves (a defnable) computation problem/set of problems\\
-Domain-specificity of tasks \\
-Service/platform derives benefit from participation \\
+Activities involve creative production of content \\
+Activities involve subjective appraisal of content \\
+Activities involve solving (a definable) computation task or set of problems\\
+Tasks are domain-specific \\
+The machine owner (service/platform) derives benefit from participation \\
+Activities and tasks are pre-defined or participant defined \\
 Variation in types of contributions and tasks \\

### 2013-02-24 latex/paper_new.tex
working on 4.3

--- 2013-02-24 latex/paper_new.tex
@@
 \hline
-Task involves creative production or creation of content \\
-Task involves subjective appraisal of content \\
-Task solves (a defnable) computation problem/set of problems\\
-Domain-specificity of tasks \\
-Service/platform derives benefit from participation \\
+Activities involve creative production of content \\
+Activities involve subjective appraisal of content \\
+Activities involve solving (a definable) computation task or set of problems\\
+Tasks are domain-specific \\
+The machine owner (service/platform) derives benefit from participation \\
+Activities and tasks are pre-defined or participant-defined \\
 Variation in types of contributions and tasks \\
@@
 \hspace{1cm} to ``be social'' \\
+\hspace{1cm} for the benefit of a specific group of people who need help \\ 
+\hspace{1cm} for the benefit of society as a whole \\ 
 Participation is motivated by extrinsic reward (payment, status etc) \\
-Participation is done for the benefit of a specific third-party \\
-Participation is done for the benefit of society as a whole \\
 \hline
@@
 \hline
-Participants' (geographic) location is relevant to the service \\
+Participants' participate to fulfill needs of a role  \\
+\hspace{1cm} work-related/professional role
+\hspace{1cm} home/family related
+\hspace{1cm} social role
+\hspace{1cm} leisure/entertainment role
+Participants' physical location is relevant to the service \\
 Participation is done via: \\
@@
 \subsubsection{Tasks, Problems and Activiites}
-The first set of constructs pertain to \emph{what} the participants do
-individually and collectively within the system.  This task could
-involve creation of content, subjective appraisal of content, posing
-of problems, group solving of problems, and so on.  Tasks relate to a
-pre-defined problem space or domain, or may be support a general,
-broad collection of activities.  Finally, a question of whether a
-direct benefit is derived by the platform/machine provider other than
-the benefits gained by participants through participation.
+The first set of constructs pertain to \emph{what} the participants
+do, both individually within the system, and collectively as a whole.
+Individual activities might include creation of content, subjective
+appraisal of existing content, posing of problems, solving of
+problems, andf so on.  The resulting computation, looked at from a
+macroscopic whole, might be the identification of high-quality
+artistic content, insight, or creative works, collective problem
+solving, and so on.  Such tasks might be pre-defined by the system
+designers, or brought to the system by participants.  Futher, these
+tasks might be of a specific type or class, or of many varied types or
+encompass a general class of activities.  To capture these
+difeferences, we include a construct that differentiates machines with
+pre-defined problem spaces or are open to participant-provided tasks
+and activities.  Finally, we ask whether the benefit derived by
+participants is distinct from that gained by machine/platform owners
+and service providers.
+
+\subsubsection{Motivation}
+The third set of themes pertains to the motivation that sustains
+continued participation in these systems.  We identiy 6 different
+kinds of common sources of intrinsic motivation; first, that
+participation is fun, second that it accomplishes an activity that the
+participant enjoys or wanted to get done, third, that it satisfies the
+desire to gain or share knowledge, or, fourth, or to be social.  The
+last two pertain to philanthropy, whether participation is seen to
+benefit a particular group of disadvantaged people or individuals who
+need assistance, or, finally, if participation is beneficial to
+society as a whole.  Finally, we ask whether extrinsic motivational
+factors also contributed to sustained participation - such factors
+could include money reward, status, recognition and so on
+
+\subsection{Context of Participation}
+The final set of constructs pertains to the context(s) of each
+participant's greater (life-) activities that participation is
+performed.  This context could be in a work context, home/family
+context, for leisure, and at a desk or away from a computer (in a
+mobile context).  PArticipation could be conducted via apps, web
+browsers, sensors and other technolgoies.  Due to the predominance of
+``geo-locative'' services that incorporate information about the
+user's location, we have included a construct identifying the degree
+to which the user's location is used by the machine.
 

### 2013-02-24 latex/paper_new.tex
written sec

@@
 
-\subsection{Context of Participation}
+\subsubsection{Context of Participation}
 The final set of constructs pertains to the context(s) of each
@@
 
-\subsubsection{Users and Roles}
-
+% \subsubsection{Users and Roles}
 % {\bf The roles of the participants and the social network they form}
-
 % \subsubsection{Participation (How users interact with the Social Machine)}
-
-\subsubsection{Participation - \emph{How and when user participate}}
-
-\subsubsection{Problems Solved and Benefit Derived}
-
+%\subsubsection{Participation - \emph{How and when user participate}}
+%\subsubsection{Problems Solved and Benefit Derived}
 %The result is what we call `participation' constructs. These constructs arise from use of a

### 2013-02-26 latex/paper_new.tex
written sec
intermediary version

@@
 The success of the Web has changed this relationship: we now see configurations of people interacting with content and with each other, typified by social Web sites.
-Rather than drawing a line through such Web-based systems to separate the human and machine parts (as Computer Science has traditionally done), we can now draw a line around them  and  treat  each  such  assembly  as  a  'social  machine'  -  a  machine  in which  digital  and  human components are seamlessly interwoven.
+Rather than drawing a line through such Web-based systems to separate the human and digital parts (as computer science has traditionally done), we can now draw a line around them  and  treat  each  such  compound  as  a  'social  machine'  -  a  machine in which the two aspects are seamlessly interwoven.
 This crucial transition in thinking acknowledges the reality of current socio-technical systems, and is essential to underpin any understanding of the science and engineering of their future development towards pervasive ecosystems of co-evolving social machines.
 
+Essentially social machines can be characterized as assemblies of manually executed and machine-driven (as in 'automatized') services and the interaction of such services.
+A  traditional  database,  with  users  looking  up  records  independently  of  each  other  and  an  administrator responsible for the management of the content, has some of the right ingredients, but there is really no social element in this strict provider-consumer relationship.
+When we fill out a form (e.g., health information, birds spotted in my garden, new construction sites on my way to work) the systems supporting this activity are minimal forms of social machines  because the users  are  part  of  the  'social  computation', in this case the data creation and collection process facilitate through the site.
+This social component  becomes  richer when  the database  is  curated by members  of  the broader community (e.g., Wikipedia) and when  the social network adds value implicitly (Amazon) or explicitly (Facebook) to the overall system through their individual or joint activities.
+
 A number of prominent examples aside, today we still see a divide between conventional IT systems dedicated to data- and computation-intensive tasks, and Web $2.0$ sites offering some combination of well-known participatory features, in which user-generated content and the underlying social network evolve dynamically and hand-in-hand.
-As technology becomes more and more ubiquitous, many of the challenges we will witness in any domain of our life, business, and society will need solutions that rely on both of these axes: a sophisticated  combination  of  data-intensive, complex automation and deep community involvement.
-This suggests the need for new types of systems to tackle these challenges, and to achieve requires a through understanding of the science and engineering of social machines.
+However, as technology becomes more and more ubiquitous, many of the challenges we witness in any domain of our life, business, and society will soon require solutions that rely on both of these axes: a sophisticated  combination  of  data-intensive, complex automation and deep community involvement.
+This suggests the need for new types of systems to tackle these emerging challenges, and these systems will not be able to be built and used in a sustainable way without a through understanding of the science and engineering of (the continuum of) social machines.
 
-As a first step, we propose a classification framework for social machines, which can be used to describe, study, and compare this expanding field of interdisciplinary research and the increasing number of systems combining human and computational intelligence to serve purposes as diverse as traffic management, healthcare, media and journalism, or design and natural sciences.
+As an early step towards achieving this principled understanding, we propose in this paper an initial outline for a classification framework for social machines.
+The primary aim of the framework is to identify and define the constructs to describe, study, and compare this expanding field of interdisciplinary research, as well as the increasing number of systems combining human and computational intelligence that are continuously being created in areas as diverse as traffic management, healthcare, media and journalism, or design and natural sciences.
 
-In this paper we introduce the current version of the framework, which is the result of a knowledge elicitation process performed internally in our lab.
-Our aim is to present it to the broader community and extend and revise it based on their feedback in order to achieve a shared understanding, terminology, and set of descriptive features to enable a systematic study of existing and future social machines.
-Using a common descriptive classification allows researchers to observe the effects of specific technical properties of a system on social behavior, identify design and evolution patterns, learn about social network formation and dynamics, and devise incentive mechanisms to encourage large-scale participation.
-A second audience group targeted by our work are developers of social machines.
-For them, the framework may provide a theoretical grounding for the realization of new systems, informing system design and operation in terms of critical features and predicted user behavior and community development.
+The current version of our framework is the result of a knowledge elicitation process performed internally in our lab.
+Our goal is to present it to the broader community and extend and revise it based on their feedback in order to achieve a shared understanding of the basic notions and of the associated terminology, and to consensually lay out the structure of the design space that will enable a systematic study of existing and future social machines.
+We anticipate that the framework will be a useful tool for both researchers in social and computer sciences, and for developers and operators of social machines.
+Using a common conceptualization allows the former to become familiar with the landscape, and identify topics of research that so far have remained unexplored or under-explored.
+They are given a theoretical grounding to observe the effects of specific technical properties of a system configuration on social behavior, discover design and evolution patterns of such systems, learn about social network formation and dynamics, and devise incentive mechanisms to encourage wide participation.
+For developers and operators, the framework may inform the engineering of new systems in terms of critical features and community development.
 
-\section{What are social machines?}
-Social machines can be basically characterized as assemblies of human and automatic components and their interaction.
-A  traditional  database,  with  users  looking  up  records  independently  of  each  other  and  an  administrator responsible for content, has some of the right ingredients but there is really no social component in this strict provider-consumer relationship.
-When we fill out a form (e.g., health information, birds spotted in my garden) it  is minimally  a  social machine  because humans  are  part  of  the  'social  computation'.
-This  becomes  richer when  the database  is  curated by members  of  the  community  (e.g., Wikipedia) and when  the  social network adds value implicitly (Amazon) or explicitly (Facebook).
-We anticipate many categories of social machine in this continuum, and they may manifest themselves as Web site or app, or indeed in the Internet of Things.
+This paper makes two main contributions:
+
+\begin{itemize}
+  \item We introduce the notion of 'social machines' and position it in connection to similar ideas and initiatives.
+  \item We propose a set of constructs to capture the most important features of social machines and to compare existing instances thereof.
+\end{itemize}
+
+The remainder of this paper is organized as follows.
+In order to pin down what makes up a social machine, and establish the boundaries of this emerging field of research, we need to first understand the relationship between social machines and related topics such as human computation, collective intelligence, and crowdsourcing; Section \ref{sec:comparison} gives a summary of this comparative analysis.
+In Section \ref{sec:framework} we then present the framework in terms of the methodology used to devise it, the main clusters of constructs used for classification, and an example of using them to compare a selection of popular social machines via grid analysis.
+We conclude with an outlook of our future development activities, including details about the evaluation of the framework.
+In addition, we plan to engage with the broader community to set up a 'social machine' bringing together researchers from various disciplines who will contribute to the 'computation' of a joint and widely agreed classification framework for the field.
 
 \section{Comparison with related fields}
+\label{sec:comparison}
+Since the invention of the Turing machine we have experienced a 'paradigm shift' in the usage of computers, moving away from purely calculative devices to facilitators of a wide range of human interactions.
+The emergence of 'Computer-supported Collaboration Work (CSCW)' \cite{grudin1994computer} is representative for the early days of this trend.
+CSCW addresses 'how collaborative activities and their coordination can be supported by means of computer systems' \cite{grudin1994computer}, whereas the initial concept evolved towards the more broader field of 'Computer-supported collaboration' (CSC).
 
-A Social Machine includes human and machine two major components, therefore the study of social machine is an interdisciplinary endeavour.
-Since the Turning machine was invented, there has been a `paradigm shift' \cite{kuhn1996structure} in using computers, where the computer facilitates human communication rather acts as a purely calculative device.
-The Computer-supported Collaboration Work (CSCW) \cite{grudin1994computer} emerged in 1984 represents such shift.
-It addresses ``how collaborative activities and their coordination can be supported by means of computer systems''.
-This concept later evolved into a broader sense of  Computer-supported collaboration (CSC).
-The Social computing \cite{parameswaran2007research} is another shift, which refers systems that support ``the gathering, representation, processing, use, and dissemination of information that is distributed across social collectivities such as teams, communities, organizations, and markets''.
-To compare with the CSC, the social computing has more focus in `computing' rather than `collaboration'.
-Similarly, the Human Computation \cite{von2009human} is another paradigm towards the computational direction, which utilize human processing power to solve problems that computers cannot yet solve.
-All these three mentioned fields overlap in their concepts and are mainly popularised in the computer science.
-In the Web 2.0 era, there were few generic concepts/buzzword emerged on the idea of utilising the human power which are not only used in computer science, but also in other fields such as social science, business etc.
-Crowding sourcing \cite{howe2006crowdsourcing} as an example, describes the idea that taking a job traditionally performed by a designated agent and outsourcing it to an undefined, generally large group of people in the form of an open call.
-Open Innovation \cite{chesbrough2008open} advocates the similar idea with empathizing on the innovation in firms.
-Surowiecki also argued in the book of the wisdom of crowds \cite{surowiecki2005wisdom} that the aggregation of information in groups, resulting in decisions that are often better than could have been made by any single member of group.
-Collective Intelligence described by Malone \cite{malone2009harnessing}, defines an even more broadly concept that ``groups of individuals doing things collectively that seem intelligent.''
-The concept overlapping between human computation, crowdsourcing, social computing and collective intelligence is discussed in \cite{quinn2011human}.
-In  social machines  the  social and  the  technical are equal  partners.
-While  crowdsourcing  could  be  viewed  as  'human  subroutines'  and  'citizen  sensors',  the social machine perspective is one of incentivised participation and not humans enslaved to computers.
-We define the scope of social machine in a venn digram in Figure \ref{socialmachine}.
-
-The study of social machines is naturally an interdisciplinary endeavor.
-Since the Turning machine was invented, we have experienced a paradigm shift in computer science from computers as purely calculative devices into a worldwide interconnected computer infrastructure which facilitates almost any type of human endeavor.
-The Computer-supported Collaboration Work (CSCW) \cite{grudin1994computer} emerged in 1984 represents such shift.
-It addresses ``how collaborative activities and their coordination can be supported by means of computer systems''.
-This concept later evolved into a broader sense of  Computer-supported collaboration (CSC).
-The Social computing \cite{parameswaran2007research} is another shift, which refers systems that support ``the gathering, representation, processing, use, and dissemination of information that is distributed across social collectivities such as teams, communities, organizations, and markets''.
-To compare with the CSC, the social computing has more focus in `computing' rather than `collaboration'.
-Similarly, the Human Computation \cite{von2009human} is another paradigm towards the computational direction, which utilize human processing power to solve problems that computers cannot yet solve.
-All these three mentioned fields overlap in their concepts and are mainly popularised in the computer science.
-In the Web 2.0 era, there were few generic concepts/buzzword emerged on the idea of utilising the human power which are not only used in computer science, but also in other fields such as social science, business etc.
-Crowding sourcing \cite{howe2006crowdsourcing} as an example, describes the idea that taking a job traditionally performed by a designated agent and outsourcing it to an undefined, generally large group of people in the form of an open call.
-Open Innovation \cite{chesbrough2008open} advocates the similar idea with empathizing on the innovation in firms.
-Surowiecki also argued in the book of the wisdom of crowds \cite{surowiecki2005wisdom} that the aggregation of information in groups, resulting in decisions that are often better than could have been made by any single member of group.
-Collective Intelligence described by Malone \cite{malone2009harnessing}, defines an even more broadly concept that ``groups of individuals doing things collectively that seem intelligent.''
-The concept overlapping between human computation, crowdsourcing, social computing and collective intelligence is discussed in \cite{quinn2011human}.
-In  social machines  the  social and  the  technical are equal  partners.
-While  crowdsourcing  could  be  viewed  as  'human  subroutines'  and  'citizen  sensors',  the social machine perspective is one of incentivised participation and not humans enslaved to computers.
-We define the scope of social machine in a conceptual digram in Figure \ref{socialmachine}.
+The Web as a global platform for information access and sharing marked a second essential milestone; in particular, through principles and technologies promoted through Web $2.0$ and the Mobile Web.
+These developments led to amazing growth in terms of the amounts of content available online and mass participation.
+They are responsible for hundreds of millions of users all over the globe creating high-quality encyclopedias, publishing Terabytes of multimedia content, contributing to world-class software, and lively taking part in defining the agenda of many aspects of our society.
+This trend towards 'prosumerism' is finding more and more adopters in the public and private sector as well, as governments and enterprises not only become active in open initiatives, but encourage the participation of their customers and employees in taking decisions related to organizational management, product development, services offers, and policies.
+In this context, a number of terms are used to refer to the ways people interact with each other and with applications: 'wisdom of the crowds', 'collective intelligence', 'open innovation', 'crowdsourcing', 'human computation', and 'social computing'.
+They are related, but not synonymous with 'social machines', as explained in the following (see also Figure \ref{socialmachine}).
 
@@
 \includegraphics[width=8cm]{img/socialmachinescope.png}
-\caption{Social machines and related areas} \label{social machine}
+\caption{Social machines and related areas} \label{fig:social machine}
 \end{center}
@@
 
+Wisdom of the crowds \cite{surowiecki2005wisdom} refers to a principle for decision making that takes into account information and opinions of a group of people rather than individuals; the use of specific technologies, most notably Web $2.0$, has made it possible for such processes to be carried out at scales hardly conceivable in the past, and to involve highly diverse and geographically distributed participants.
+A similar concept, though broader scoped, is collective intelligence, defined in  \cite{malone2009harnessing} as 'groups of individuals doing things collectively that seem intelligent.' The main difference to what we refer to as 'social machines' is hidden in the the automation part of the latter.
+In a social machine, human and computational intelligence coalesce in order to achieve a given purpose.
+The overall setting may slightly vary in terms of the interaction patterns between the two types of services and the way their outcomes are used; however, where wisdom of the crowds and collective intelligence clearly place their focus on identifying the situations in which groups of people perform better than individuals, social machines are concerned with the study and realization of hybrid systems where the two types of components co-exist.
+As such, theoretical and empirical insights from the two related areas are useful to understand the dynamics of the social structures underlying such a system, but they are definitely not the only ingredient needed to build operational social machines.
+The question of how human and automatic services can be brought together to achieve optimal results, as well as the actual engineering exercise by which a system is developed, tested, and updated are equally important.
+
+One of the direct consequences of the popularity of the wisdom of the crowds idea was a stronger investment worldwide in open innovation, which can be seen as an application of the concept to business environments, or, in the words of the authors, as a 'a paradigm that assumes that firms can and should use external ideas as well as internal ideas, and internal and external paths to market, as the firms look to advance their technology' \cite{chesbrough2003}.
+At a more general level, this kind of creativity could be leveraged in almost any domain which benefits from diversity, while in the same time ensuring in-time access of a potentially infinite pool of skills and resources not available before the advent of social computing technologies.
+The term crowdsourcing is typically associated with this larger collection of situations, in which 'a job traditionally performed by a designated agent [...] [is outsourced] to an undefined, generally large group of people in the form of an open call' \cite{howe2006crowdsourcing}.
+Human computation applies human processing power to tackle technical tasks that computers (still) find challenging \cite{von2009human}, typically in areas such as visual, audio, and natural language understanding.
+This sort of tasks are an important part of today's crowdsourcing projects landscape, in particular on so-called 'microtask' platforms such as Amazon's Mechanical Turk\footnote{\url{https://www.mturk.com/}} or CrowdFlower\footnote{\url{http://crowdflower.com/}}, which offer small financial rewards to an anonymous crowd engaged with atomic units of work in the range of seconds to minutes to complete.
+The key difference between social machines and open innovation and crowdsourcing is again in the computational part.
+The novelty these two approaches bring in lies in using a much larger pool of human resources as traditional work environments \cite{quinn2011human}.
+There are clear socio-economic implications that their adopters need to deal with in order to optimally make use of this wealth of resources; technology may be needed to assist specific aspects of crowdsourcing projects, from evaluating and rewarding the results produced by the crowd to consolidating and aggregating them into a complete solution.
+A principled combination of human and computational capabilities is less in the focus of such systems and the technical means that support them, by contrast to social machines.
+In comparison to human computation, social machines cover a wider range of scenarios.
+Human computation is AI-centric and uses people to perform tasks that computers are not (yet) able to tackle (in terms of accuracy) \cite{quinn2011human}; by comparison, we see many successful examples of social machines in which the role of machines is rather to facilitate interactions within groups of people or communities of interest.
+
+An analogous line of reasoning could be followed to point out the overlaps between social machines and social computing \cite{parameswaran2007research}.
+The latter is an area of computer science which refers systems that support 'the gathering, representation, processing, use, and dissemination of information that is distributed across social collectivities such as teams, communities, organizations, and markets' \cite{parameswaran2007research}.
+As such, compared to the general concept of 'Computer-supported collaboration', social computing puts a greater emphasis on the information management capabilities of groups and communities, and less on the way these capabilities emerge as a joint effort.
+This distinction is even stronger in the case of social machines, which look at the social and the technical components as equal and necessary partners, and study the ways they could be best combined to master the challenges of future socio-technical systems.
+
+We now turn to a description of the main classification dimensions and features relevant to describe, study, and compare social machines.
+
 \section{Classification of social machines}
+\label{sec:framework}
 
-\subsection{Methodology} \label{meth}
-
+\subsection{Methodology}
+\label{sec:methodology}
 The classification framework presented in this paper was created through knowledge elicitation ~\cite{knowledgeelicitation}. In particular we used the {\it repertory grid} elicitation
@@
 The user then rates every element with value from $1$ to $5$ on this construct, where $1$ represents an element that is purely
-'For Science', and $5$ one that mainly serves  'socialising' purposes.
+'For Science', and $5$ one that mainly serves  'socializing' purposes.
 Elements can also be rated with values between the poles if they may correspond to both or neither.
@@
 While determining the intersection of elements was straightforward, the consolidation of the
-constructs required a more thorough process. We embarked on a manual process of grouping the constructs into rough clusters, based around the areas they cover. For example, we used construct
-clusters including: Users, Motivation, Popularity, Technology, and Purpose. We then examined
-each construct to determine which were equivalent, and whether we could re-word or subsume
-existing constructs to cover the same aspects as a construct, with the aim to remove any
-redundancy, to cut the list of constructs down to a manageable number, but to ensure that
-all of the constructs that were elicited were somehow represented in the final set. This
-process involved four of the authors discussing the choices, and finally agreeing on a
-consolidated set of $31$ constructs.
+constructs required a more thorough process.
+We manually grouped the constructs into rough clusters, based around the areas they cover.
+We examined each construct to determine which were equivalent, and whether we could re-word or subsume existing constructs to cover the same aspects.
+The aim of this exercise was twofold: to remove any redundancy and cut the classification space down to a manageable size, while ensuring that all constructs that were elicited were represented in the final set.
+This process involved four of the authors discussing the choices, and resulted in a consolidated set of $31$ constructs organized in four different clusters:
 
-%Even with this reduced set of constructs, if a single person were to classify all 56
-%elements, this would still require 1736 individual decisions, and hence is unfeasible to
-%expect full coverage of the elements using all of these constructs, despite how representative
-%they are.
-%
-%[something here about doing 2 grids each and verifying the results?]
-%
-%To extract greater usefulness, and the ability to use our constructs to classify social
-%machines we re-evaluated the constructs and the semantics that they represent into richer
-%constructs that are not constrained by the 1-5 rating of repertory grid constructs.
+\begin{description}
+  \item [Popularity]: this cluster mainly covered descriptive information about a given system, such as its maturity and number of users.
+For the purpose of our analysis we decided to discard the associated constructs, as they
+  \item [Tasks and purpose]
+  \item [Motivation and incentives]
+  \item [Technology]
+\end{description}
+
+The paper presents the current version of the framework.
+We envision a reiteration of the main clusters and constructs based on discussions with and feedback from various scientific audiences.
+We provide initial evidence of the usefulness of the framework as a tool to  examine commonalities and differences between social machines in Section \ref{sec:usage}, in which we applied the repertory grid elicitation technique to a set of $20$ social machines instances ranked by popularity according to the Alexa service.\footnote{\url{www.alexa.com/}} A more thorough evaluation will
 
@@
 \hline
-{\bf Context of Participation} \\
+{\bf Context(s) of Participation} \\
 \hline
 Participants' participate to fulfill needs of a role  \\
-\hspace{1cm} work-related/professional role
-\hspace{1cm} home/family related
-\hspace{1cm} social role
-\hspace{1cm} leisure/entertainment role
-Participants' physical location is relevant to the service \\
+\hspace{1cm} work-related/professional role \\
+\hspace{1cm} home/family related \\
+\hspace{1cm} social role \\
+\hspace{1cm} leisure/entertainment role \\
 Participation is done via: \\
@@
 Participation is done in a mobile context \\
+\hspace{1cm} Physical location is relevant to the service \\
 \hline
@@
 
-\subsubsection{Tasks, Problems and Activiites}
+\subsubsection{Tasks, Problems and Activities}
 The first set of constructs pertain to \emph{what} the participants
@@
 \subsection{Using the classification framework}
-
+\label{sec:usage}
 In order to demonstrate usage of our framework's constructs, we ranked our social machines
@@
 This process involved ranking each social machine against the constructs, rating each construct
-from 1 to 5. In total, 680 ranking were made, and the
+from 1 to 5. In total, $680$ ranking were made, and the
 results are illustrated in a grid and accompanying dendrograms for element and construct

### 2013-02-26 latex/paper_new.tex
intermediary version

@@
 We envision a reiteration of the main clusters and constructs based on discussions with and feedback from various scientific audiences.
-We provide initial evidence of the usefulness of the framework as a tool to  examine commonalities and differences between social machines in Section \ref{sec:usage}, in which we applied the repertory grid elicitation technique to a set of $20$ social machines instances ranked by popularity according to the Alexa service.\footnote{\url{www.alexa.com/}} A more thorough evaluation will
+We provide initial evidence of the usefulness of the framework as a tool to  examine commonalities and differences between social machines in Section \ref{sec:usage}, in which we applied the repertory grid elicitation technique to a set of $20$ social machines instances ranked by popularity according to the Alexa service.\footnote{\url{www.alexa.com/}} A more thorough evaluation will assess the extent to which the classification constructs identified can be meaningfully used to
 

### 2013-02-26 latex/paper_new.tex
another intermediary version

--- 2013-02-26 latex/paper_new.tex
@@
 \begin{center}
-\includegraphics[width=8cm]{img/socialmachinescope.png}
+\includegraphics[width=0.3\textwidth]{img/socialmachinescope.png}
 \caption{Social machines and related areas} \label{fig:social machine}
@@
 that differentiate between elements - for example,  prompting the user to create a construct that
-differentiates between {\it GalaxyZoo} and {\it Facebook}, as two prominent example of social machines our classification framework should apply to.
+differentiates between {\it GalaxyZoo}\footnote{In GalaxyZoo people classify galaxies with collective performance as good as professional astronomers, see \url{http://www.galaxyzoo.org}.} and {\it Facebook}, as two prominent example of social machines our classification framework should apply to.
 The user describes the opposing poles of the construct - in this case, the user may decide on 'For Science' and 'For Socialising' in order to capture the core distinction between the two system.
@@
 The aim of this exercise was twofold: to remove any redundancy and cut the classification space down to a manageable size, while ensuring that all constructs that were elicited were represented in the final set.
-This process involved four of the authors discussing the choices, and resulted in a consolidated set of $31$ constructs organized in four different clusters:
-
-\begin{description}
-  \item [Popularity]: this cluster mainly covered descriptive information about a given system, such as its maturity and number of users.
-For the purpose of our analysis we decided to discard the associated constructs, as they
-  \item [Tasks and purpose]
-  \item [Motivation and incentives]
-  \item [Technology]
-\end{description}
+This process involved four of the authors discussing the choices, and resulted in a consolidated set of $31$ constructs organized in four different clusters: popularity, tasks and purpose, participants and roles, and motivation and incentives.
+In our analysis we will not further consider the first cluster, which primarily covered descriptive information about a given system, such as its perceived level of maturity and current number of users.
+Further on, the constructs were not evenly distributed across clusters; whereas some aspects such as roles of participants occurred frequently (in various flavors) in the grids provided by the ten experts, others such as the different types of workflows synthesizing human and automatized capabilities of a social machine were less taken into account.
+As discussed in Section \ref{sec:comparison} we believe that the interaction between these two types of services is an essential part of the theory and practice of social machines, being one of the elements that distinguishes them from related areas which are biased towards one or the other.
+Extending the framework with constructs capturing these aspects (see, for instance, \cite{quinn2011human} for examples in the context of human computation) is part of our future work.
 
@@
 We envision a reiteration of the main clusters and constructs based on discussions with and feedback from various scientific audiences.
-We provide initial evidence of the usefulness of the framework as a tool to  examine commonalities and differences between social machines in Section \ref{sec:usage}, in which we applied the repertory grid elicitation technique to a set of $20$ social machines instances ranked by popularity according to the Alexa service.\footnote{\url{www.alexa.com/}} A more thorough evaluation will
+We provide initial evidence of the usability of the framework as a tool to  examine commonalities and differences between social machines in Section \ref{sec:usage}, in which we applied the repertory grid elicitation technique to a set of $20$ social machines instances ranked by popularity according to the Alexa service.\footnote{\url{www.alexa.com/}} A more thorough evaluation will determine the extent to which the classification constructs identified can be meaningfully used by researchers and system designers and operators.
+We will test the completeness, correctness, and comprehensibility of the constructs in experiments in which a new set of social machines will be classified by framework users; we will ask the participants to assess the quality of the framework along these general dimensions, and measure inter-annotator agreement to learn about the usefulness of the classifications produced.
 
-\subsection{The Polyarchical Relationship of Social Machines}
+\subsection{The polyarchical relationship of social machines}
+When defining the boundaries of what we call social machines, one important observation was related to the distinction between platforms and technologies such as wikis and GWAP ('games-with-a-purpose'), which enable social machines to be created, and their instantiations into social machines that were brought to life for a specific purpose, such as Wikipedia and EteRNA\footnote{\url{http://eterna.cmu.edu}}, a game in which  participants design RNA  that is evaluated automatically using simulators.
+A second observation is concerned with the relationship between broader and narrower-scoped instances of social machines; general-purpose examples such as Facebook, Twitter or Amazon, or even the Web, enable the formation of more specific social machines and communities within them, that are working to solve solve different problems.
+An example illustrating this relationship is the Obama'12 US Presidential campaign~\cite{obamakieron}, which relied on Twitter and Facebook as basic social machines; most forms of customer relationship management and community engagement carried out in the digital space use similar tools.
+Finally, social machines are interconnected into a greater ecosystem, both at the social and at the computation levels.
+For example, insights on the classification of space objects gained through GalaxyZoo may influence the content of the corresponding articles, and the associated editorial processes, within Wikipedia.
+In the same time,a large number of communities within Wikipedia focus specific topics and activities, such as those contributing to science and technology.\footnote{Visualization of Wikipedia Science Communities: \url{http://www.olihb.com/WikiCommunities/}}
 
-While determining a set of social machines, one clear distinction was between social machine frameworks, such as MediaWiki and Ushahidi, which enable
-social machines to be created, and instantiations of those frameworks into social machines that have been created for a specific purpose, such as Wikipedia
-and Ushahidi Haiti. However, through further investigation, we discovered that the distinction between framework and instantiations was not the only
-differentiator. Specifically, that some social machines could also have sub-machines or communities within them, that are working to solve different problems,
-within the context of the larger machine. For example, Wikipedia is an instantiation of the MediaWiki framework, but there are a large number of
-communities within Wikipedia that work on domain-specific areas, such as those contributing in areas of science and technology\footnote{Visualisation of Wikipedia Science Communities: \url{http://www.olihb.com/WikiCommunities/}}.
-However, we must also note that wikipedia is primarily split into communities of language, with the English, German and French wikipedias having the largest number of articles.
-It is therefore possible to analyse wikipedia into a hierarchy, from MediaWiki, to the languages of Wikipedia, down to the domain-specific communities.
-We explored this idea further, to include other social machines, and links where social machines utilise other social machines in order to function.
-For example, the social machine used by the Obama '12 US Presidential campaign~\cite{obamakieron}, which relied on social machines such as Twitter and Facebook.
-
-Looking at the set of social machines in a polyarchy leads to a broad/specific relationship emerging that lets us talk about behaviours at various levels of granularity.
-We propose looking at nested machines with ``The Web'' as one of several potential roots, with the next level down consisting of sub-platforms (such as Facebook, Twitter, MediaWiki, Ushahidi platforms) that spawn more specific social machines.
-A resulting polyarchy is shown in Figure~\ref{polyarchy}, which classifies levels as Infrastructure, Frameworks, Services and Causes/Groups.
+Looking at the set of social machines in a polyarchy leads to a broad/specific relationship emerging that lets us talk about behavior at various levels of granularity.
+We propose looking at nested machines with 'The Web' as one of several potential roots, with the next level down consisting of sub-platforms (such as Facebook, Twitter, MediaWiki, Ushahidi) that spawn more specific social machines.
+A resulting polyarchy is shown in Figure~\ref{polyarchy}, which classifies levels as infrastructure, frameworks, services, and projects and initiatives.
 
@@
 \includegraphics[width=8.5cm]{img/polyarchy.pdf}
-\caption{A polyarchy of Social Machines, illustrating the infrastructure and frameworks used by social machines, and machine-machine usage.} \label{polyarchy}
+\caption{A polyarchy of social machines, illustrating the infrastructure and frameworks used by social machines, and machine-machine usage.} \label{polyarchy}
 \end{center}
@@
 how to they differ dynamically?
+How do certain design decisions taken at the level of the infrastructure, frameworks and service propagate into narrower-focused systems that are built on top of them?
+Similarly, how will such decisions affect a broader ecosystem of social machines, each with their own, though overlapping, purposes and communities?
 
-One originally unforeseen classification is that of the causes that utilise Social Machines.
-Through a particularly popular cause, large numbers of participants can be mobilised, for
-either a short period of time, or for long-term engagement, using multiple social machines
-in order to further their cause. Examples of causes can be seen in Figure~\ref{polyarchy},
-at the lowest level of the polyarchy. One such example is the cause of ``Open Access'' to
-academic publications~\cite{harnad2001self}. This cause mobilises academic authors to
-self-archive their own papers online so they are available for free, in addition to hosted
-versions behind publishers pay-walls. In order to achieve this aim, the social machine of
-``open access'' typically uses the social machines of institutional repositories provided
-by the authors' associated institutions. As with other social machines, these social
-machines typically use off-the-shelf implementations (in this case the most popular are
-EPrints~\cite{eprints} and DSpace~\cite{dspace}), which run as web sites on the world wide
-web. This cause is mature, growing year-on-year and has global coverage.
+The polyarchy is complementary to the classification constructs we elicitated.
+The constructs apply to all levels of the polyarchy, though some of them might be more easier to understand and describe in the context of some levels.
+For example, instances of social machines residing at the bottom of the polyarchy are likely to be built with a very concrete audience in mind, and as such the mechanisms to incentivize participation are likely to be clearer and more straightforward to study and adjust than in cases in which such boundaries cannot be defined, such as Facebook, or even Wikipedia.
+In the same time, these systems will have to decide among an array of diverse basic services and frameworks for their implementations; the impact of such engineering decisions is difficult to predict in great detail as a thorough understanding of what makes certain social machines more successful than others, in terms of technology choices and beyond, is largely missing for most application domains.
 
+%One such example is the 'open access' movement in scholarly publishing
+%~\cite{harnad2001self}.
+This cause mobilizes academic authors to self-archive their own papers online so they are available for free, in addition to hosted versions behind publishers pay-walls.
+In order to achieve this aim, the social machine of 'open access' typically uses the social machines of institutional repositories provided
+%by the authors' associated institutions. As with other social machines, these social
+%machines typically use off-the-shelf implementations (in this case the most popular are
+%EPrints~\cite{eprints} and DSpace~\cite{dspace}), which run as web sites on the world wide
+%web. This cause is mature, growing year-on-year and has global coverage.
+
+%One originally unforeseen classification is that of the causes that utilise Social Machines.
+%Through a particularly popular cause, large numbers of participants can be mobilised, for
+%either a short period of time, or for long-term engagement, using multiple social machines
+%in order to further their cause. Examples of causes can be seen in Figure~\ref{polyarchy},
+%at the lowest level of the polyarchy. 
 %\subsubsection{Multi-faceted motivation, and who-gets-what benefit}
@@
 \subsection{Constructs}
-
 The clustering process described in Section~\ref{meth} yielded both a
 reduction in the overall number of constructs (in particular an
-elimination of overrepresented areas) but also identificaiton of four
+elimination of over-represented areas) and the identification of four
 simple themes: sets of constructs pertaining to users and roles within
@@
 performed through and using the machine, third, motivation for
-participation, finally, the contexts of participation.  Each of these
-clusters contain 5-8 constructs as illustrated in Table
+participation, finally, the contexts of participation.
+
+Each of these clusters contain between five and eight constructs as illustrated in Table
 \ref{table:constructs}; we describe each of these themes in the
@@
 \hline
-{\bf Users and Roles} \\
-\hline
-Generality of Audience \\
-Participant autonomy \\
-Participant anonymity \\
-Extent of hierarchical organisation of roles \\
-Clear separation of roles/responsibilities among participants \\
-\hline
-{\bf Tasks} \\
+
+{\bf Tasks, purpose and context of participation} \\
 \hline
@@
 Variation in types of contributions and tasks \\
-\hline
-{\bf Motivation} \\
-\hline
-Participants are intrinsically motivated:\\
-\hspace{1cm} to gain/share knowledge \\
-\hspace{1cm} to ``get something done'' \\
-\hspace{1cm} to ``for fun/entertainment''\\
-\hspace{1cm} to ``be social'' \\
-\hspace{1cm} for the benefit of a specific group of people who need help \\
-\hspace{1cm} for the benefit of society as a whole \\
-Participation is motivated by extrinsic reward (payment, status etc) \\
-\hline
-{\bf Context(s) of Participation} \\
-\hline
 Participants' participate to fulfill needs of a role  \\
@@
 \hline
+{\bf Participants, roles and incentives} \\
+\hline
+Generality of audience \\
+Participant autonomy \\
+Participant anonymity \\
+Extent of hierarchical organization of roles \\
+Clear separation of roles/responsibilities among participants \\
+Participants are intrinsically motivated:\\
+\hspace{1cm} to gain/share knowledge \\
+\hspace{1cm} to ``get something done'' \\
+\hspace{1cm} to ``for fun/entertainment''\\
+\hspace{1cm} to ``be social'' \\
+\hspace{1cm} for the benefit of a specific group of people who need help \\
+\hspace{1cm} for the benefit of society as a whole \\
+Participation is motivated by extrinsic reward (payment, status etc) \\
+\hline
+
 \end{tabular}
@@
 
-\subsubsection{Users and Roles}
-The first set of constructs pertain to the human participants of
-social machines and the way(s) participants are organised within them.
-One construct identifies whether participants constitute members of
-the general public, or are of a specific demongraphic, external group
-or organisation, occupation, or of a particular expertise.  Two
-constructs, \emph{autonomy} and \emph{anonymity} pertain to the degree
-to which interaction is constrained among participants, and to the
-degree to which participants' identities are used and disclosed within
-the machine. The final two constructs pertain to the roles: first,
-as to whether multiple roles are pre-differentiated by the machine or
-whether all participants essentially initially assume the same role,
-second whether roles (pre-defined or emergent) are archical.
-
-\subsubsection{Tasks, Problems and Activities}
+\subsubsection{System view: tasks, purpose and context of participation}
 The first set of constructs pertain to \emph{what} the participants
@@
 
-\subsubsection{Motivation}
+The final set of constructs pertains to the context(s) of each
+participant's greater (life-) activities that participation is
+performed.  This context could be in a work context, home/family
+context, for leisure, and at a desk or away from a computer (in a
+mobile context).  PArticipation could be conducted via apps, web
+browsers, sensors and other technolgoies.  Due to the predominance of
+``geo-locative'' services that incorporate information about the
+user's location, we have included a construct identifying the degree
+to which the user's location is used by the machine.
+
+\subsubsection{Social view: users, roles, motivation and incentives}
+
+The first set of constructs pertain to the human participants of
+social machines and the way(s) participants are organised within them.
+One construct identifies whether participants constitute members of
+the general public, or are of a specific demongraphic, external group
+or organisation, occupation, or of a particular expertise.  Two
+constructs, \emph{autonomy} and \emph{anonymity} pertain to the degree
+to which interaction is constrained among participants, and to the
+degree to which participants' identities are used and disclosed within
+the machine. The final two constructs pertain to the roles: first,
+as to whether multiple roles are pre-differentiated by the machine or
+whether all participants essentially initially assume the same role,
+second whether roles (pre-defined or emergent) are archical.
+
 The third set of themes pertains to the motivation that sustains
@@
 could include money reward, status, recognition and so on
-
-\subsubsection{Context of Participation}
-The final set of constructs pertains to the context(s) of each
-participant's greater (life-) activities that participation is
-performed.  This context could be in a work context, home/family
-context, for leisure, and at a desk or away from a computer (in a
-mobile context).  PArticipation could be conducted via apps, web
-browsers, sensors and other technolgoies.  Due to the predominance of
-``geo-locative'' services that incorporate information about the
-user's location, we have included a construct identifying the degree
-to which the user's location is used by the machine.
 

### 2013-02-26 latex/paper_new.tex
pre-final version

@@
 When we fill out a form (e.g., health information, birds spotted in my garden, new construction sites on my way to work) the systems supporting this activity are minimal forms of social machines  because the users  are  part  of  the  'social  computation', in this case the data creation and collection process facilitate through the site.
-This social component  becomes  richer when  the database  is  curated by members  of  the broader community (e.g., Wikipedia) and when  the social network adds value implicitly (Amazon) or explicitly (Facebook) to the overall system through their individual or joint activities.
+This social component  becomes  richer when  the database  is  curated by members  of  the broader community (e.g., Wikipedia) and when  the social network adds value implicitly (e.g., Amazon) or explicitly (e.g., Facebook) to the overall system through their individual or joint activities.
 
@@
 
-This paper makes two main contributions:
+This paper makes two main contributions: First, we introduce the notion of 'social machines' and position it in connection to similar ideas and initiatives.
+Second we propose a set of constructs to capture the most important features of social machines and to compare existing instances thereof.
 
-\begin{itemize}
-  \item We introduce the notion of 'social machines' and position it in connection to similar ideas and initiatives.
-  \item We propose a set of constructs to capture the most important features of social machines and to compare existing instances thereof.
-\end{itemize}
-
-The remainder of this paper is organized as follows.
+The remaining sections are organized as follows.
 In order to pin down what makes up a social machine, and establish the boundaries of this emerging field of research, we need to first understand the relationship between social machines and related topics such as human computation, collective intelligence, and crowdsourcing; Section \ref{sec:comparison} gives a summary of this comparative analysis.
@@
 In this context, a number of terms are used to refer to the ways people interact with each other and with applications: 'wisdom of the crowds', 'collective intelligence', 'open innovation', 'crowdsourcing', 'human computation', and 'social computing'.
-They are related, but not synonymous with 'social machines', as explained in the following (see also Figure \ref{socialmachine}).
+They are related, but not synonymous with 'social machines', as explained in the following (see also Figure \ref{fig:socialmachine}).
 
@@
 \begin{center}
-\includegraphics[width=0.3\textwidth]{img/socialmachinescope.png}
+\includegraphics[width=0.25\textwidth]{img/socialmachinescope.png}
 \caption{Social machines and related areas} \label{fig:social machine}
@@
 technique~\cite{kelly} in order to derive an initial set of {\it elements}, which represent instances of social machines, and {\it constructs}, which capture their most important characteristics.
-
 In repertory grid elicitation a software tool is used to ask users to describe constructs
@@
 differentiates between {\it GalaxyZoo}\footnote{In GalaxyZoo people classify galaxies with collective performance as good as professional astronomers, see \url{http://www.galaxyzoo.org}.} and {\it Facebook}, as two prominent example of social machines our classification framework should apply to.
-The user describes the opposing poles of the construct - in this case, the user may decide on 'For Science' and 'For Socialising' in order to capture the core distinction between the two system.
+The user describes the opposing poles of the construct - in this case, the user may decide on 'For Science' and 'For socialising' in order to capture the core distinction between the two system.
 The user then rates every element with value from $1$ to $5$ on this construct, where $1$ represents an element that is purely
@@
 \begin{center}
-\includegraphics[width=8.5cm]{img/polyarchy.pdf}
-\caption{A polyarchy of social machines, illustrating the infrastructure and frameworks used by social machines, and machine-machine usage.} \label{polyarchy}
+\includegraphics[width=0.4\textwidth]{img/polyarchy.pdf}
+\caption{A polyarchy of social machines, illustrating the infrastructure and frameworks used by social machines, and machine-machine usage} \label{polyarchy}
 \end{center}
@@
 
-The polyarchy is complementary to the classification constructs we elicitated.
+The polyarchy is complementary to the classification constructs we elicited.
 The constructs apply to all levels of the polyarchy, though some of them might be more easier to understand and describe in the context of some levels.
@@
 reduction in the overall number of constructs (in particular an
-elimination of over-represented areas) and the identification of four
-simple themes: sets of constructs pertaining to users and roles within
+elimination of over-represented areas) and the identification of three
+simple themes: a
+
+sets of constructs pertaining to users and roles within
 each machine, second, the nature of tasks, problems and activities
@@
 \begin{center}
+\begin{scriptsize}
 \begin{tabular}{|p{8cm}|}
 \hline
-
 {\bf Tasks, purpose and context of participation} \\
@@
 Tasks are domain-specific \\
-The machine owner (service/platform) derives benefit from participation \\
+The machine owner derives benefit from participation \\
 Activities and tasks are pre-defined or participant-defined \\
@@
 \hspace{1cm} mobile devices \\
-\hspace{1cm} web browsers \\
+\hspace{1cm} Web browsers \\
 \hspace{1cm} apps \\
@@
 \hline
-{\bf Participants, roles and incentives} \\
+{\bf Participants and roles} \\
 \hline
@@
 Extent of hierarchical organization of roles \\
-Clear separation of roles/responsibilities among participants \\
+Clear separation of roles among participants \\
+{\bf Motivation and incentives} \\
+\hline
 Participants are intrinsically motivated:\\
@@
 \hspace{1cm} for the benefit of society as a whole \\
-Participation is motivated by extrinsic reward (payment, status etc) \\
+Participation is motivated by extrinsic reward (payment, status) \\
 \hline
@@
 \end{tabular}
+\end{scriptsize}
 \end{center}
-\caption{Consolidated constructs of social machines.} \label{table:constructs}
+\caption{Consolidated constructs of social machines} \label{table:constructs}
 \end{table}
 
-\subsubsection{System view: tasks, purpose and context of participation}
+\subsubsection{Tasks, purpose, and context of participation}
 The first set of constructs pertain to \emph{what} the participants
@@
 appraisal of existing content, posing of problems, solving of
-problems, andf so on.  The resulting computation, looked at from a
+problems, and so on.  The resulting 'computation', looked at from a
 macroscopic whole, might be the identification of high-quality
@@
 solving, and so on.  Such tasks might be pre-defined by the system
-designers, or brought to the system by participants.  Futher, these
+designers, or brought to the system by participants.  Further, these
 tasks might be of a specific type or class, or of many varied types or
-encompass a general class of activities.  To capture these
-difeferences, we include a construct that differentiates machines with
-pre-defined problem spaces or are open to participant-provided tasks
-and activities.  Finally, we ask whether the benefit derived by
+encompass a general class of activities.  To capture this diversity, 
+we include a construct that differentiates machines with pre-defined problem spaces or are open to participant-provided tasks and activities.  Finally, we ask whether the benefit derived by
 participants is distinct from that gained by machine/platform owners
@@
 
-The final set of constructs pertains to the context(s) of each
-participant's greater (life-) activities that participation is
-performed.  This context could be in a work context, home/family
-context, for leisure, and at a desk or away from a computer (in a
-mobile context).  PArticipation could be conducted via apps, web
-browsers, sensors and other technolgoies.  Due to the predominance of
-``geo-locative'' services that incorporate information about the
-user's location, we have included a construct identifying the degree
-to which the user's location is used by the machine.
+The second set of constructs pertains to the context(s) of each participant's greater (life-) activities in which the interaction with the system occurs.
+This context could be in a work context, home/family context, for leisure, and at a desk or away from a computer (in a mobile context).
+Participation could be conducted via apps, Web browsers, sensors and other technologies.
+Due to the predominance of geolocation services, one of the constructs collected through knowledge elicitation identifies the degree to which the user's location is used by the machine.
 
-\subsubsection{Social view: users, roles, motivation and incentives}
+\subsubsection{Participants and roles}
+This set of constructs pertain to the human participants of social machines and the way(s) participants are organized within them.
+One construct identifies whether participants constitute members of the general public, or are of a specific demographic, external group or organization, occupation, or of a particular expertise.
+Two constructs, \emph{autonomy} and \emph{anonymity} pertain to the degree to which interaction is constrained among participants, and to the degree to which participants' identities are used and disclosed within the machine.
+Additional two constructs pertain to the roles: first, as to whether multiple roles are pre-differentiated by the machine or
+whether all participants essentially initially assume the same role, second whether roles (pre-defined or emergent) are hierarchical.
 
-The first set of constructs pertain to the human participants of
-social machines and the way(s) participants are organised within them.
-One construct identifies whether participants constitute members of
-the general public, or are of a specific demongraphic, external group
-or organisation, occupation, or of a particular expertise.  Two
-constructs, \emph{autonomy} and \emph{anonymity} pertain to the degree
-to which interaction is constrained among participants, and to the
-degree to which participants' identities are used and disclosed within
-the machine. The final two constructs pertain to the roles: first,
-as to whether multiple roles are pre-differentiated by the machine or
-whether all participants essentially initially assume the same role,
-second whether roles (pre-defined or emergent) are archical.
-
-The third set of themes pertains to the motivation that sustains
-continued participation in these systems.  We identiy 6 different
+\subsubsection{Motivation and incentives}
+Related to the social structure  motivation that sustains
+continued participation in these systems.  We identify six different
 kinds of common sources of intrinsic motivation; first, that
@@
 factors also contributed to sustained participation - such factors
-could include money reward, status, recognition and so on
+could include money reward, status, recognition and so on \cite{kuznetsov2006motivations,lakhani2003hackers,tagging}.
+%Peer-to-peer filesharing~\cite{p2p}; Tagging~\cite{tagging}; and Virtual
+%Communities~\cite{ardichvili2003motivation,Moore:2007:UMM:1235000.1235035}.
 
@@
 \label{sec:usage}
-In order to demonstrate usage of our framework's constructs, we ranked our social machines
-by their Alexa ranking, and performed a full repertory grid elicitation exercise over them.
-This process involved ranking each social machine against the constructs, rating each construct
-from 1 to 5. In total, $680$ ranking were made, and the
+In order to demonstrate usage of the constructs we ordered our social machines
+by their Alexa ranking, and performed a full repertory grid elicitation exercise over the first $20$ of the $56$ collected in the first phase of the methodology (see Section \ref{sec:methodology}).
+This process assessing each social machine against the constructs by choosing a rate between $1$ and $5$.
+This led to a total of $680$ ratings, and the
 results are illustrated in a grid and accompanying dendrograms for element and construct
@@
 \begin{center}
-\includegraphics[width=4cm]{img/dendrogram-elements.png}
-\caption{Dendrogram of elements (social machines) from a repertory grid exercise of the top 20 (ranked by Alexa) social machines from our element set, against our consolidated constructs.} \label{dendrogram}
+\includegraphics[width=0.2\textwidth]{img/dendrogram-elements.png}
+\caption{Dendrogram of elements (social machines) from a repertory grid exercise of the top $20$ (ranked by Alexa) social machines from our element set, against our consolidated constructs} \label{dendrogram}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=8cm]{img/dendrogram-constructs.png}
-\caption{Dendrogram of constructs from a repertory grid exercise of the top 20 (ranked by Alexa) social machines from our element set, against our consolidated constructs.} \label{dendrogram}
+\includegraphics[width=0.4\textwidth]{img/dendrogram-constructs.png}
+\caption{Dendrogram of constructs from a repertory grid exercise of the top $20$ (ranked by Alexa) social machines from our element set, against our consolidated constructs} \label{dendrogram}
 \end{center}
@@
 From the dendrograms we can get a sense of similarities and separation of social machines. For
-instance, it is clear that YouTube, Vimeo, Reddit and Digg are similar, which is logical, because
-they are all social machines that focus on sharing content. Likewise, Quora and StackOverflow
-are also similar, and again, this makes sense, because they are both question answering social
-machines. There is also a dendrogram for the constructs, to show correlations between them. This
-dendrogram indicates that there is a correlation between the ``Single Participant Roles'' and
-``No Hierarchy in Participant Roles'' constructs, which is because they refer to the same
-aspect of social machines. Another, but less obvious similarity is between the ``Large variety of
-social features'' and ``Benefit NOT for the world/society'', which indicates that of the social
-machines we have rated, there is a correlation between social features and social machines that
-do not benefit the world at large, and vice-versa.
+instance, it is clear that YouTube, Vimeo, Reddit, and Digg are similar, which is not unexpected, due to their focus on sharing content.
+Likewise, the exercise confirms commonalities between Quora and StackOverflow, which can be easily understood, given their question answering purpose.
+There is also a dendrogram for the constructs, which can be used, as more data is collected, as an indicator for correlation between constructs.
+In this particular case, dendrogram suggests that there is a correlation between the 'Single Participant Roles' and 'No Hierarchy in Participant Roles' constructs, which is because they refer to the same aspect of social machines.
+Another, but less obvious connection is between the 'Large variety of social features' and 'Benefit NOT for the world/society', which requires further investigation.
 
-We envisage two key initial uses for this framework: for classifying social machines; and for
-predicting behaviour of new/future social machines. For classifying social machines, it will
-enable direct comparisons to be made with existing machines, as well as provide a platform
-for discussions about social machines. We also think that the data from a large-scale classification
-exercise may be useful for researchers who wish to apply prediction techniques (such as SVMs)
-to social machines, their behaviour and impact.
+We envisage three key use cases for our framework.
+First and foremost the framework forms the basis for a coherent and comprehensive description and classification of social machines.
+The knowledge elicitation technique referred to in this paper is just one of the many tools which can assist this exercise.
+In addition, it enables direct comparisons between systems, and provides a platform for terminologically consistent discussions about the field as a whole and specific instantiations.
+Finally, we also think that with more and more classification data becoming available, for instance, also through functionality that automatically monitors and computes the values of the some of the constructs introduced earlier, researchers will wish to apply prediction techniques (such as SVMs) to social machines, their behavior, and impact.
 
 \section{Future work: building a social machine around the classification framework}
+In this paper we presented a first outline of a possible organization structure for the design space of social machines.
+We introduced the key features of the notion of social machines, and explained how this notion differs from related areas such as collective intelligence, crowdsourcing and human computing.
+We then discussed the basic constructs of a classification framework created through the repertory grid elicitation technique.
+In its current version the framework consists of $31$ constructs clustered according to the main components of social machines: social and machine-driven services, and the interactions between them.
+Our future work will mostly be dedicated to the revision of this initial list of constructs.
+Most notably, we would like to add a number of constructs to emphasize the importance of the interaction aspect in the theory, engineering, and operation of social machines.
+This includes constructs related to the human-computer workflows supported (e.g., human intelligence replacing automatic algorithms as in GWAPs vs the different bots available on Wikipedia complementing collaborative manual editing), the way inputs from the crowd are validated, aggregated and turned into a useful outcome, as well as issues related collaboration and coordination, and interface and communication design.
+In addition, we plan to engage with the broader researcher community by setting up a social machine to facilitate discussions and feedback collection.
+One specific idea that is currently under development is to build a microtask-like environment, including specific game elements, in which participants are asked to provide answers to atomic challenges that rate and compare a pair of social machine instances according to a construct of our framework.
 
 \section{Acknowledgments}
-
 This work is supported under SOCIAM: The Theory and Practice of Social Machines.

### 2013-02-26 latex/paper_new.tex
corrected some missing references
version sent to Nigel and Wendy

@@
 \includegraphics[width=0.25\textwidth]{img/socialmachinescope.png}
-\caption{Social machines and related areas} \label{fig:social machine}
+\caption{Social machines and related areas} \label{fig:socialmachine}
 \end{center}
@@
 \subsection{Constructs}
-The clustering process described in Section~\ref{meth} yielded both a
+The clustering process described in Section \ref{sec:methodology} yielded both a
 reduction in the overall number of constructs (in particular an

### 2013-02-26 latex/paper_new.tex
Fix quotation marks all being end-quotes

@@
 
-This paper provides a framework for...
+social machines, overview of what they are, their history
+    people interacting with content and with each other
 
@@
 \section{Social machines: an introduction}
-Once upon a time 'machines' were programmed by programmers and used by users.
+Once upon a time `machines' were programmed by programmers and used by users.
 The success of the Web has changed this relationship: we now see configurations of people interacting with content and with each other, typified by social Web sites.
-Rather than drawing a line through such Web-based systems to separate the human and digital parts (as computer science has traditionally done), we can now draw a line around them  and  treat  each  such  compound  as  a  'social  machine'  -  a  machine in which the two aspects are seamlessly interwoven.
+Rather than drawing a line through such Web-based systems to separate the human and digital parts (as computer science has traditionally done), we can now draw a line around them  and  treat  each  such  compound  as  a  `social  machine'  -  a  machine in which the two aspects are seamlessly interwoven.
 This crucial transition in thinking acknowledges the reality of current socio-technical systems, and is essential to underpin any understanding of the science and engineering of their future development towards pervasive ecosystems of co-evolving social machines.
 
-Essentially social machines can be characterized as assemblies of manually executed and machine-driven (as in 'automatized') services and the interaction of such services.
+Essentially social machines can be characterised as assemblies of manually executed and machine-driven (as in `automatised') services and the interaction of such services.
 A  traditional  database,  with  users  looking  up  records  independently  of  each  other  and  an  administrator responsible for the management of the content, has some of the right ingredients, but there is really no social element in this strict provider-consumer relationship.
-When we fill out a form (e.g., health information, birds spotted in my garden, new construction sites on my way to work) the systems supporting this activity are minimal forms of social machines  because the users  are  part  of  the  'social  computation', in this case the data creation and collection process facilitate through the site.
+When we fill out a form (e.g., health information, birds spotted in my garden, new construction sites on my way to work) the systems supporting this activity are minimal forms of social machines  because the users  are  part of the `social  computation', in this case the data creation and collection process facilitate through the site.
 This social component  becomes  richer when  the database  is  curated by members  of  the broader community (e.g., Wikipedia) and when  the social network adds value implicitly (e.g., Amazon) or explicitly (e.g., Facebook) to the overall system through their individual or joint activities.
@@
 
-This paper makes two main contributions: First, we introduce the notion of 'social machines' and position it in connection to similar ideas and initiatives.
+This paper makes two main contributions: First, we introduce the notion of `social machines' and position it in connection to similar ideas and initiatives.
 Second we propose a set of constructs to capture the most important features of social machines and to compare existing instances thereof.
@@
 We conclude with an outlook of our future development activities, including details about the evaluation of the framework.
-In addition, we plan to engage with the broader community to set up a 'social machine' bringing together researchers from various disciplines who will contribute to the 'computation' of a joint and widely agreed classification framework for the field.
+In addition, we plan to engage with the broader community to set up a `social machine' bringing together researchers from various disciplines who will contribute to the `computation' of a joint and widely agreed classification framework for the field.
 
@@
 \label{sec:comparison}
-Since the invention of the Turing machine we have experienced a 'paradigm shift' in the usage of computers, moving away from purely calculative devices to facilitators of a wide range of human interactions.
-The emergence of 'Computer-supported Collaboration Work (CSCW)' \cite{grudin1994computer} is representative for the early days of this trend.
-CSCW addresses 'how collaborative activities and their coordination can be supported by means of computer systems' \cite{grudin1994computer}, whereas the initial concept evolved towards the more broader field of 'Computer-supported collaboration' (CSC).
+Since the invention of the Turing machine we have experienced a `paradigm shift' in the usage of computers, moving away from purely calculative devices to facilitators of a wide range of human interactions.
+The emergence of `Computer-supported Collaboration Work (CSCW)' \cite{grudin1994computer} is representative for the early days of this trend.
+CSCW addresses `how collaborative activities and their coordination can be supported by means of computer systems' \cite{grudin1994computer}, whereas the initial concept evolved towards the more broader field of `Computer-supported collaboration' (CSC).
 
@@
 These developments led to amazing growth in terms of the amounts of content available online and mass participation.
-They are responsible for hundreds of millions of users all over the globe creating high-quality encyclopedias, publishing Terabytes of multimedia content, contributing to world-class software, and lively taking part in defining the agenda of many aspects of our society.
-This trend towards 'prosumerism' is finding more and more adopters in the public and private sector as well, as governments and enterprises not only become active in open initiatives, but encourage the participation of their customers and employees in taking decisions related to organizational management, product development, services offers, and policies.
-In this context, a number of terms are used to refer to the ways people interact with each other and with applications: 'wisdom of the crowds', 'collective intelligence', 'open innovation', 'crowdsourcing', 'human computation', and 'social computing'.
-They are related, but not synonymous with 'social machines', as explained in the following (see also Figure \ref{fig:socialmachine}).
+They are responsible for hundreds of millions of users all over the globe creating high-quality encyclopaedias, publishing Terabytes of multimedia content, contributing to world-class software, and lively taking part in defining the agenda of many aspects of our society.
+This trend towards `prosumerism' is finding more and more adopters in the public and private sector as well, as governments and enterprises not only become active in open initiatives, but encourage the participation of their customers and employees in taking decisions related to organisational management, product development, services offers, and policies.
+In this context, a number of terms are used to refer to the ways people interact with each other and with applications: `wisdom of the crowds', `collective intelligence', `open innovation', `crowdsourcing', `human computation', and `social computing'.
+They are related, but not synonymous with `social machines', as explained in the following (see also Figure \ref{fig:socialmachine}).
 
@@
 Wisdom of the crowds \cite{surowiecki2005wisdom} refers to a principle for decision making that takes into account information and opinions of a group of people rather than individuals; the use of specific technologies, most notably Web $2.0$, has made it possible for such processes to be carried out at scales hardly conceivable in the past, and to involve highly diverse and geographically distributed participants.
-A similar concept, though broader scoped, is collective intelligence, defined in  \cite{malone2009harnessing} as 'groups of individuals doing things collectively that seem intelligent.' The main difference to what we refer to as 'social machines' is hidden in the the automation part of the latter.
+A similar concept, though broader scoped, is collective intelligence, defined in  \cite{malone2009harnessing} as `groups of individuals doing things collectively that seem intelligent.' The main difference to what we refer to as `social machines' is hidden in the the automation part of the latter.
 In a social machine, human and computational intelligence coalesce in order to achieve a given purpose.
@@
 
-One of the direct consequences of the popularity of the wisdom of the crowds idea was a stronger investment worldwide in open innovation, which can be seen as an application of the concept to business environments, or, in the words of the authors, as a 'a paradigm that assumes that firms can and should use external ideas as well as internal ideas, and internal and external paths to market, as the firms look to advance their technology' \cite{chesbrough2003}.
+One of the direct consequences of the popularity of the wisdom of the crowds idea was a stronger investment worldwide in open innovation, which can be seen as an application of the concept to business environments, or, in the words of the authors, as a `a paradigm that assumes that firms can and should use external ideas as well as internal ideas, and internal and external paths to market, as the firms look to advance their technology' \cite{chesbrough2003}.
 At a more general level, this kind of creativity could be leveraged in almost any domain which benefits from diversity, while in the same time ensuring in-time access of a potentially infinite pool of skills and resources not available before the advent of social computing technologies.
-The term crowdsourcing is typically associated with this larger collection of situations, in which 'a job traditionally performed by a designated agent [...] [is outsourced] to an undefined, generally large group of people in the form of an open call' \cite{howe2006crowdsourcing}.
+The term crowdsourcing is typically associated with this larger collection of situations, in which `a job traditionally performed by a designated agent [...] [is outsourced] to an undefined, generally large group of people in the form of an open call' \cite{howe2006crowdsourcing}.
 Human computation applies human processing power to tackle technical tasks that computers (still) find challenging \cite{von2009human}, typically in areas such as visual, audio, and natural language understanding.
-This sort of tasks are an important part of today's crowdsourcing projects landscape, in particular on so-called 'microtask' platforms such as Amazon's Mechanical Turk\footnote{\url{https://www.mturk.com/}} or CrowdFlower\footnote{\url{http://crowdflower.com/}}, which offer small financial rewards to an anonymous crowd engaged with atomic units of work in the range of seconds to minutes to complete.
+This sort of tasks are an important part of today's crowdsourcing projects landscape, in particular on so-called `microtask' platforms such as Amazon's Mechanical Turk\footnote{\url{https://www.mturk.com/}} or CrowdFlower\footnote{\url{http://crowdflower.com/}}, which offer small financial rewards to an anonymous crowd engaged with atomic units of work in the range of seconds to minutes to complete.
 The key difference between social machines and open innovation and crowdsourcing is again in the computational part.
@@
 An analogous line of reasoning could be followed to point out the overlaps between social machines and social computing \cite{parameswaran2007research}.
-The latter is an area of computer science which refers systems that support 'the gathering, representation, processing, use, and dissemination of information that is distributed across social collectivities such as teams, communities, organizations, and markets' \cite{parameswaran2007research}.
-As such, compared to the general concept of 'Computer-supported collaboration', social computing puts a greater emphasis on the information management capabilities of groups and communities, and less on the way these capabilities emerge as a joint effort.
+The latter is an area of computer science which refers systems that support `the gathering, representation, processing, use, and dissemination of information that is distributed across social collectivities such as teams, communities, organizations, and markets' \cite{parameswaran2007research}.
+As such, compared to the general concept of `Computer-supported collaboration', social computing puts a greater emphasis on the information management capabilities of groups and communities, and less on the way these capabilities emerge as a joint effort.
 This distinction is even stronger in the case of social machines, which look at the social and the technical components as equal and necessary partners, and study the ways they could be best combined to master the challenges of future socio-technical systems.
@@
 differentiates between {\it GalaxyZoo}\footnote{In GalaxyZoo people classify galaxies with collective performance as good as professional astronomers, see \url{http://www.galaxyzoo.org}.} and {\it Facebook}, as two prominent example of social machines our classification framework should apply to.
-The user describes the opposing poles of the construct - in this case, the user may decide on 'For Science' and 'For socialising' in order to capture the core distinction between the two system.
+The user describes the opposing poles of the construct - in this case, the user may decide on `For Science' and `For socialising' in order to capture the core distinction between the two system.
 The user then rates every element with value from $1$ to $5$ on this construct, where $1$ represents an element that is purely
-'For Science', and $5$ one that mainly serves  'socializing' purposes.
+`For Science', and $5$ one that mainly serves  `socialising' purposes.
 Elements can also be rated with values between the poles if they may correspond to both or neither.
@@
 \subsection{The polyarchical relationship of social machines}
-When defining the boundaries of what we call social machines, one important observation was related to the distinction between platforms and technologies such as wikis and GWAP ('games-with-a-purpose'), which enable social machines to be created, and their instantiations into social machines that were brought to life for a specific purpose, such as Wikipedia and EteRNA\footnote{\url{http://eterna.cmu.edu}}, a game in which  participants design RNA  that is evaluated automatically using simulators.
+When defining the boundaries of what we call social machines, one important observation was related to the distinction between platforms and technologies such as wikis and GWAP (`games-with-a-purpose'), which enable social machines to be created, and their instantiations into social machines that were brought to life for a specific purpose, such as Wikipedia and EteRNA\footnote{\url{http://eterna.cmu.edu}}, a game in which  participants design RNA  that is evaluated automatically using simulators.
 A second observation is concerned with the relationship between broader and narrower-scoped instances of social machines; general-purpose examples such as Facebook, Twitter or Amazon, or even the Web, enable the formation of more specific social machines and communities within them, that are working to solve solve different problems.
@@
 Looking at the set of social machines in a polyarchy leads to a broad/specific relationship emerging that lets us talk about behavior at various levels of granularity.
-We propose looking at nested machines with 'The Web' as one of several potential roots, with the next level down consisting of sub-platforms (such as Facebook, Twitter, MediaWiki, Ushahidi) that spawn more specific social machines.
+We propose looking at nested machines with `The Web' as one of several potential roots, with the next level down consisting of sub-platforms (such as Facebook, Twitter, MediaWiki, Ushahidi) that spawn more specific social machines.
 A resulting polyarchy is shown in Figure~\ref{polyarchy}, which classifies levels as infrastructure, frameworks, services, and projects and initiatives.
@@
 
-%One such example is the 'open access' movement in scholarly publishing
+%One such example is the `open access' movement in scholarly publishing
 %~\cite{harnad2001self}.
 This cause mobilizes academic authors to self-archive their own papers online so they are available for free, in addition to hosted versions behind publishers pay-walls.
-In order to achieve this aim, the social machine of 'open access' typically uses the social machines of institutional repositories provided
+In order to achieve this aim, the social machine of `open access' typically uses the social machines of institutional repositories provided
 %by the authors' associated institutions. As with other social machines, these social
@@
 appraisal of existing content, posing of problems, solving of
-problems, and so on.  The resulting 'computation', looked at from a
+problems, and so on.  The resulting `computation', looked at from a
 macroscopic whole, might be the identification of high-quality
@@
 %% {\bf The types of tasks undertaken by the social component and the resulting contributions:}Social machines, just as any computing system, are built with a given purpose in mind, and exhibit functionality by which this purpose can be achieved.
-Users and the emerging social network interact with the system and with each other to carry out specific tasks by which they contribute to the 'computation'.
+Users and the emerging social network interact with the system and with each other to carry out specific tasks by which they contribute to the `computation'.
 This includes the creation and refinement of digital artifacts by executing given actions, possibly in collaboration with other participants.
@@
 There is also a dendrogram for the constructs, which can be used, as more data is collected, as an indicator for correlation between constructs.
-In this particular case, dendrogram suggests that there is a correlation between the 'Single Participant Roles' and 'No Hierarchy in Participant Roles' constructs, which is because they refer to the same aspect of social machines.
-Another, but less obvious connection is between the 'Large variety of social features' and 'Benefit NOT for the world/society', which requires further investigation.
+In this particular case, dendrogram suggests that there is a correlation between the `Single Participant Roles' and `No Hierarchy in Participant Roles' constructs, which is because they refer to the same aspect of social machines.
+Another, but less obvious connection is between the `Large variety of social features' and `Benefit NOT for the world/society', which requires further investigation.
 
@@
 %
-% ACM needs 'a single self-contained file'!
+% ACM needs `a single self-contained file'!
 %

### 2013-02-26 latex/paper_new.tex
Correct use of dashes and removed weird double spaces throughout

@@
 The success of the Web has changed this relationship: we now see configurations of people interacting with content and with each other, typified by social Web sites.
-Rather than drawing a line through such Web-based systems to separate the human and digital parts (as computer science has traditionally done), we can now draw a line around them  and  treat  each  such  compound  as  a  `social  machine'  -  a  machine in which the two aspects are seamlessly interwoven.
+Rather than drawing a line through such Web-based systems to separate the human and digital parts (as computer science has traditionally done), we can now draw a line around them and treat each such compound as a `social machine' --- a machine in which the two aspects are seamlessly interwoven.
 This crucial transition in thinking acknowledges the reality of current socio-technical systems, and is essential to underpin any understanding of the science and engineering of their future development towards pervasive ecosystems of co-evolving social machines.
@@
 In repertory grid elicitation a software tool is used to ask users to describe constructs
-that differentiate between elements - for example,  prompting the user to create a construct that
+that differentiate between elements -- for example, prompting the user to create a construct that
 differentiates between {\it GalaxyZoo}\footnote{In GalaxyZoo people classify galaxies with collective performance as good as professional astronomers, see \url{http://www.galaxyzoo.org}.} and {\it Facebook}, as two prominent example of social machines our classification framework should apply to.
-The user describes the opposing poles of the construct - in this case, the user may decide on `For Science' and `For socialising' in order to capture the core distinction between the two system.
+The user describes the opposing poles of the construct -- in this case, the user may decide on `For Science' and `For socialising' in order to capture the core distinction between the two system.
 The user then rates every element with value from $1$ to $5$ on this construct, where $1$ represents an element that is purely
@@
 \subsubsection{Participants and roles}
-This set of constructs pertain to the human participants of social machines and the way(s) participants are organized within them.
-One construct identifies whether participants constitute members of the general public, or are of a specific demographic, external group or organization, occupation, or of a particular expertise.
+This set of constructs pertain to the human participants of social machines and the way(s) participants are organised within them.
+One construct identifies whether participants constitute members of the general public, or are of a specific demographic, external group or organisation, occupation, or of a particular expertise.
 Two constructs, \emph{autonomy} and \emph{anonymity} pertain to the degree to which interaction is constrained among participants, and to the degree to which participants' identities are used and disclosed within the machine.
@@
 society as a whole. Finally, we ask whether extrinsic motivational
-factors also contributed to sustained participation - such factors
+factors also contributed to sustained participation -- such factors
 could include money reward, status, recognition and so on \cite{kuznetsov2006motivations,lakhani2003hackers,tagging}.
@@
 % \subsubsection{Participation (How users interact with the Social Machine)}
-%\subsubsection{Participation - \emph{How and when user participate}}
+%\subsubsection{Participation -- \emph{How and when user participate}}
 %\subsubsection{Problems Solved and Benefit Derived}
@@
 In addition, it enables direct comparisons between systems, and provides a platform for terminologically consistent discussions about the field as a whole and specific instantiations.
-Finally, we also think that with more and more classification data becoming available, for instance, also through functionality that automatically monitors and computes the values of the some of the constructs introduced earlier, researchers will wish to apply prediction techniques (such as SVMs) to social machines, their behavior, and impact.
+Finally, we also think that with more and more classification data becoming available, for instance, also through functionality that automatically monitors and computes the values of the some of the constructs introduced earlier, researchers will wish to apply prediction techniques (such as SVMs) to social machines, their behaviour, and impact.
 
 \section{Future work: building a social machine around the classification framework}
-In this paper we presented a first outline of a possible organization structure for the design space of social machines.
+In this paper we presented a first outline of a possible organisation structure for the design space of social machines.
 We introduced the key features of the notion of social machines, and explained how this notion differs from related areas such as collective intelligence, crowdsourcing and human computing.

### 2013-02-26 latex/paper_new.tex
British English

@@
 % --- Author Metadata here ---
-\conferenceinfo{Workshop on the theory and practice of social machines @ WWW2013}{2013, Rio de Janeiro, Brazil}
+\conferenceinfo{SOCM2013: Workshop on Theory and Practice of Social Machines, WWW2013}{2013, Rio de Janeiro, Brazil}
 %\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
@@
 We anticipate that the framework will be a useful tool for both researchers in social and computer sciences, and for developers and operators of social machines.
-Using a common conceptualization allows the former to become familiar with the landscape, and identify topics of research that so far have remained unexplored or under-explored.
-They are given a theoretical grounding to observe the effects of specific technical properties of a system configuration on social behavior, discover design and evolution patterns of such systems, learn about social network formation and dynamics, and devise incentive mechanisms to encourage wide participation.
+Using a common conceptualisation allows the former to become familiar with the landscape, and identify topics of research that so far have remained unexplored or under-explored.
+They are given a theoretical grounding to observe the effects of specific technical properties of a system configuration on social behaviour, discover design and evolution patterns of such systems, learn about social network formation and dynamics, and devise incentive mechanisms to encourage wide participation.
 For developers and operators, the framework may inform the engineering of new systems in terms of critical features and community development.
@@
 
-The remaining sections are organized as follows.
+The remaining sections are organised as follows.
 In order to pin down what makes up a social machine, and establish the boundaries of this emerging field of research, we need to first understand the relationship between social machines and related topics such as human computation, collective intelligence, and crowdsourcing; Section \ref{sec:comparison} gives a summary of this comparative analysis.
@@
 In a social machine, human and computational intelligence coalesce in order to achieve a given purpose.
-The overall setting may slightly vary in terms of the interaction patterns between the two types of services and the way their outcomes are used; however, where wisdom of the crowds and collective intelligence clearly place their focus on identifying the situations in which groups of people perform better than individuals, social machines are concerned with the study and realization of hybrid systems where the two types of components co-exist.
+The overall setting may slightly vary in terms of the interaction patterns between the two types of services and the way their outcomes are used; however, where wisdom of the crowds and collective intelligence clearly place their focus on identifying the situations in which groups of people perform better than individuals, social machines are concerned with the study and realisation of hybrid systems where the two types of components co-exist.
 As such, theoretical and empirical insights from the two related areas are useful to understand the dynamics of the social structures underlying such a system, but they are definitely not the only ingredient needed to build operational social machines.
@@
 An analogous line of reasoning could be followed to point out the overlaps between social machines and social computing \cite{parameswaran2007research}.
-The latter is an area of computer science which refers systems that support `the gathering, representation, processing, use, and dissemination of information that is distributed across social collectivities such as teams, communities, organizations, and markets' \cite{parameswaran2007research}.
+The latter is an area of computer science which refers systems that support `the gathering, representation, processing, use, and dissemination of information that is distributed across social collectivities such as teams, communities, organisations, and markets' \cite{parameswaran2007research}.
 As such, compared to the general concept of `Computer-supported collaboration', social computing puts a greater emphasis on the information management capabilities of groups and communities, and less on the way these capabilities emerge as a joint effort.
@@
 The aim of this exercise was twofold: to remove any redundancy and cut the classification space down to a manageable size, while ensuring that all constructs that were elicited were represented in the final set.
-This process involved four of the authors discussing the choices, and resulted in a consolidated set of $31$ constructs organized in four different clusters: popularity, tasks and purpose, participants and roles, and motivation and incentives.
+This process involved four of the authors discussing the choices, and resulted in a consolidated set of $31$ constructs organised in four different clusters: popularity, tasks and purpose, participants and roles, and motivation and incentives.
 In our analysis we will not further consider the first cluster, which primarily covered descriptive information about a given system, such as its perceived level of maturity and current number of users.
-Further on, the constructs were not evenly distributed across clusters; whereas some aspects such as roles of participants occurred frequently (in various flavors) in the grids provided by the ten experts, others such as the different types of workflows synthesizing human and automatized capabilities of a social machine were less taken into account.
+Further on, the constructs were not evenly distributed across clusters; whereas some aspects such as roles of participants occurred frequently (in various flavours) in the grids provided by the ten experts, others such as the different types of workflows synthesising human and automatised capabilities of a social machine were less taken into account.
 As discussed in Section \ref{sec:comparison} we believe that the interaction between these two types of services is an essential part of the theory and practice of social machines, being one of the elements that distinguishes them from related areas which are biased towards one or the other.
@@
 For example, insights on the classification of space objects gained through GalaxyZoo may influence the content of the corresponding articles, and the associated editorial processes, within Wikipedia.
-In the same time,a large number of communities within Wikipedia focus specific topics and activities, such as those contributing to science and technology.\footnote{Visualization of Wikipedia Science Communities: \url{http://www.olihb.com/WikiCommunities/}}
+In the same time,a large number of communities within Wikipedia focus specific topics and activities, such as those contributing to science and technology.\footnote{Visualisation of Wikipedia Science Communities: \url{http://www.olihb.com/WikiCommunities/}}
 
-Looking at the set of social machines in a polyarchy leads to a broad/specific relationship emerging that lets us talk about behavior at various levels of granularity.
+Looking at the set of social machines in a polyarchy leads to a broad/specific relationship emerging that lets us talk about behaviour at various levels of granularity.
 We propose looking at nested machines with `The Web' as one of several potential roots, with the next level down consisting of sub-platforms (such as Facebook, Twitter, MediaWiki, Ushahidi) that spawn more specific social machines.
@@
 The constructs apply to all levels of the polyarchy, though some of them might be more easier to understand and describe in the context of some levels.
-For example, instances of social machines residing at the bottom of the polyarchy are likely to be built with a very concrete audience in mind, and as such the mechanisms to incentivize participation are likely to be clearer and more straightforward to study and adjust than in cases in which such boundaries cannot be defined, such as Facebook, or even Wikipedia.
+For example, instances of social machines residing at the bottom of the polyarchy are likely to be built with a very concrete audience in mind, and as such the mechanisms to incentivise participation are likely to be clearer and more straightforward to study and adjust than in cases in which such boundaries cannot be defined, such as Facebook, or even Wikipedia.
 In the same time, these systems will have to decide among an array of diverse basic services and frameworks for their implementations; the impact of such engineering decisions is difficult to predict in great detail as a thorough understanding of what makes certain social machines more successful than others, in terms of technology choices and beyond, is largely missing for most application domains.
@@
 %~\cite{harnad2001self}.
-This cause mobilizes academic authors to self-archive their own papers online so they are available for free, in addition to hosted versions behind publishers pay-walls.
+This cause mobilises academic authors to self-archive their own papers online so they are available for free, in addition to hosted versions behind publishers pay-walls.
 In order to achieve this aim, the social machine of `open access' typically uses the social machines of institutional repositories provided
@@
 Variation in types of contributions and tasks \\
-Participants' participate to fulfill needs of a role \\
+Participants' participate to fulfil needs of a role \\
 \hspace{1cm} work-related/professional role \\
@@
 Participant anonymity \\
-Extent of hierarchical organization of roles \\
+Extent of hierarchical organisation of roles \\
 Clear separation of roles among participants \\
@@
 Users and the emerging social network interact with the system and with each other to carry out specific tasks by which they contribute to the `computation'.
-This includes the creation and refinement of digital artifacts by executing given actions, possibly in collaboration with other participants.
+This includes the creation and refinement of digital artefacts by executing given actions, possibly in collaboration with other participants.
 For example, in GalaxyZoo\footnote{\url{http://www.galaxyzoo.org}} people classify galaxies with collective performance as good as professional astronomers, while in reCaptcha\footnote{\url{http://www.google.com/recaptcha}} they refine the data that describes digital manuscripts.
@@
 instance, it is clear that YouTube, Vimeo, Reddit, and Digg are similar, which is not unexpected, due to their focus on sharing content.
-Likewise, the exercise confirms commonalities between Quora and StackOverflow, which can be easily understood, given their question answering purpose.
+Likewise, the exercise confirms commonalities between Quora and Stack Overflow, which can be easily understood, given their question answering purpose.
 There is also a dendrogram for the constructs, which can be used, as more data is collected, as an indicator for correlation between constructs.

### 2013-02-26 latex/paper_new.tex
Abstract from dan

@@
 
-social machines, overview of what they are, their history
-    people interacting with content and with each other
+The state of the art in human interaction with computation systems blurs the line between
+computation performed by machine logic and algorithms, and those that result from input by
+humans, arising from their own neurological processes and life experience. Current
+socio-technical systems, known as `social machines' exploit the large-scale interaction of
+humans with machines, motivated for numerous aims and purposes across the spectrums of
+financial gain, charitable aid and simply for fun. In this paper we explore
+the landscape of social machines, both past and present, towards the aim of defining an
+initial classification framework. Through a number of knowledge elicitation
+and refinement exercises we have identified the polyarchical relationship between
+infrastructure, social machines and large-scale social initiatives. Our initial framework
+describes classification constructs in the areas of {\it participation}, {\it participants} and
+{\it motivation}. We present an initial classification of the largest social machines, as
+an initial validation and demonstration of the use of our identified constructs.
+There is rich potential for performing
+analysis on the behaviour and phenomenology of social machines, and analysis of their
+growth and outputs over time. Our future work is to elicit additional opinion,
+classifications and validation from a wider audience, to produce a formal framework to
+classify and aid in analysis of social machines.
 

### 2013-02-26 latex/paper_new.tex
Added missing hline in table

@@
 Clear separation of roles among participants \\
+\hline
 {\bf Motivation and incentives} \\

### 2013-03-05 latex/paper_new.tex
new image and final reading

--- 2013-02-26 latex/paper_new.tex
@@
 \alignauthor
-Authors\\
+Nigel Shadbolt,^{1} Daniel Alexander Smith,^{2} Elena Simperl,^{3} Max Van Kleek,^{4} Yang Yang,^{5} Wendy Hall^{6}
        \affaddr{Web and Internet Science Group}\\
@@
        \affaddr{Southampton, UK}\\
-       \email{\{a,b,c,d,wh,nrs\}@ecs.soton.ac.uk}
+       \email{\{^{1}nrs, ^{2}ds, ^{3}es, ^{4}emax, ^{5}yy, ^{6}wh@ecs.soton.ac.uk}
 }
+
 % There's nothing stopping you putting the seventh, eighth, etc.
@@
 
-social machines, overview of what they are, their history
-    people interacting with content and with each other
+The state of the art in human interaction with computational systems blurs the line between
+computations performed by machine logic and algorithms, and those that result from input by
+humans, arising from their own psychological processes and life experience. Current
+socio-technical systems, known as `social machines' exploit the large-scale interaction of
+humans with machines. Interactions that are motivated by numerous goals and purposes
+including financial gain, charitable aid, and simply for fun. In this paper we explore
+the landscape of social machines, both past and present, with the aim of defining an
+initial classificatory framework. Through a number of knowledge elicitation
+and refinement exercises we have identified the polyarchical relationship between
+infrastructure, social machines, and large-scale social initiatives. Our initial framework
+describes classification constructs in the areas of {\it contributions}, {\it participants}, and
+{\it motivation}. We present an initial characterization of some of the most popular social machines, as
+an initial validation and demonstration of the use of the identified constructs.
+We believe that it is important to undertake an analysis of the behaviour and phenomenology of social machines, and of their growth and evolution over time.
+Our future work will seek to elicit additional opinions, classifications and validation from a wider audience, to produce a comprehensive framework for the description, analysis and comparison of social machines.
 
@@
 Rather than drawing a line through such Web-based systems to separate the human and digital parts (as computer science has traditionally done), we can now draw a line around them and treat each such compound as a `social machine' --- a machine in which the two aspects are seamlessly interwoven.
+This was the insight behind Berners-Lee's original characterisation of the concept of a social machine \cite{bernerslee1999}.
 This crucial transition in thinking acknowledges the reality of current socio-technical systems, and is essential to underpin any understanding of the science and engineering of their future development towards pervasive ecosystems of co-evolving social machines.
@@
 A traditional database, with users looking up records independently of each other and an administrator responsible for the management of the content, has some of the right ingredients, but there is really no social element in this strict provider-consumer relationship.
-When we fill out a form (e.g., health information, birds spotted in my garden, new construction sites on my way to work) the systems supporting this activity are minimal forms of social machines because the users are part of the `social computation', in this case the data creation and collection process facilitate through the site.
-This social component becomes richer when the database is curated by members of the broader community (e.g., Wikipedia) and when the social network adds value implicitly (e.g., Amazon) or explicitly (e.g., Facebook) to the overall system through their individual or joint activities.
+When we fill out a form (e.g., health information, birds spotted in a garden, new construction sites on the way to work) the systems supporting this activity are minimal forms of social machines because the users are part of the `social computation', in this case the data creation and collection process facilitated through the site.
+This social component becomes richer when the database is curated by members of the broader community (e.g., Wikipedia) and when the social network adds value implicitly (e.g., Amazon) or explicitly (e.g., Facebook) to the overall system through individual or joint activities of the participants.
 
-A number of prominent examples aside, today we still see a divide between conventional IT systems dedicated to data- and computation-intensive tasks, and Web $2.0$ sites offering some combination of well-known participatory features, in which user-generated content and the underlying social network evolve dynamically and hand-in-hand.
-However, as technology becomes more and more ubiquitous, many of the challenges we witness in any domain of our life, business, and society will soon require solutions that rely on both of these axes: a sophisticated combination of data-intensive, complex automation and deep community involvement.
-This suggests the need for new types of systems to tackle these emerging challenges, and these systems will not be able to be built and used in a sustainable way without a through understanding of the science and engineering of (the continuum of) social machines.
+In general we still see a divide between conventional IT systems dedicated to data- and computation-intensive tasks, and Web $2.0$ sites offering some combination of well-known participatory features, in which user-generated content and the underlying social network evolve dynamically and hand-in-hand.
+However, as technology becomes more and more ubiquitous, many of the challenges we face will increasingly require solutions that rely on both of these axes: a sophisticated combination of data-intensive, complex automation, and deep community involvement.
+This suggests the need for new types of systems to tackle these emerging challenges, and these systems will not be able to be built and used in a sustainable way without a thorough understanding of the science and engineering of (the continuum of) social machines.
 
-As an early step towards achieving this principled understanding, we propose in this paper an initial outline for a classification framework for social machines.
-The primary aim of the framework is to identify and define the constructs to describe, study, and compare this expanding field of interdisciplinary research, as well as the increasing number of systems combining human and computational intelligence that are continuously being created in areas as diverse as traffic management, healthcare, media and journalism, or design and natural sciences.
+As an early step towards achieving this principled understanding, we propose in this paper a first outline for a classification framework for social machines.
+The primary aim of the framework is to identify and define the constructs to describe, study, and compare this expanding field of interdisciplinary research.
+We seek to better describe the increasing number of systems combining human and computational intelligence --- systems that are continuously being created in areas as diverse as traffic management, healthcare, media and journalism, or design and natural sciences.
 
 The current version of our framework is the result of a knowledge elicitation process performed internally in our lab.
-Our goal is to present it to the broader community and extend and revise it based on their feedback in order to achieve a shared understanding of the basic notions and of the associated terminology, and to consensually lay out the structure of the design space that will enable a systematic study of existing and future social machines.
-We anticipate that the framework will be a useful tool for both researchers in social and computer sciences, and for developers and operators of social machines.
+Our goal is to present it to the broader community and then to extend and revise it based on feedback in order to achieve a shared understanding of the basic notions and of the associated terminology.
+We aim to develop a descriptive framework to enable a systematic study of existing and future social machines.
+We anticipate that the framework will be a useful tool for both researchers in social and computer sciences, and for developers and operators of such systems.
 Using a common conceptualisation allows the former to become familiar with the landscape, and identify topics of research that so far have remained unexplored or under-explored.
-They are given a theoretical grounding to observe the effects of specific technical properties of a system configuration on social behaviour, discover design and evolution patterns of such systems, learn about social network formation and dynamics, and devise incentive mechanisms to encourage wide participation.
+They are given a means to observe the effects of specific technical properties of a system configuration on social behaviour, discover design and evolution patterns, learn about social network formation and dynamics, and devise incentive mechanisms to encourage wide participation.
 For developers and operators, the framework may inform the engineering of new systems in terms of critical features and community development.
 
-This paper makes two main contributions: First, we introduce the notion of `social machines' and position it in connection to similar ideas and initiatives.
+This paper makes two main contributions: First, we introduce the notion of `social machines' and how they relate to other concepts.
 Second we propose a set of constructs to capture the most important features of social machines and to compare existing instances thereof.
@@
 The remaining sections are organised as follows.
-In order to pin down what makes up a social machine, and establish the boundaries of this emerging field of research, we need to first understand the relationship between social machines and related topics such as human computation, collective intelligence, and crowdsourcing; Section \ref{sec:comparison} gives a summary of this comparative analysis.
-In Section \ref{sec:framework} we then present the framework in terms of the methodology used to devise it, the main clusters of constructs used for classification, and an example of using them to compare a selection of popular social machines via grid analysis.
-We conclude with an outlook of our future development activities, including details about the evaluation of the framework.
-In addition, we plan to engage with the broader community to set up a `social machine' bringing together researchers from various disciplines who will contribute to the `computation' of a joint and widely agreed classification framework for the field.
+In order to pin down what constitutes a social machine, and establish the boundaries of this emerging field of research, we need to first understand the relationship between social machines and related topics such as human computation, collective intelligence, and crowdsourcing; Section \ref{sec:comparison} provides a summary of this comparative analysis.
+In Section \ref{sec:framework} we present the framework in terms of the methodology used to devise it, the main clusters of constructs used for classification, and an example of using them to compare a selection of popular social machines via repertory grid analysis.
+We conclude with our proposed future work, including details concerning the evaluation of the framework and reference our plans to engage with the broader community to set up a `social machine' bringing together researchers from various disciplines to help formulate a widely agreed framework for the field.
 
@@
 \label{sec:comparison}
-Since the invention of the Turing machine we have experienced a `paradigm shift' in the usage of computers, moving away from purely calculative devices to facilitators of a wide range of human interactions.
-The emergence of `Computer-supported Collaboration Work (CSCW)' \cite{grudin1994computer} is representative for the early days of this trend.
-CSCW addresses `how collaborative activities and their coordination can be supported by means of computer systems' \cite{grudin1994computer}, whereas the initial concept evolved towards the more broader field of `Computer-supported collaboration' (CSC).
+Turing machines were once conceived as logical automata.
+We have experienced a `paradigm shift' in the usage of computers, moving away from this narrow characterisation to one according to which machines facilitate a wide range of human interactions.
+The emergence of `Computer-supported Collaborative Work (CSCW)' \cite{grudin1994computer} is representative of the early phases of this trend.
+This initial concept has evolved to become the broader field of `Computer-supported collaboration' (CSC).
 
 The Web as a global platform for information access and sharing marked a second essential milestone; in particular, through principles and technologies promoted through Web $2.0$ and the Mobile Web.
-These developments led to amazing growth in terms of the amounts of content available online and mass participation.
-They are responsible for hundreds of millions of users all over the globe creating high-quality encyclopaedias, publishing Terabytes of multimedia content, contributing to world-class software, and lively taking part in defining the agenda of many aspects of our society.
-This trend towards `prosumerism' is finding more and more adopters in the public and private sector as well, as governments and enterprises not only become active in open initiatives, but encourage the participation of their customers and employees in taking decisions related to organisational management, product development, services offers, and policies.
+These developments led to amazing growth in terms of the amounts of content available online and the extent of mass participation.
+They are responsible for hundreds of millions of users all over the globe creating high-quality encyclopedias, publishing Terabytes of multimedia content, contributing to world-class software, and active participation in defining the agenda for many aspects of our society.
+This trend towards `prosumerism' is finding more and more adopters in the public and private sector.
+Governments and enterprises are not only becoming active in open initiatives, but encourage the participation of their customers and employees in taking decisions related to organisational management, product development, service offerings, and policy formulation.
 In this context, a number of terms are used to refer to the ways people interact with each other and with applications: `wisdom of the crowds', `collective intelligence', `open innovation', `crowdsourcing', `human computation', and `social computing'.
-They are related, but not synonymous with `social machines', as explained in the following (see also Figure \ref{fig:socialmachine}).
+As we argue below, these terms are related, but not synonymous with `social machines' (see also Figure \ref{fig:socialmachine}).
 
@@
 \begin{center}
-\includegraphics[width=0.25\textwidth]{img/socialmachinescope.png}
+\includegraphics[width=0.5\textwidth]{img/socialmachinesscope_new.png}
 \caption{Social machines and related areas} \label{fig:socialmachine}
@@
 
-Wisdom of the crowds \cite{surowiecki2005wisdom} refers to a principle for decision making that takes into account information and opinions of a group of people rather than individuals; the use of specific technologies, most notably Web $2.0$, has made it possible for such processes to be carried out at scales hardly conceivable in the past, and to involve highly diverse and geographically distributed participants.
-A similar concept, though broader scoped, is collective intelligence, defined in \cite{malone2009harnessing} as `groups of individuals doing things collectively that seem intelligent.' The main difference to what we refer to as `social machines' is hidden in the the automation part of the latter.
+Wisdom of the crowds \cite{surowiecki2005wisdom} refers to a principle for decision making that takes into account the information and opinions of a group of people rather than individuals; the use of specific technologies, most notably Web $2.0$, has made it possible for such processes to be carried out at scales inconceivable in the past, and to involve highly diverse and geographically distributed participants.
+A similar concept, though broader scoped, is collective intelligence, defined in \cite{malone2009harnessing} as `groups of individuals doing things collectively that seem intelligent.' The main difference to what we refer to as `social machines' is with regard to the extent and role of automation.
 In a social machine, human and computational intelligence coalesce in order to achieve a given purpose.
-The overall setting may slightly vary in terms of the interaction patterns between the two types of services and the way their outcomes are used; however, where wisdom of the crowds and collective intelligence clearly place their focus on identifying the situations in which groups of people perform better than individuals, social machines are concerned with the study and realisation of hybrid systems where the two types of components co-exist.
-As such, theoretical and empirical insights from the two related areas are useful to understand the dynamics of the social structures underlying such a system, but they are definitely not the only ingredient needed to build operational social machines.
-The question of how human and automatic services can be brought together to achieve optimal results, as well as the actual engineering exercise by which a system is developed, tested, and updated are equally important.
+While wisdom of the crowds and collective intelligence place their focus on identifying the situations in which groups of people perform better than individuals, social machines are concerned with the study and realisation of hybrid systems in which the two types of components co-exist.
+As such, theoretical and empirical insights from these two related areas are useful to understand the dynamics of the social structures underlying a social machine, but they are definitely not the only ingredient needed to build operational systems.
+The question of how human and automatic services can be brought together to achieve optimal results, as well as the actual engineering by which a system is developed, tested, and updated are equally important.
 
 One of the direct consequences of the popularity of the wisdom of the crowds idea was a stronger investment worldwide in open innovation, which can be seen as an application of the concept to business environments, or, in the words of the authors, as a `a paradigm that assumes that firms can and should use external ideas as well as internal ideas, and internal and external paths to market, as the firms look to advance their technology' \cite{chesbrough2003}.
-At a more general level, this kind of creativity could be leveraged in almost any domain which benefits from diversity, while in the same time ensuring in-time access of a potentially infinite pool of skills and resources not available before the advent of social computing technologies.
+At a more general level, this kind of creativity could be leveraged in almost any domain which benefits from diversity, whilst at the same time ensuring in-time access to a potentially infinite pool of skills and resources not available before the advent of social computing technologies.
 The term crowdsourcing is typically associated with this larger collection of situations, in which `a job traditionally performed by a designated agent [...] [is outsourced] to an undefined, generally large group of people in the form of an open call' \cite{howe2006crowdsourcing}.
 Human computation applies human processing power to tackle technical tasks that computers (still) find challenging \cite{von2009human}, typically in areas such as visual, audio, and natural language understanding.
-This sort of tasks are an important part of today's crowdsourcing projects landscape, in particular on so-called `microtask' platforms such as Amazon's Mechanical Turk\footnote{\url{https://www.mturk.com/}} or CrowdFlower\footnote{\url{http://crowdflower.com/}}, which offer small financial rewards to an anonymous crowd engaged with atomic units of work in the range of seconds to minutes to complete.
-The key difference between social machines and open innovation and crowdsourcing is again in the computational part.
-The novelty these two approaches bring in lies in using a much larger pool of human resources as traditional work environments \cite{quinn2011human}.
+These kinds of task are an important part of today's crowdsourcing landscape, in particular on so-called `microtask' platforms such as Amazon's Mechanical Turk\footnote{\url{https://www.mturk.com/}} or CrowdFlower\footnote{\url{http://crowdflower.com/}}, which offer small financial rewards to an anonymous crowd engaged with atomic units of work that take in the range of seconds to minutes to complete.
+The key difference between social machines and open innovation and crowdsourcing is at the level of the interaction between the social and the machine-driven processing components.
+The novelty these two other approaches bring in lies in their use of a much larger pool of human resources than traditional work environments \cite{quinn2011human}.
 There are clear socio-economic implications that their adopters need to deal with in order to optimally make use of this wealth of resources; technology may be needed to assist specific aspects of crowdsourcing projects, from evaluating and rewarding the results produced by the crowd to consolidating and aggregating them into a complete solution.
-A principled combination of human and computational capabilities is less in the focus of such systems and the technical means that support them, by contrast to social machines.
+These aspects are equally important for social machines, which, however, also look into the principled combination of human and computational capabilities and the technical means to support them.
 In comparison to human computation, social machines cover a wider range of scenarios.
@@
 
-An analogous line of reasoning could be followed to point out the overlaps between social machines and social computing \cite{parameswaran2007research}.
-The latter is an area of computer science which refers systems that support `the gathering, representation, processing, use, and dissemination of information that is distributed across social collectivities such as teams, communities, organisations, and markets' \cite{parameswaran2007research}.
+An analogous line of reasoning illustrates the overlap between social machines and social computing \cite{parameswaran2007research}.
+The latter is an area of computer science that refers to systems that support `the gathering, representation, processing, use, and dissemination of information that is distributed across social collectivities such as teams, communities, organisations, and markets' \cite{parameswaran2007research}.
 As such, compared to the general concept of `Computer-supported collaboration', social computing puts a greater emphasis on the information management capabilities of groups and communities, and less on the way these capabilities emerge as a joint effort.
-This distinction is even stronger in the case of social machines, which look at the social and the technical components as equal and necessary partners, and study the ways they could be best combined to master the challenges of future socio-technical systems.
+This distinction is even stronger in the case of social machines, which regard the social and the technical components as equal and necessary partners, and study the ways they could be best combined to master the challenges of future socio-technical systems.
 
@@
 technique~\cite{kelly} in order to derive an initial set of {\it elements}, which represent instances of social machines, and {\it constructs}, which capture their most important characteristics.
+The technique originated in psychology where it was used to elicit the implicit knowledge used by human subjects to conceptualise the world around them.
+
 In repertory grid elicitation a software tool is used to ask users to describe constructs
-that differentiate between elements -- for example, prompting the user to create a construct that
+that differentiate a set of elements -- for example, prompting the user to create a construct that
 differentiates between {\it GalaxyZoo}\footnote{In GalaxyZoo people classify galaxies with collective performance as good as professional astronomers, see \url{http://www.galaxyzoo.org}.} and {\it Facebook}, as two prominent example of social machines our classification framework should apply to.
-The user describes the opposing poles of the construct -- in this case, the user may decide on `For Science' and `For socialising' in order to capture the core distinction between the two system.
-The user then rates every element with value from $1$ to $5$ on this construct, where $1$ represents an element that is purely
-`For Science', and $5$ one that mainly serves `socialising' purposes.
-Elements can also be rated with values between the poles if they may correspond to both or neither.
-The repertory grid software then identifies elements that are the least differentiated when requesting new constructs, in an attempt to elicit as much knowledge as possible from the user.
-We applied this technique to create an initial set of examples of social machines and to classify them according to specific features.
-To do so, we asked then computer science researchers familiar with the field to create their own repertory grids, and populate the elements from their own knowledge, and created the constructs using the standard repertory grid elicitation technique described above.
+The user describes the opposing poles of the construct --- in this case, the user may decide on `For Science' and `For socialising' in order to capture the core distinction between the two system.
+The user then rates every element with value from $1$ to $5$ on this construct, where $1$ represents an element that is purely `For Science', and $5$ one that mainly serves `socialising' purposes.
+In triadic presentation the repertory grid software lists three elements and asks the user to generate a construct where one of the elements is different to the other two.
+We asked computer science researchers familiar with the field to create their own repertory grids, generating elements from their own knowledge, and creating constructs using this technique.
 This exercise led to $10$ grids, the union of which comprised a total of $56$ unique elements (social machines)
 and $117$ different constructs (classifying factors).
-As the aim of this initial phase was to understand how people perceive the notion of social machine and their most distinctive properties, we allowed the participants to choose the social machines they are familiar with and describe them in their own terms.
+As the aim of this initial phase was to understand how people perceive the notion of a social machine and their most distinctive properties, we allowed the participants to choose the systems they are familiar with and describe them in their own terms.
 
@@
 We manually grouped the constructs into rough clusters, based around the areas they cover.
-We examined each construct to determine which were equivalent, and whether we could re-word or subsume existing constructs to cover the same aspects.
-The aim of this exercise was twofold: to remove any redundancy and cut the classification space down to a manageable size, while ensuring that all constructs that were elicited were represented in the final set.
+We examined each construct to determine which were equivalent, and whether we could re-word or subsume existing constructs to reduce redundancies and overlap.
+Our aim was to cut the classification space down to a manageable size, while ensuring that all constructs that were initially elicited were represented in the final set.
 This process involved four of the authors discussing the choices, and resulted in a consolidated set of $31$ constructs organised in four different clusters: popularity, tasks and purpose, participants and roles, and motivation and incentives.
-In our analysis we will not further consider the first cluster, which primarily covered descriptive information about a given system, such as its perceived level of maturity and current number of users.
-Further on, the constructs were not evenly distributed across clusters; whereas some aspects such as roles of participants occurred frequently (in various flavours) in the grids provided by the ten experts, others such as the different types of workflows synthesising human and automatised capabilities of a social machine were less taken into account.
+In our analysis we did not further consider the first cluster, which primarily covered descriptive information about a given system, such as its perceived level of maturity and current number of users.
+It was apparent that the constructs were not evenly distributed across clusters; whereas some aspects such as roles of participants occurred frequently (in various flavours) in the grids provided by the ten experts, others such as the different types of workflows synthesising human and automatised capabilities of a social machine were less common.
 As discussed in Section \ref{sec:comparison} we believe that the interaction between these two types of services is an essential part of the theory and practice of social machines, being one of the elements that distinguishes them from related areas which are biased towards one or the other.
@@
 The paper presents the current version of the framework.
-We envision a reiteration of the main clusters and constructs based on discussions with and feedback from various scientific audiences.
+We envision iterating the main clusters and constructs based on discussions with and feedback from various scientific audiences.
 We provide initial evidence of the usability of the framework as a tool to examine commonalities and differences between social machines in Section \ref{sec:usage}, in which we applied the repertory grid elicitation technique to a set of $20$ social machines instances ranked by popularity according to the Alexa service.\footnote{\url{www.alexa.com/}} A more thorough evaluation will determine the extent to which the classification constructs identified can be meaningfully used by researchers and system designers and operators.
@@
 For example, insights on the classification of space objects gained through GalaxyZoo may influence the content of the corresponding articles, and the associated editorial processes, within Wikipedia.
-In the same time,a large number of communities within Wikipedia focus specific topics and activities, such as those contributing to science and technology.\footnote{Visualisation of Wikipedia Science Communities: \url{http://www.olihb.com/WikiCommunities/}}
+At the same time,a large number of communities within Wikipedia focus specific topics and activities, such as those contributing to science and technology.\footnote{Visualisation of Wikipedia Science Communities: \url{http://www.olihb.com/WikiCommunities/}}
 
@@
 This approach enables us to start with a more detailed analysis of certain levels over others;
-and seeing what similarities flow up and down the polyarchy. For example, what do
+and to see what similarities flow up and down the polyarchy. For example, what do
 specific instances of Ushahidi/Zooniverse/MediaWiki have in common with other instances, and
-how to they differ dynamically?
+how do they differ dynamically?
 How do certain design decisions taken at the level of the infrastructure, frameworks and service propagate into narrower-focused systems that are built on top of them?
@@
 For example, instances of social machines residing at the bottom of the polyarchy are likely to be built with a very concrete audience in mind, and as such the mechanisms to incentivise participation are likely to be clearer and more straightforward to study and adjust than in cases in which such boundaries cannot be defined, such as Facebook, or even Wikipedia.
-In the same time, these systems will have to decide among an array of diverse basic services and frameworks for their implementations; the impact of such engineering decisions is difficult to predict in great detail as a thorough understanding of what makes certain social machines more successful than others, in terms of technology choices and beyond, is largely missing for most application domains.
+At the same time, these systems will have to decide among an array of diverse basic services and frameworks for their implementations; the impact of such engineering decisions is difficult to predict in great detail as a thorough understanding of what makes certain social machines more successful than others, in terms of technology choices and beyond, is largely missing for most application domains.
 
@@
 elimination of over-represented areas) and the identification of three
-simple themes: a
-
-sets of constructs pertaining to users and roles within
-each machine, second, the nature of tasks, problems and activities
-performed through and using the machine, third, motivation for
-participation, finally, the contexts of participation.
+themes: first the nature of tasks, problems and activities
+performed through and using the machine, second the users and roles within
+each machine, and third, motivation for participation.
 
@@
 \hspace{1cm} to ``get something done'' \\
-\hspace{1cm} to ``for fun/entertainment''\\
+\hspace{1cm} to ``be for fun/entertainment''\\
 \hspace{1cm} to ``be social'' \\
@@
 encompass a general class of activities. To capture this diversity,
-we include a construct that differentiates machines with pre-defined problem spaces or are open to participant-provided tasks and activities. Finally, we ask whether the benefit derived by
+we include a construct that differentiates machines with pre-defined problem spaces or those that are open to participant-provided tasks and activities. Finally, we ask whether the benefit derived by
 participants is distinct from that gained by machine/platform owners
@@
 This set of constructs pertain to the human participants of social machines and the way(s) participants are organised within them.
-One construct identifies whether participants constitute members of the general public, or are of a specific demographic, external group or organisation, occupation, or of a particular expertise.
-Two constructs, \emph{autonomy} and \emph{anonymity} pertain to the degree to which interaction is constrained among participants, and to the degree to which participants' identities are used and disclosed within the machine.
-Additional two constructs pertain to the roles: first, as to whether multiple roles are pre-differentiated by the machine or
+One construct identifies whether participants constitute members of the general public, or are of a specific demographic, external group or organisation, occupation, or possess a particular expertise.
+Two constructs, \emph{autonomy} and \emph{anonymity} pertain to the degree to which interaction is constrained among participants, and the degree to which participants' identities are used and disclosed within the machine.
+An additional two constructs pertain to the roles: first, as to whether multiple roles are pre-differentiated by the machine or
 whether all participants essentially initially assume the same role, second whether roles (pre-defined or emergent) are hierarchical.
@@
 \subsubsection{Motivation and incentives}
-Related to the social structure motivation that sustains
+These constructs relate to the social structure and motivation that sustains
 continued participation in these systems. We identify six different
@@
 participation is fun, second that it accomplishes an activity that the
-participant enjoys or wanted to get done, third, that it satisfies the
-desire to gain or share knowledge, or, fourth, or to be social. The
+participant enjoys or wants to get done, third, that it satisfies the
+desire to gain or share knowledge, or, fourth, the desire to be social. The
 last two pertain to philanthropy, whether participation is seen to
@@
 by their Alexa ranking, and performed a full repertory grid elicitation exercise over the first $20$ of the $56$ collected in the first phase of the methodology (see Section \ref{sec:methodology}).
-This process assessing each social machine against the constructs by choosing a rate between $1$ and $5$.
+This process assessing each social machine against the constructs by choosing a rating between $1$ and $5$.
 This led to a total of $680$ ratings, and the
@@
 \begin{center}
-\includegraphics[width=0.2\textwidth]{img/dendrogram-elements.png}
-\caption{Dendrogram of elements (social machines) from a repertory grid exercise of the top $20$ (ranked by Alexa) social machines from our element set, against our consolidated constructs} \label{dendrogram}
+\includegraphics[width=0.5\textwidth]{img/dendrogram-both.png}
+\caption{Dendrograms of constructs, in blue, and elements (social machines), in red, from a repertory grid exercise of the top $20$ (ranked by Alexa) social machines from our element set, against our consolidated constructs.} \label{dendrogram}
 \end{center}
@@
 
-\begin{figure}[htb]
-\begin{center}
-\includegraphics[width=0.4\textwidth]{img/dendrogram-constructs.png}
-\caption{Dendrogram of constructs from a repertory grid exercise of the top $20$ (ranked by Alexa) social machines from our element set, against our consolidated constructs} \label{dendrogram}
-\end{center}
-\end{figure}
+%\begin{figure}[htb]
+%\begin{center}
+%\includegraphics[width=0.2\textwidth]{img/dendrogram-elements.png}
+%\caption{Dendrogram of elements (social machines) from a repertory grid exercise of the top $20$ (ranked by Alexa) social machines from our element set, against our consolidated constructs} \label{dendrogram}
+%\end{center}
+%\end{figure}
+%
+%\begin{figure}[htb]
+%\begin{center}
+%\includegraphics[width=0.4\textwidth]{img/dendrogram-constructs.png}
+%\caption{Dendrogram of constructs from a repertory grid exercise of the top $20$ (ranked by Alexa) social machines from our element set, against our consolidated constructs} \label{dendrogram}
+%\end{center}
+%\end{figure}
 
-From the dendrograms we can get a sense of similarities and separation of social machines. For
+From the dendrograms we can get a sense of similarities and differences between social machines. For
 instance, it is clear that YouTube, Vimeo, Reddit, and Digg are similar, which is not unexpected, due to their focus on sharing content.
@@
 There is also a dendrogram for the constructs, which can be used, as more data is collected, as an indicator for correlation between constructs.
-In this particular case, dendrogram suggests that there is a correlation between the `Single Participant Roles' and `No Hierarchy in Participant Roles' constructs, which is because they refer to the same aspect of social machines.
+In this particular case, our dendrogram suggests that there is a correlation between the `Single Participant Roles' and `No Hierarchy in Participant Roles' constructs, which is because they refer to the same aspect of social machines.
 Another, but less obvious connection is between the `Large variety of social features' and `Benefit NOT for the world/society', which requires further investigation.
 
-We envisage three key use cases for our framework.
+We envisage three use cases for our framework.
 First and foremost the framework forms the basis for a coherent and comprehensive description and classification of social machines.
@@
 In addition, it enables direct comparisons between systems, and provides a platform for terminologically consistent discussions about the field as a whole and specific instantiations.
-Finally, we also think that with more and more classification data becoming available, for instance, also through functionality that automatically monitors and computes the values of the some of the constructs introduced earlier, researchers will wish to apply prediction techniques (such as SVMs) to social machines, their behaviour, and impact.
+Finally, we also think that with more and more classification data becoming available, for instance, with functionality that automatically monitors and computes the values of the some of the constructs introduced earlier (see work on the Web Observatory REF), researchers will wish to apply prediction techniques (such as SVMs) to social machines, their behaviour, and impact.
 
 \section{Future work: building a social machine around the classification framework}
-In this paper we presented a first outline of a possible organisation structure for the design space of social machines.
-We introduced the key features of the notion of social machines, and explained how this notion differs from related areas such as collective intelligence, crowdsourcing and human computing.
+In this paper we presented a first outline of a description space for social machines.
+We introduced the key features of social machines, and explained how this concept differs from related areas such as collective intelligence, crowdsourcing, and human computing.
 We then discussed the basic constructs of a classification framework created through the repertory grid elicitation technique.
 In its current version the framework consists of $31$ constructs clustered according to the main components of social machines: social and machine-driven services, and the interactions between them.
-Our future work will mostly be dedicated to the revision of this initial list of constructs.
+Our future work will seek to revise this initial list of constructs.
 Most notably, we would like to add a number of constructs to emphasize the importance of the interaction aspect in the theory, engineering, and operation of social machines.
@@
 In addition, we plan to engage with the broader researcher community by setting up a social machine to facilitate discussions and feedback collection.
-One specific idea that is currently under development is to build a microtask-like environment, including specific game elements, in which participants are asked to provide answers to atomic challenges that rate and compare a pair of social machine instances according to a construct of our framework.
+One specific idea that is currently under development is to build a microtask environment, including specific game elements, in which participants are asked to provide answers to atomic challenges that rate and compare a pair of social machine instances according to a construct in our framework.
+This classificatory work is part of a larger project in which we seek to understand the theory and practice of building social machines and where we might in future anticipate what makes social machines scale or fail.
 

### 2013-04-04 latex/13socm07-shadbolt.tex
Updated to version submitted to sheridan

--- 2013-03-05 latex/paper_new.tex
@@
 %
-% Technical questions _only_ to
+% Technical questions _only_ tob
 % Gerald Murray (murray@hq.acm.org)
@@
 
-\documentclass{sig-alternate}
+\documentclass{www13-companion-accepted}
 
+\clubpenalty=10000 
+\widowpenalty = 10000
+
+\usepackage{ifpdf}
+\ifpdf
+\setlength{\pdfpagewidth}{8.5in}
+\setlength{\pdfpageheight}{11in}
+\else
+\fi
 \usepackage{url}
@@
 \alignauthor
-Nigel Shadbolt,^{1} Daniel Alexander Smith,^{2} Elena Simperl,^{3} Max Van Kleek,^{4} Yang Yang,^{5} Wendy Hall^{6}
+Nigel R. Shadbolt, Daniel A. Smith, Elena Simperl, Max Van Kleek, Yang Yang, Wendy Hall\\
        \affaddr{Web and Internet Science Group}\\
-       \affaddr{University of Southampton}\\
-       \affaddr{Southampton, UK}\\
-       \email{\{^{1}nrs, ^{2}ds, ^{3}es, ^{4}emax, ^{5}yy, ^{6}wh@ecs.soton.ac.uk}
+       \affaddr{University of Southampton, UK}\\
+       \email{\{nrs,ds,es6g12,emax,yy1402,wh\}@ecs.soton.ac.uk}
 }
-
 % There's nothing stopping you putting the seventh, eighth, etc.
@@
 
-The state of the art in human interaction with computational systems blurs the line between
-computations performed by machine logic and algorithms, and those that result from input by
-humans, arising from their own psychological processes and life experience. Current
-socio-technical systems, known as `social machines' exploit the large-scale interaction of
-humans with machines. Interactions that are motivated by numerous goals and purposes
-including financial gain, charitable aid, and simply for fun. In this paper we explore
-the landscape of social machines, both past and present, with the aim of defining an
-initial classificatory framework. Through a number of knowledge elicitation
-and refinement exercises we have identified the polyarchical relationship between
-infrastructure, social machines, and large-scale social initiatives. Our initial framework
-describes classification constructs in the areas of {\it contributions}, {\it participants}, and
-{\it motivation}. We present an initial characterization of some of the most popular social machines, as
-an initial validation and demonstration of the use of the identified constructs.
+The state of the art in human interaction with computational systems blurs the line between computations performed by machine logic and algorithms, and those that result from input by humans, arising from their own psychological processes and life experience.
+Current socio-technical systems, known as `social machines' exploit the large-scale interaction of humans with machines.
+Interactions that are motivated by numerous goals and purposes including financial gain, charitable aid, and simply for fun.
+In this paper we explore the landscape of social machines, both past and present, with the aim of defining an initial classificatory framework.
+Through a number of knowledge elicitation and refinement exercises we have identified the polyarchical relationship between infrastructure, social machines, and large-scale social initiatives.
+Our initial framework describes classification constructs in the areas of {\it contributions}, {\it participants}, and {\it motivation}.
+We present an initial characterisation of some of the most popular social machines, as demonstration of the use of the identified constructs.
 We believe that it is important to undertake an analysis of the behaviour and phenomenology of social machines, and of their growth and evolution over time.
@@
 % A category with the (minimum) three required fields
-%\category{H.4}{Information Systems Applications}{Miscellaneous}
+\category{H.1.2}{Information Systems}{Models and Principles}[User/Machine Systems]
 %A category including the fourth, optional field follows...
@@
 
-%\terms{Theory}
+%\terms{Social Machines, Web Science}
 
-%\keywords{ACM proceedings, \LaTeX, text tagging}
+\keywords{Social Machines;  Web Science}
 
-\section{Social machines: an introduction}
+\section{What Are Social Machines}
 Once upon a time `machines' were programmed by programmers and used by users.
@@
 When we fill out a form (e.g., health information, birds spotted in a garden, new construction sites on the way to work) the systems supporting this activity are minimal forms of social machines because the users are part of the `social computation', in this case the data creation and collection process facilitated through the site.
-This social component becomes richer when the database is curated by members of the broader community (e.g., Wikipedia) and when the social network adds value implicitly (e.g., Amazon) or explicitly (e.g., Facebook) to the overall system through individual or joint activities of the participants.
+This social component becomes richer when the database is curated by members of the broader community (e.g., Wikipedia) and when the social network adds value implicitly (e.g., Amazon) or explicitly (e.g., Facebook) to the overall system through the individual or joint activities of the participants.
 
 In general we still see a divide between conventional IT systems dedicated to data- and computation-intensive tasks, and Web $2.0$ sites offering some combination of well-known participatory features, in which user-generated content and the underlying social network evolve dynamically and hand-in-hand.
-However, as technology becomes more and more ubiquitous, many of the challenges we face will increasingly require solutions that rely on both of these axes: a sophisticated combination of data-intensive, complex automation, and deep community involvement.
+However, as technology becomes more and more ubiquitous, many of the challenges we face will increasingly require solutions that rely on both of these axes: a sophisticated combination of data-intensive, complex automation and deep community involvement.
 This suggests the need for new types of systems to tackle these emerging challenges, and these systems will not be able to be built and used in a sustainable way without a thorough understanding of the science and engineering of (the continuum of) social machines.
@@
 We aim to develop a descriptive framework to enable a systematic study of existing and future social machines.
-We anticipate that the framework will be a useful tool for both researchers in social and computer sciences, and for developers and operators of such systems.
+We anticipate that the framework will be a useful tool for both researchers in social and computer sciences, and for developers and operators of social machines.
 Using a common conceptualisation allows the former to become familiar with the landscape, and identify topics of research that so far have remained unexplored or under-explored.
-They are given a means to observe the effects of specific technical properties of a system configuration on social behaviour, discover design and evolution patterns, learn about social network formation and dynamics, and devise incentive mechanisms to encourage wide participation.
+They are given a means to observe the effects of specific technical properties of a system configuration on social behaviour, discover design and evolution patterns of such systems, learn about social network formation and dynamics, and devise incentive mechanisms to encourage wide participation.
 For developers and operators, the framework may inform the engineering of new systems in terms of critical features and community development.
@@
 In Section \ref{sec:framework} we present the framework in terms of the methodology used to devise it, the main clusters of constructs used for classification, and an example of using them to compare a selection of popular social machines via repertory grid analysis.
-We conclude with our proposed future work, including details concerning the evaluation of the framework and reference our plans to engage with the broader community to set up a `social machine' bringing together researchers from various disciplines to help formulate a widely agreed framework for the field.
+We conclude with our proposed future work, including details concerning the evaluation of the framework.
+Finally we reference our plans to engage with the broader community to set up a `social machine' bringing together researchers from various disciplines to help formulate a widely agreed classification framework for the field.
 
@@
 Turing machines were once conceived as logical automata.
-We have experienced a `paradigm shift' in the usage of computers, moving away from this narrow characterisation to one according to which machines facilitate a wide range of human interactions.
+We have experienced a `paradigm shift' in the usage of computers, moving away from this narrow characterisation to one where machines facilitate a wide range of human interactions.
 The emergence of `Computer-supported Collaborative Work (CSCW)' \cite{grudin1994computer} is representative of the early phases of this trend.
@@
 These developments led to amazing growth in terms of the amounts of content available online and the extent of mass participation.
-They are responsible for hundreds of millions of users all over the globe creating high-quality encyclopedias, publishing Terabytes of multimedia content, contributing to world-class software, and active participation in defining the agenda for many aspects of our society.
+They are responsible for hundreds of millions of users all over the globe creating high-quality encyclopaedias, publishing Terabytes of multimedia content, contributing to world-class software, and active participation in defining the agenda for many aspects of our society.
 This trend towards `prosumerism' is finding more and more adopters in the public and private sector.
@@
 In this context, a number of terms are used to refer to the ways people interact with each other and with applications: `wisdom of the crowds', `collective intelligence', `open innovation', `crowdsourcing', `human computation', and `social computing'.
-As we argue below, these terms are related, but not synonymous with `social machines' (see also Figure \ref{fig:socialmachine}).
+As we argue below they are related, but not synonymous with `social machines' (see also Figure \ref{fig:socialmachine}).
 
@@
 \begin{center}
-\includegraphics[width=0.5\textwidth]{img/socialmachinesscope_new.png}
+\includegraphics[width=0.5\textwidth]{img/Socialmachine1.png}
 \caption{Social machines and related areas} \label{fig:socialmachine}
@@
 In a social machine, human and computational intelligence coalesce in order to achieve a given purpose.
-While wisdom of the crowds and collective intelligence place their focus on identifying the situations in which groups of people perform better than individuals, social machines are concerned with the study and realisation of hybrid systems in which the two types of components co-exist.
-As such, theoretical and empirical insights from these two related areas are useful to understand the dynamics of the social structures underlying a social machine, but they are definitely not the only ingredient needed to build operational systems.
+Wisdom of the crowds (WoC) and collective intelligence (CI) clearly place their focus on identifying the situations in which groups of people perform better than individuals, social machines are concerned with the study and realisation of hybrid systems where the two types of components co-exist.
+As such, theoretical and empirical insights from WoC and CI are useful to understand the dynamics of the social structures underlying such a system, but they are definitely not the only ingredient needed to build operational social machines.
 The question of how human and automatic services can be brought together to achieve optimal results, as well as the actual engineering by which a system is developed, tested, and updated are equally important.
@@
 Human computation applies human processing power to tackle technical tasks that computers (still) find challenging \cite{von2009human}, typically in areas such as visual, audio, and natural language understanding.
-These kinds of task are an important part of today's crowdsourcing landscape, in particular on so-called `microtask' platforms such as Amazon's Mechanical Turk\footnote{\url{https://www.mturk.com/}} or CrowdFlower\footnote{\url{http://crowdflower.com/}}, which offer small financial rewards to an anonymous crowd engaged with atomic units of work that take in the range of seconds to minutes to complete.
+These sort of tasks are an important part of today's crowdsourcing landscape, in particular on so-called `microtask' platforms such as Amazon's Mechanical Turk\footnote{\url{https://www.mturk.com/}} or CrowdFlower\footnote{\url{http://crowdflower.com/}}, which offer small financial rewards to an anonymous crowd engaged with atomic units of work that take in the range of seconds to minutes to complete.
 The key difference between social machines and open innovation and crowdsourcing is at the level of the interaction between the social and the machine-driven processing components.
-The novelty these two other approaches bring in lies in their use of a much larger pool of human resources than traditional work environments \cite{quinn2011human}.
+The novelty these two other approaches bring in lies in their use of a much larger pool of human resources then traditional work environments \cite{quinn2011human}.
 There are clear socio-economic implications that their adopters need to deal with in order to optimally make use of this wealth of resources; technology may be needed to assist specific aspects of crowdsourcing projects, from evaluating and rewarding the results produced by the crowd to consolidating and aggregating them into a complete solution.
@@
 An analogous line of reasoning illustrates the overlap between social machines and social computing \cite{parameswaran2007research}.
-The latter is an area of computer science that refers to systems that support `the gathering, representation, processing, use, and dissemination of information that is distributed across social collectivities such as teams, communities, organisations, and markets' \cite{parameswaran2007research}.
+The latter is an area of computer science which refers to systems that support `the gathering, representation, processing, use, and dissemination of information that is distributed across social collectivities such as teams, communities, organisations, and markets' \cite{parameswaran2007research}.
 As such, compared to the general concept of `Computer-supported collaboration', social computing puts a greater emphasis on the information management capabilities of groups and communities, and less on the way these capabilities emerge as a joint effort.
@@
 that differentiate a set of elements -- for example, prompting the user to create a construct that
-differentiates between {\it GalaxyZoo}\footnote{In GalaxyZoo people classify galaxies with collective performance as good as professional astronomers, see \url{http://www.galaxyzoo.org}.} and {\it Facebook}, as two prominent example of social machines our classification framework should apply to.
-The user describes the opposing poles of the construct --- in this case, the user may decide on `For Science' and `For socialising' in order to capture the core distinction between the two system.
-The user then rates every element with value from $1$ to $5$ on this construct, where $1$ represents an element that is purely `For Science', and $5$ one that mainly serves `socialising' purposes.
-In triadic presentation the repertory grid software lists three elements and asks the user to generate a construct where one of the elements is different to the other two.
+differentiates between {\it GalaxyZoo}\footnote{In GalaxyZoo people classify galaxies with collective performance as good as professional astronomers, see: \\ \url{http://www.galaxyzoo.org}.} and {\it Facebook}, as two prominent example of social machines our classification framework should apply to.
+The user describes the opposing poles of the construct --- in this case, the user may decide on `For Science' and `For socialising' in order to capture the core distinction between the two systems.
+The user then rates every element with a value from $1$ to $5$ on this construct, where $1$ represents an element that is purely
+`For Science', and $5$ one that mainly serves `socialising' purposes.
+In triadic presentation the repertory grid software presents three elements and asks the user to generate a construct where one of the elements is different to the other two.
 We asked computer science researchers familiar with the field to create their own repertory grids, generating elements from their own knowledge, and creating constructs using this technique.
@@
 and $117$ different constructs (classifying factors).
-As the aim of this initial phase was to understand how people perceive the notion of a social machine and their most distinctive properties, we allowed the participants to choose the systems they are familiar with and describe them in their own terms.
+As the aim of this initial phase was to understand how people perceive the notion of social machine and their most distinctive properties, we allowed the participants to choose the social machines they are familiar with and describe them in their own terms.
 
@@
 This process involved four of the authors discussing the choices, and resulted in a consolidated set of $31$ constructs organised in four different clusters: popularity, tasks and purpose, participants and roles, and motivation and incentives.
-In our analysis we did not further consider the first cluster, which primarily covered descriptive information about a given system, such as its perceived level of maturity and current number of users.
+In our analysis we will not further consider the first cluster, which primarily covered descriptive information about a given system, such as its perceived level of maturity and current number of users.
 It was apparent that the constructs were not evenly distributed across clusters; whereas some aspects such as roles of participants occurred frequently (in various flavours) in the grids provided by the ten experts, others such as the different types of workflows synthesising human and automatised capabilities of a social machine were less common.
@@
 
+\vfill\eject
 \subsection{Constructs}
@@
 themes: first the nature of tasks, problems and activities
-performed through and using the machine, second the users and roles within
-each machine, and third, motivation for participation.
+performed through and using the machine, second sets of constructs pertaining to users and roles within
+each machine, third, motivation for participation.
 
@@
 
-\section{Future work: building a social machine around the classification framework}
+\section{Future work}
 In this paper we presented a first outline of a description space for social machines.
-We introduced the key features of social machines, and explained how this concept differs from related areas such as collective intelligence, crowdsourcing, and human computing.
+We introduced the key features of social machines, and explained how this concept differs from related areas such as collective intelligence, crowdsourcing and human computing.
 We then discussed the basic constructs of a classification framework created through the repertory grid elicitation technique.
@@
 Our future work will seek to revise this initial list of constructs.
-Most notably, we would like to add a number of constructs to emphasize the importance of the interaction aspect in the theory, engineering, and operation of social machines.
+Most notably, we would like to add a number of constructs to emphasise the importance of the interaction aspect in the theory, engineering, and operation of social machines.
 This includes constructs related to the human-computer workflows supported (e.g., human intelligence replacing automatic algorithms as in GWAPs vs the different bots available on Wikipedia complementing collaborative manual editing), the way inputs from the crowd are validated, aggregated and turned into a useful outcome, as well as issues related collaboration and coordination, and interface and communication design.
@@
 The SOCIAM Project is funded by the UK Engineering and Physical Sciences Research Council (EPSRC) under grant number EP/J017728/1 and comprises the Universities of Southampton, Oxford and Edinburgh.
+The authors would like to thank the entire SOCIAM team for their insightful feedback and lively discussions.
 

### 2013-04-04 latex/old_table.tex
Updated to version submitted to sheridan

\begin{table}[htb]
\begin{center}
\begin{tabular}{|p{8cm}|}
\hline
{\bf Tasks} \\
\hline
Variety of types of contributions \\
Domain specificity \\
Participation is done to `get something done' \\
Participation is done `for fun' \\
Participation is done to exchange knowledge \\
Participation is done to be social \\
Participants are intrinsically motivated \\
Participants are motivated by extrinsic reward \\
\hline
{\bf Users and Roles} \\
\hline
Generality of Audience \\
Participant autonomy \\
Participant anonymity \\
Extent of hierarchical organisation of roles \\
Clear separation of roles/responsibilities among participants \\
\hline
{\bf Participation} \\
\hline
Participants' (geographic) location is used by the service \\
Participation is done via: \\
\hspace{1cm} mobile devices \\
\hspace{1cm} web browsers \\
\hspace{1cm} apps \\
\hspace{1cm} sensors/sensing (location sensing and wearable devices) \\
Participation is done in a mobile context \\
\hline
{\bf Computation, Output and Benefit} \\
\hline
Output solves a computational problem \\
Output requires human insight \\
Output is a subjective appraisal \\
Participants directly benefit from participating \\
Service/Platform derives benefit from participant participation \\
Participants'' friends/social network benefit from participation \\
Participation is for the benefit of a specific person/group other than the participant or platform owner \\
Participation is for the benefit of society/the world at large  \\
\hline
\end{tabular}
\end{center}
\caption{Consolidated constructs of social machines.} \label{table:constructs}
\end{table}

### 2013-04-04 latex/paper_new.tex
Updated to version submitted to sheridan

@@
 \alignauthor
-Nigel Shadbolt,^{1} Daniel Alexander Smith,^{2} Elena Simperl,^{3} Max Van Kleek,^{4} Yang Yang,^{5} Wendy Hall^{6}
+Nigel R. Shadbolt, Daniel A. Smith, Elena Simperl, Max Van Kleek, Yang Yang, \\ Wendy Hall\\
        \affaddr{Web and Internet Science Group}\\
-       \affaddr{University of Southampton}\\
-       \affaddr{Southampton, UK}\\
-       \email{\{^{1}nrs, ^{2}ds, ^{3}es, ^{4}emax, ^{5}yy, ^{6}wh@ecs.soton.ac.uk}
+       \affaddr{University of Southampton, UK}\\
+       \email{\{nrs,ds,es6g12,emax,yy1402,wh\}@ecs.soton.ac.uk}
 }
-
 % There's nothing stopping you putting the seventh, eighth, etc.
@@
 \begin{center}
-\includegraphics[width=0.5\textwidth]{img/socialmachinesscope_new.png}
+\includegraphics[width=0.5\textwidth]{img/Socialmachine1}
 \caption{Social machines and related areas} \label{fig:socialmachine}
@@
 \begin{center}
-\includegraphics[width=0.4\textwidth]{img/polyarchy.pdf}
+\includegraphics[width=0.4\textwidth]{img/polyarchy}
 \caption{A polyarchy of social machines, illustrating the infrastructure and frameworks used by social machines, and machine-machine usage} \label{polyarchy}
@@
 \begin{center}
-\includegraphics[width=0.5\textwidth]{img/dendrogram-both.png}
+\includegraphics[width=0.5\textwidth]{img/dendrogram-both}
 \caption{Dendrograms of constructs, in blue, and elements (social machines), in red, from a repertory grid exercise of the top $20$ (ranked by Alexa) social machines from our element set, against our consolidated constructs.} \label{dendrogram}

### 2013-04-04 latex/old2/13socm07-shadbolt.tex
Merge of papers to final version with edits sent to Sheridan

--- 2013-04-04 latex/paper_new.tex
@@
 
-\title{Towards a classification framework for social machines}
+\title{Towards a Classification Framework for Social Machines}
 %\subtitle{[Extended Abstract]
@@
 \bibliographystyle{abbrv}
-\bibliography{sigproc} % sigproc.bib is the name of the Bibliography in this case
+%\bibliography{sigproc} % sigproc.bib is the name of the Bibliography in this case
 % You must have a proper ".bib" file
@@
 
+\begin{thebibliography}{10}
+
+\bibitem{bernerslee1999}
+T.~Berners-Lee and M.~Fischetti.
+\newblock {\em Weaving the web: The original design and ultimate destiny of the
+  world wide web by its inventor}.
+\newblock Harper, San Francisco, 1999.
+
+\bibitem{chesbrough2003}
+H.~Chesbrough.
+\newblock {\em {Open innovation: The new imperative for creating and profiting
+  from technology}}.
+\newblock 2003.
+
+\bibitem{grudin1994computer}
+J.~Grudin.
+\newblock Computer-supported cooperative work: History and focus.
+\newblock {\em Computer}, 27(5):19--26, 1994.
+
+\bibitem{howe2006crowdsourcing}
+J.~Howe.
+\newblock Crowdsourcing: A definition.
+\newblock {\em Crowdsourcing: Tracking the rise of the amateur}, 2006.
+
+\bibitem{kelly}
+G.~A. Kelly.
+\newblock {The Psychology Of Personal Constructs}.
+\newblock 1955.
+
+\bibitem{kuznetsov2006motivations}
+S.~Kuznetsov.
+\newblock Motivations of contributors to wikipedia.
+\newblock {\em ACM SIGCAS computers and society}, 36(2):1, 2006.
+
+\bibitem{lakhani2003hackers}
+K.~Lakhani and R.~Wolf.
+\newblock Why hackers do what they do: Understanding motivation and effort in
+  free/open source software projects.
+\newblock 2003.
+
+\bibitem{malone2009harnessing}
+T.~Malone, R.~Laubacher, and C.~Dellarocas.
+\newblock Harnessing crowds: Mapping the genome of collective intelligence.
+\newblock 2009.
+
+\bibitem{obamakieron}
+K.~O'Hara.
+\newblock {Social Machine Politics Are Here to Stay}.
+\newblock {\em {IEEE Internet Computing}}, 17(2), March/April 2013.
+
+\bibitem{parameswaran2007research}
+M.~Parameswaran and A.~B. Whinston.
+\newblock Research issues in social computing.
+\newblock {\em Journal of the Association for Information Systems},
+  8(6):336--350, 2007.
+
+\bibitem{quinn2011human}
+A.~J. Quinn and B.~B. Bederson.
+\newblock Human computation: a survey and taxonomy of a growing field.
+\newblock In {\em Proceedings of the 2011 annual conference on Human factors in
+  computing systems}, pages 1403--1412. ACM, 2011.
+
+\bibitem{knowledgeelicitation}
+N.~R. Shadbolt and A.~M. Burton.
+\newblock {Knowledge Elicitation}.
+\newblock {\em Evaluation of Human Work: Practical Ergonomics Methodology},
+  1990.
+
+\bibitem{surowiecki2005wisdom}
+J.~Surowiecki.
+\newblock {\em The wisdom of crowds}.
+\newblock Anchor, 2005.
+
+\vfill\eject
+
+\bibitem{von2009human}
+L.~Von~Ahn.
+\newblock Human computation.
+\newblock In {\em Design Automation Conference, 2009. DAC'09. 46th ACM/IEEE},
+  pages 418--419. IEEE, 2009.
+
+\bibitem{tagging}
+A.~Zollers.
+\newblock Emerging motivations for tagging: Expression, performance, and
+  activism.
+\newblock In {\em Workshop on Tagging and Metadata for Social Information
+  Organization, held at the 16th International World Wide Web Conference},
+  2007.
+
+\end{thebibliography}
+
 \balancecolumns % GM June 2007

### 2013-04-04 latex/old2/old/paper_new.tex
Merge of papers to final version with edits sent to Sheridan

--- 2013-02-26 latex/paper_new.tex
@@
 
-\begin{figure}[htb]
-\begin{center}
-\includegraphics[width=0.4\textwidth]{img/dendrogram-constructs.png}
-\caption{Dendrogram of constructs from a repertory grid exercise of the top $20$ (ranked by Alexa) social machines from our element set, against our consolidated constructs} \label{dendrogram}
-\end{center}
-\end{figure}
+%\begin{figure}[htb]
+%\begin{center}
+%\includegraphics[width=0.2\textwidth]{img/dendrogram-elements.png}
+%\caption{Dendrogram of elements (social machines) from a repertory grid exercise of the top $20$ (ranked by Alexa) social machines from our element set, against our consolidated constructs} \label{dendrogram}
+%\end{center}
+%\end{figure}
+%
+%\begin{figure}[htb]
+%\begin{center}
+%\includegraphics[width=0.4\textwidth]{img/dendrogram-constructs.png}
+%\caption{Dendrogram of constructs from a repertory grid exercise of the top $20$ (ranked by Alexa) social machines from our element set, against our consolidated constructs} \label{dendrogram}
+%\end{center}
+%\end{figure}
 

### 2013-04-04 latex/old2/sig-alternate.tex
Merge of papers to final version with edits sent to Sheridan

--- 2013-02-15 latex/sig-alternate.tex
@@
 
-Wikipedia
+%Wikipedia
+%
+%GalaxyZoo (?)
+%
+%Ushahidi
+%
+%Facebook
+%
+%MechanicalTurk
+%
+%GWAPs
 
-GalaxyZoo (?)
+In order to demonstrate usage of our framework's constructs, we ranked our social machines
+by their Alexa ranking, and performed a full repertory grid elicitation exercise. The
+results are illustrated in a grid and accompanying dendrograms for element and construct
+similarity in Figure~\ref{dendrogram}.
 
-Ushahidi
+\begin{figure*}
+\begin{center}
+\includegraphics[width=18cm]{img/dendrogram.png}
+\caption{Dendrogram from a repertory grid exercise of the top 20 (ranked by Alexa) social machines from our element set, against our consolidated constructs.} \label{dendrogram}
+\end{center}
+\end{figure*}
 
-Facebook
-
-MechanicalTurk
-
-GWAPs
+[needs some discussion on what this shows.]
 
